### 2020-07-10 hpp.tex
Journal paper about HPP software.

\section{Introduction}
This \paper describes a software platform taylored for manipulation planning.
The main contributions are:
\begin{itemize}
\item an original and general implementation of prehensile manipulation based on
  non-linear constraints,
\item an original solver for non-linear constraints that is able to handle implicit and explicit constraints,
\item a manipulation planning algorithm that is able to tackle a great variety of manipulation planning problems,
\item an open-source software suite that implements all the above following state of the art development tools and methods.
\end{itemize}

The \paper is organized as follows.
\partie~\ref{partie:preliminary} introduces some preliminary notions like kinematic chains and Lie groups that are used to model the configuration space of each joint.

Each \partie is implemented by one or several software packages. Instead of providing formulas for some values need to be computed, we provide a link to the C++ implementation.

\section{Preliminaries}\label{partie:preliminary}
\subsection{Kinematic chains and Lie groups}

A kinematic chain is commonly understood as a set of rigid-body links connected
to each other by joints. Each joint has one degree of freedom either in
rotation or in translation. A configuration of the kinematic chain is represented by a vector. Each component of the vector represents the angular or linear value of the correspondig joint.

This representation although well suited for fixed base manipulator
arms is not well adapted for robots with a mobile base like wheeled
mobile robots, aerial robots or legged robots, since the mobility of the base
cannot be correctly represented by translation or rotation joints.
Representing a freeflying object by three virtual translations followed by
three virtual rotations called roll, pitch and yaw is indeed a poor workaround.

\subsubsection{Kinematic chain}

A \textit{kinematic chain} is a tree of \textit{joints} where each
joint represents a mobility of a rigid-body link with respect to
another link or with respect to the world reference frame. To each joint is
associated a configuration space called the \textit{joint configuration space}.
The most common joints with their respective configuration spaces are
\begin{itemize}
\item linear \textit{translation} with configuration space $\reals$,
\item bounded \textit{rotation} with configuration space  $\reals$,
\item unbounded \textit{rotation} with configuration space  $SO(2)$,
\item \textit{planar} with configuration space $SE(2)$,
\item \textit{freeflyer} with configuration space $SE(3)$.
\end{itemize}
$SO(n)$ and $SE(n)$ respectively stand for \textit{special orthogonal group} and \textit{special Euclidean group}. They represents the group of rigid-body transformations and the group of rotations in $\reals^n$.

\subsubsection{Lie groups}

The joint configuration spaces enumerated in the previous paragraph: $\mathbb{R}^n$, $SE(n)$ are all Lie groups. The group operation is $+$ for $\mathbb{R}^n$, and composition denoted as $"."$ for $SE(n)$. We refer to~\cite{LMS94} -- Appendix A for a thorough definition of Lie groups. We expose here only properties that are useful for the sequel of the \paper.

For any Lie group $\liegroup$ with neutral element $\neutral$, the tangent space at the neutral element $T_{\neutral}\liegroup$ of the group naturally maps to the tangent space at any point of the group. This means that any \textit{velocity} $\mathbf{v}\in T_{\neutral}\liegroup$ uniquely defines
\begin{enumerate}
\item a velocity $\mathbf{w}\in T_{g}\liegroup$ at any point $g$ of the group, and thus,
\item a vector field on the tangent space $T\liegroup$, and
\item by integration during unit time of the latter vector field, starting from the origin, a new point $g_{1}\in\liegroup$.
\end{enumerate}
Item 1 above is called the \textit{transport} of velocity $\mathbf{v}$ to $g$.
Item 3 above is called the exponential map of $\liegroup$ and is denoted by $\exp$.

\paragraph{Geometric interpretations}
\begin{itemize}
\item $\reals$ (and by trivial generalization $\reals^n$): the neutral element is $0$. The tangent space at 0 is isomorphic to $\reals$ and
  $$
  \forall \theta\in\reals,\  \exp(\theta) = \theta.
  $$
\item $SE(3)$: an element $g$ of $SE(3)$ can be seen as the position of a moving frame in a fixed reference frame. A point $\mathbf{x}\in\reals^3$ is mapped to $g(\mathbf{x})$. Note that $\mathbf{x}$ is also the coordinate vector of $g(\mathbf{x})$ in the moving frame $g$. If $\mathbf{v}$, $\omega$ are a linear and an angular velocities at the origin, $(\mathbf{v},\omega)$ is transported to $g$
  as the same linear and angular velocities expressed in the moving frame. In other words, if
  \begin{equation}\label{eq:homogeneous matrix}
  M=\left(\begin{array}{ll} R & \mathbf{t}\\ 0 & 1\end{array}\right)
    \end{equation}
  with $R\in SO(3)$ and $\mathbf{t}\in\reals^3$  is the homogeneous matrix
  representing $g$, and $(\mathbf{v},\mathbf{\omega})$ is a velocity in $T_{I_3}SE(3)$, the velocity transported to $g$ corresponds to the linear velocity $R\mathbf{v}$ and $R\omega$ of the moving frame.
  Integral curves of the vector field mentioned in item 2 above correspond to helicoidal motions of constant velocity expressed in the moving frame.
\end{itemize}
$SE(2)$, $SO(3)$, and $SO(2)$ are (or are isomorphic to) subgroups of $SE(3)$ and follow the same geometrical interpretation.

\paragraph{Vector representations}
Each Liegroup element is represented by a vector. Rotations are represented by
unit quaternions.

Therefore elements of $SE(3)$ are represented by a vector in $\reals^7$ where the 3 first components represent the image of the origin (vector $\mathbf{t}$ in Equation~\ref{eq:homogeneous matrix}), the 4 last components represent a unit quaternion.

Elements of $SO(3)$ are represented by a unit vector of dimension 4.

Elements of $SE(2)$ are represented by a vector of dimension 4. The 2 first components represent the image of the origin. The 2 last components represent the cosine and sine of the rotation angle. Therefore the homogeneous matrix associated to $\conf=(q_1,q_2,q_3,q_4)$ is
$$
M = \left(\begin{array}{lll}
  q_3 & -q_4 & q_1 \\
  q_4 &  q_3 & q_2 \\
  0   &  0  & 1
\end{array}\right).
$$
Table~\ref{tab:vector representation} compiles this information.
\begin{table}
  \begin{tabular}{|l|c|c|}
    \hline
    Lie group type & configuration & velocity\\
    \hline
    $SE(3)$ & $(x_1,x_2,x_3,p_1,\cdots,p_4)$ & $\dconf=(\mathbf{v},\omega)$ \\
            & $\in\reals^7$                 & $\in\reals^6$ \\
    $SE(2)$ & $(x_1,x_2,\cos\theta,\sin\theta)$ & $\dconf=(\mathbf{v},\dot{\theta})$\\
            & $\in\reals^4$                 & $\in\reals^3$ \\
    $SO(3)$ & $(p_1,p_2,p_3,p_4)\in\reals^4$ & $\dconf=\omega\in\reals^3$\\
    $SO(2)$ & $(\cos\theta,\sin\theta)\in\reals^2$ & $\dconf=\dot{\theta}\in\reals$\\
    \hline
  \end{tabular}
  \caption{Main Lie group types and their vector representations. Notice that the dimensions of the configuration representation and of the velocity representation may differ.}
  \label{tab:vector representation}
\end{table}

\paragraph{Exponential map}

As expressed earlier, following a constant velocity $\dot{\conf}$ from the neutral element of a joint configuration space leads to another configuration denoted as
$$
\conf = \exp(\dconf).
$$
In some cases, we may specify in subscript the Lie group that is used: \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/explog.hpp#L34}{$\exp_{SO(3)}$}, \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/explog.hpp#L236}{$\exp_{SE(3)}$}.

For all Lie groups $\reals$, $SO(n)$, $SE(n)$, the exponential map is surjective. This means that for any $\conf\in\liegroup$, there exists $\mathbf{v}\in T_{\neutral}\liegroup$, such that $\conf = \exp(\mathbf{v})$.
Although $\exp$ is not injective, choosing the smallest norm $\mathbf{v}$ uniquely defines function $\log$ from $\liegroup$ to $T_{\neutral}\liegroup$, up to some singularities where several candidates $\mathbf{v}$ are of equal norms.
Again, we may specify the Lie group that is used: \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/log.hxx#L112}{$\log_{SE(3)}$}, \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/log.hxx#L15}{$\log_{SO(3)}$}

\paragraph{Sum and difference notations}

Following a constant velocity $\dconf\in T_{\neutral}\liegroup$ starting from $\conf_0\in\liegroup$, leads to
$$
\conf_1 = q_0.\exp(\dconf).
$$
Note that if $\liegroup=\reals$, we write
$$
\conf_1 = \conf_0 + \dconf,
$$
since the Lie group operator of $\reals$ is $+$ and $\exp_{\reals}$ is the identity.
In order to homogeneize notation, we define the following operators:
\begin{eqnarray}\label{eq:plus}
  \conf_0 \oplus \dconf &\triangleq& \conf_0.\exp(\dconf),\\
  \label{eq:minus}
  \conf_1 \ominus \conf_0 &\triangleq& \log(\conf_0^{-1}.\conf_1).
\end{eqnarray}

\subsubsection{Robot configuration space}

Given a kinematic chain with joints $(J_1, \cdots, J_{njoints})$, ordered in such
a way that each joint has an index bigger than its parent in the tree, the configuration space of the robot is the cartesian product of the joint configuration spaces.
$$
\CS \triangleq \CS_{J_1}\times \cdots \times \CS_{J_{njoints}}.
$$
$\CS$ naturally inherits the Lie group structure of the joint configuration spaces through the Cartesian product.
We denote by $nq_i$, $nv_i$ the sizes of the configuration and velocity vector representations of joint $J_i$, as defined in Table~\ref{tab:vector representation}.
The configuration and velocity of the robot can thus be represented by vectors of size $nq$ and $nv$ such that
$$
nq = \sum_{i=1}^{njoints} nq_i, \ \ \ nv = \sum_{i=1}^{njoints} nv_i
$$
We denote by $iq_i$, and $iv_i$ the indices of joint $i$ in the robot configuration and velocity vectors.
$$
iq_i = \sum_{j=1}^{i-1} nq_j\ \ \ \ iv_i = \sum_{j=1}^{i-1} nv_j
$$
With these definitions and notation, the linear interpolation between two robot
configurations $\conf_0$ and $\conf_1$ is naturally written:
$$
\conf(t) = \conf_0 \oplus t (\conf_1 - \conf_0)
$$
This formula generalizes the linear interpolation to robots with freeflying bases, getting rid of singularities of roll -- pitch -- yaw parameterization.
%% begin impl
Cartesian products of Lie groups are represented by class \href{https://gepettoweb.laas.fr/hpp/hpp-pinocchio/doxygen-html/classhpp_1_1pinocchio_1_1LiegroupSpace.html}{\texttt{LiegroupSpace}}. Elements of these spaces are represented by classes
\begin{itemize}
\item \href{https://gepettoweb.laas.fr/hpp/hpp-pinocchio/doxygen-html/classhpp_1_1pinocchio_1_1LiegroupElementBase.html}{\texttt{LiegroupElement}}
\item \href{https://gepettoweb.laas.fr/hpp/hpp-pinocchio/doxygen-html/classhpp_1_1pinocchio_1_1LiegroupElementBase.html}{\texttt{LiegroupElementRef}}, and
\item \href{https://gepettoweb.laas.fr/hpp/hpp-pinocchio/doxygen-html/classhpp_1_1pinocchio_1_1LiegroupElementBase.html}{\texttt{LiegroupElementConstRef}}.
  \end{itemize}
%% end impl

\section{Non-linear constraints and solvers}

Some tasks require the robot to enforce some non-linear constraints. Foot contact on the ground for a humanoid robot, center of mass projection on a horizontal plane, gaze constraint are a few examples.

\subsection{Non-linear constraints}

\begin{definition}\textbf{Non-linear constraint.} A non linear constraint is defined by a differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and is written
\begin{equation}\label{eq:constraint}
h(\conf) = 0.
\end{equation}
\end{definition}
If the robot is subject to several numerical constraints, $h_1,\cdots,h_k$ with values in $\reals^{m_1}\cdots\reals^{m_k}$, these constraints are equivalent to a single constraint $h$ with values in $\reals^{m}$, where $m=\sum_{i=1}^k m_i$, such that
$$
h(\conf) \triangleq \left(\begin{array}{c} h_1(\conf) \\ \vdots \\ h_k(\conf)
\end{array}\right).
$$
It may be useful to use a non zero right hand side for the same function $h$. We define for that parameterized non-linear constraints.
\begin{definition}\textbf{Parameterized non-linear constraints.}
  A parameterized non-linear constraint is defined by a differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and by a vector $\mathbf{h}_0$ of $\reals^m$ and is written
  $$
  h(\conf) = \mathbf{h}_0.
  $$
\end{definition}
%% begin impl
Differentiable functions are represented by class\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1DifferentiableFunction.html}{\texttt{DifferentiableFunction}}.
%% end impl

\paragraph{Jacobian}

In this \paper, we will make use of the term Jacobian in a generalized way.
If $h$ is a differentiable function from a Lie group $\liegroup_1$ to a Lie group $\liegroup_2$, and $\conf_1$ an element of $\liegroup_1$, we will denote by $\frac{\partial h}{\partial\conf}(\conf_1)$ the operator that maps velocities in $T_{\conf_1}\liegroup_1$ to the corresponding velocity of the image by $h$ in $T_{h(\conf_1)}\liegroup_2$.

This operator is represented by a matrix with $nv_2$ lines and $nv_1$ columns, where $nv_1$ and $nv_2$ are the dimensions of the tangent spaces of respectively $\liegroup_1$ and $\liegroup_2$.

\subsection{Newton based solver}

It is sometimes useful to produce a configuration $\conf$ that satisfies a constraint (or a set of constraints) of type~(\ref{eq:constraint}) from a configuration $\conf_0$ that does not. This action is called the \textit{projection of $\conf_0$ onto the submanifold defined by the constraint} and is performed by a numerical solver that iteratively linearizes the constraint as follows:
$$
h(\conf_{i+1}) \approx h(\conf_i) + \frac{\partial h}{\partial \conf}(\conf_{i})(\conf_{i+1}-\conf_i) = 0
$$
Iterate $\conf_{i+1}$ is computed as follows:
\begin{equation}\label{eq:HierarchicalIterative}
\conf_{i+1} = \conf_i-\alpha_i\frac{\partial h}{\partial \conf}^{+}(\conf_{i}) h(\conf_i)
\end{equation}
where $.^{+}$ denotes the Moore Penrose pseudo inverse, and $\alpha_i$ is a positive real number called the step size. Taking $\alpha_i=1$ exactly solves the linear approximation, but it may not be the best choice in general.

The computation of $\alpha_i$ is performed by a line search algorithm.
The algorithm stops when the norm of each $h_i(\conf_{i+1})$ is below a given error threshold.
%% begin impl
Class\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1HierarchicalIterative.html}{\texttt{HierarchicalIterative}} implements the above Newton method. Several linesearch methods are implemented: \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1Backtracking.html}{\texttt{Backtracking}}, \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1ErrorNormBased.html}{\texttt{ErrorNormBased}}, \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1FixedSequence.html}{\texttt{FixedSequence}}, and \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1Constant.html}{\texttt{Constant}}.
Note that to define a new constraint, the user needs to derive class \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1DifferentiableFunction.html}{\texttt{DifferentiableFunction}} and to implement methods \texttt{impl\_compute} and \texttt{impl\_jacobian}.
%% end impl

\subsection{Explicit constraints}

In manipulation planning applications, where robots manipulate objects, once an object is grasped, the position of the object can be computed explicitely from the configuration of the robot. In this case, some configuration variables of the system depends on other:
$$
\conf=(\conf_{rob},\conf_{obj})\in\CS,\ \ \ \ \conf_{obj} = g_{grasp}(\conf{rob}).
$$
Although this constraint may fit definition~(\ref{eq:constraint}) by defining
\begin{equation}\label{eq:explicit-grasp}
h(\conf) \triangleq \conf_{obj} \ominus g_{grasp}(\conf_{rob}),
\end{equation}
solving this constraint possibly with other constraints using an iterative scheme~(\ref{eq:HierarchicalIterative}) is obviously sub-optimal.
Notice that the right hand side of~(\ref{eq:explicit-grasp}) is the $\log$ of the pose of the object when grasped, with respect to the actual pose of the object.

More generally, let us denote by
\begin {itemize}
\item $I_{nq}$ the set of positive integers not greater than $nq=\dim\CS$,
\item $I$ a subset of $I_{nq}$,
\item $\bar{I}$ the complement in $I_{nq}$ of $I$,
\item $|I|$ the cardinal of $I$.
\end {itemize}
If $\conf\in\CS$ is a configuration, we denote by $\conf_{I}\in\reals^{|I|}$ the vector composed of the components of $\conf$ of increasing indices in $I$.
\paragraph {Example} if $\conf = (q_1,q_2,q_3,q_4,q_5,q_6,q_7)$ and $I=\{1,2,6\}$, then $\conf_I=(q_1,q_2,q_6)$, $\conf_{\bar{I}}=(q_3,q_4,q_5,q_7)$.

Similarly, if
\begin{itemize}
\item $m$ and $n$ are two integers,
\item $M$ and $N$ are two subsets of respectively $I_{m}$ and $I_{n}$,
\item $J$ is a matrix with $m$ rows and $n$ columns,
\end{itemize}
we denote by
\begin{equation}\label{eq:sub-matrix}
  J_{M,N}
\end{equation}
the matrix of size $|M| \times |N|$ obtained by extracting the rows of $J$ of indices in $M$ and the columns of $J$ with indices in $N$.

\paragraph{Example}
If $m=3$, $n=4$, $M=\{2,3\}$ and $N=\{1,2,4\}$,
$$
J = \left(\begin{array}{cccc}
  J_{1,1} &   J_{1,2} &   J_{1,3} &   J_{1,4}\\
  J_{2,1} &   J_{2,2} &   J_{2,3} &   J_{2,4}\\
  J_{3,1} &   J_{3,2} &   J_{3,3} &   J_{3,4}\\
  J_{4,1} &   J_{4,2} &   J_{4,3} &   J_{4,4}
\end{array}\right)
$$
then
$$
J_{M\times N} = \left(\begin{array}{ccc}
  J_{2,1} &   J_{2,2} &   J_{2,4}\\
  J_{3,1} &   J_{3,2} &   J_{3,4}
\end{array}\right)
$$

\begin{definition}\label{def:Explicit}
An explicit constraint $E=(in, out, f)$ is a mapping from $\CS$ to $\CS$, defined by the following elements:
\begin{itemize}
\item a subset of input indices $in\subset\{1,\cdots, nq\}$,
\item a subset of output indices $out\subset\{1,\cdots, nq\}$,
\item a smooth mapping $f$ from $\reals^{|in|}$ to $\reals^{|out|}$,
\end{itemize}
satisfying the following properties:
\begin{itemize}
\item $in\cap out = \emptyset$,
\item for any $\p\in\CS$, $\conf = E(\p)$ is defined by
  \begin{align*}
    &\conf_{\bar{out}} = \p_{\bar{out}}\\
    &\conf_{out} = f (\p_{in}).
  \end{align*}
\end{itemize}
\end{definition}

\subsection{Solver by substitution}

To optimize constraint resolution, we perform variable substitution when possible in order to reduce both the number of variables and the dimension of the resulting implicit constraint. We describe here the method that has been first published in~\cite{mirabel:hal-01804774}. Unlike in the former paper, the description we give in Algorithm~\ref{alg:ExplicitConstraintSet} is closer to the real implementation. Some links to the source code are indeed provided in the algorithm description.
\begin{algorithm}
  \begin{algorithmic}[1]
    \Procedure{initializeSolver}{}
    \State{$explicit$} $\gets$ empty vector of explicit constraints
    \State{$nc \gets 0$}
    \State{$args \gets$ array of size $nq$ filled with -1}
    \EndProcedure
    \Function{add}{$E=(in, out, f)$}
    \If{$args_{out}$ contains an element $\geq 0$}
    \State{return failure}
    \EndIf
    \State{queue $idxArg \gets$ elements of $in$}
    \While{$idxArg$ \textbf{not} empty}
    \State{$iArg \gets idxArg$ first element}
    \State{remove $idxArgs$ first element}
    \If{$iArg\in out$}
    \State{return failure}
    \EndIf
    \If{$args[iArg] == -1$}
    \State{continue}
    \Else
    \State{push $explicit[args[iArg]].in$ elements into $idxArg$}
    \EndIf
    \EndWhile
    \State{fill $args_{out}$ with $nc$}
    \State{explicit.add($E$); $nc\gets nc+1$}
    \State{computeOrder()}
    \State{\Return{success}}
    \EndFunction
  \end{algorithmic}
  \caption{Insertion of an explicit constraint in the solver. Line 1 is called once at initialization of the solver. $explicit$ is a vector that stores the constraints that are successfully added to the solver. $nc$ is the size of the latter. $args$ is an array that stores for each configuration variable, the index in $explicit$ of the constraint that computes this configuration variable, -1 if no constraints computes the index. Procedure \href{https://github.com/humanoid-path-planner/hpp-constraints/blob/2555dfec945575f824bd973c30c5c13ec3b67645/src/explicit-constraint-set.cc\#L144}{ADD} tests whether explicit constraint $E$ is compatible with the previously inserted constraints. Line 6 checks whether any output variable of $E$ is already computed by a previous explicit constraint. If so the procedure returns failure and the $E$ is not inserted. The Loop at line 9 recursively checks that any element of $out$ is not an input variable of a previously inserted constraint. If the loop ends without returning failure,  line 18 stores that elements of $out$ are computed by $E$ and $E$ is inserted in vector of constraints. Function \href{https://github.com/humanoid-path-planner/hpp-constraints/blob/2555dfec945575f824bd973c30c5c13ec3b67645/src/explicit-constraint-set.cc\#L301}{computeOrder} at line 20 recursively computes the order in which the explicit constraints are evaluated, following the rule that the input of a constraint should be evaluated before the output.}
  \label{alg:ExplicitConstraintSet}
\end{algorithm}
Once several compatible explicit constraints have been inserted in the solver,
they behave as a single one. For instance, if $\conf=(\conf_1,\conf_2,\conf_3)$,
$$
  \left\{\begin{array}{lcl}\conf_1 &=& f_1(\conf_2)\\
  \conf_2 &=& f_2(\conf_3)\end{array}\right. \mbox{ becomes }
  \left[\begin{array}{cc}\conf_1\\ \conf_2\end{array}\right] =
    \left[\begin{array}{cc}f_1(f_2(\conf_3))\\ f_2(\conf_3)\end{array}\right],
$$
and $f_2$ should be evaluated before $f_1$.

\paragraph{Substitution} When an explicit constraint is not successfully added following Algorithm~\ref{alg:ExplicitConstraintSet}, it is handled as an implicit constraint. Therefore, after inserting implicit and explicit constraints, the solver stores a system of equations equivalent to one explicit and one implicit constraints that we denote by:
\begin{eqnarray}\label{eq:substitution-1}
  h(\conf_{in},\conf_{out}) &=& 0, \\
  \label{eq:substitution-2}
  \conf_{out} &=& f(\conf_{in}), \mbox{ where }\\
  \label{eq:substitution-3}
  \conf &=& (\conf_{in},\conf_{out}).
\end{eqnarray}
Substituting~(\ref{eq:substitution-2}) into~(\ref{eq:substitution-1}), we define an implicit constraint on $\conf_{in}$ only:
$$
\tilde{h}(\conf_{in}) \triangleq h(\conf_{in}, f(\conf_{in})) = 0
$$
The \textit{solver by substitution} applies iteration~(\ref{eq:HierarchicalIterative}) to $\tilde{h}$, instead of $h$. We need therefore to compute the Jacobian of  $\tilde{h}$:
$$
\frac{\partial \tilde{h}}{\partial \conf_{in}}  =
\frac{\partial h}{\partial \conf_{in}} + \frac{\partial h}{\partial \conf_{out}}.
\frac{\partial f}{\partial \conf_{in}}.
$$
As the Jacobian of $h$ is provided with the implicit constraint, we need to compute $\frac{\partial f}{\partial \conf_{in}}$. Let us recall that $f$ may be the combination of several compatible explicit constraints. Let us denote by $E$ the mapping from $\CS$ to $\CS$ associated to $f$ by Definition~\ref{def:Explicit}.
Let $J$ denote the $nv\times nv$ Jacobian matrix of $E$. Then $J$ is defined by blocks as follows:
\begin{equation}\label{eq:jacobian-explicit}
\begin{array}{cc}
  J_{{in}\times{in}} = I_{|{in}|} & J_{{in}\times{out}} = 0 \\
  J_{{out}\times{in}} = \frac{\partial f}{\partial \conf_{in}} & J_{{out}\times{out}} = 0
\end{array}
\end{equation}
If $E$ is the composition of several explicit constraints $E_i=(in_i,out_i,f_i)$ of Jacobian $J_i$, $i\in I_{nc}$, for an integer $nc$, then
\begin{equation}\label{eq:jacobian-product}
J = \prod_{i=nc}^{1} J_{i},
\end{equation}
with $J_i$ obtained by expression~(\ref{eq:jacobian-explicit}) after replacing $in$, $out$, and $f$ by $in_i$, $out_i$, and $f_i$.

$\frac{\partial f}{\partial\conf_{in}}$ is then obtained by extracting from $J$ block $out\times in$.

Let us now detail the iterative computation of~(\ref{eq:jacobian-product}). Let $J$ be the product of of $J_j$ for $j$ from $nc$ to $i+1$.
Note that if $J_i$ and $J$ are square matrices of size $nv$, of the form~(\ref{eq:jacobian-explicit}), $J_i.J$ can be computed by block as follows:
\begin{eqnarray*}
  (J_{i}.J)_{in_{i}\times I_{nv}} &=& J_{in_{i}\times I_{nv}}\\
  (J_{i}.J)_{out_{i}\times I_{nv}} &=& \frac{\partial f_i}{\partial\conf_{in_{i}}}.J_{in_{i}\times I_{nv}}\\
\end{eqnarray*}
and as columns $out$ of $J$ are equal to $0$, left multiplying $J$ by $J_i$ consists in modifying only the following block of $J$:
$$
  (J_{i}.J)_{out_{i}\times in} = \frac{\partial f_i}{\partial\conf_{in_{i}}}.J_{in_{i}\times {in}}\\
$$
Other coefficients of $J_{i}J$ are equal to the corresponding coefficients of $J$.
An implementation of the above Jacobian product can be found \href{https://github.com/humanoid-path-planner/hpp-constraints/blob/e21490c8c713949bd3038dccb6fe02cf254a615f/src/explicit-constraint-set.cc#L267}{here}.

% begin impl
The solver by substitution described in this section is implemented by class \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}, that store an instance of \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1ExplicitConstraintSet.html}{\texttt{ExplicitConstraintSet}}.

\paragraph{Important remark} As mentioned in Table~\ref{tab:vector representation}, the configuration and velocity vectors may have different sizes. As a consequence, index sets $in$ and $out$ in Definition~\ref{def:Explicit} correspond to configuration vector indices, while in Expression~(\ref{eq:jacobian-explicit}) they correspond to velocity vector indices. To keep notation simple, we use the same notation for different sets.

\subsection{Constrained path}

Now that we are able to project configurations onto submanifolds defined by numerical constraints -- up to some numerical threshold, we need now to define paths on such submanifolds. The usual way of doing so is to discretize the path and to project each sample configuration. The shortcoming of this method is that is requires to choose a discretization step at path construction and to lose the continuous information of the path.

Instead, we propose an alternative architecture where paths store the constraints they are subject to and apply the constraints at path evaluation. Let $P\in C^1([0,T],\CS)$ be a path without constraint defined on an interval $[0,T]$, and $\mathbf{proj}$ a projector onto a submanifold defined by numerical constraints (\textit{i.e.} an instance of\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}).

Then the corresponding constrained path $\tilde{P}$ is defined on the same interval by
$$
\forall t\in[0,T],\ \ \ \tilde{P}(t) = \mathbf{proj}(P(t))
$$
Paths are implemented by class \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1Path.html}{\texttt{Path}}. Several implementations of unconstrained paths are provided:\\
\href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1StraightPath.html}{\texttt{StraightPath}} for linear interpolation generalized to Lie groups, \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1ReedsSheppPath.html}{\texttt{ReedsSheppPath}}, \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1DubinsPath.html}{\texttt{DubinsPath}} for non\-holo\-nomic mobile robots.

\subsubsection{Continuity of projection along a path}

Projecting configurations at path evaluation has the advantage of not losing information. In return, the projection of a continuous path may be discontinuous. Before inserting a projected path in a roadmap for instance, it is therefore necessary to detect possible discontinuities. \cite{mirabel:hal-01360409} proposes two algorithms to check whether a projected path is continuous.

\section{Manipulation Problem}

The previous sections have presented how we model kinematic chains, configurations and velocities for a given robotic system, how configurations and paths can be projected onto a submanifold of the configuration space defined by numerical constraints.

In this section, we will use these notions to represent a robotic manipulation problem.

\begin{definition}\label{def:manipulation-problem}\textbf{Prehensile manipulation problem}

  \noindent A prehensile manipulation problem is defined by
  \begin{itemize}
  \item one or several robots,
  \item one or several objects,
  \item environment contact surfaces,
  \item object contact surfaces,
  \item an initial configuration,
  \item a final configuration.
  \end{itemize}
  \textit{Admissible configurations} of the system are configurations that satisfy the following property:
  \begin{itemize}
    \item each object is either grasped by a robot, or lies in a stable contact pose.
  \end{itemize}
  \textit{Admissible motions} of the system are motions that satisfy the following property:
  \begin{itemize}
  \item configurations along the motion are admissible, and
  \item the pose of objects in stable contact is constant,
  \item the relative pose of objects grasped by a gripper with respect to the gripper is constant.
  \end{itemize}
  The solution of a prehensile manipulation problem is an admissible motion that
  links the initial and goal configurations.\\
  -----------------------------------------------------------------------
\end{definition}
We will now provide precise definitions for gripper, grasp and stable contact pose.

\subsection{Grasp}\label{subsec:grasp}

\paragraph{Configuration space} The configuration space of a manipulation problem is the Cartesian product of the configuration spaces of the robots and of the objects.
$$
\CS = \CS_{r_1}\times\CS_{r_{nr}}\times SE(3)^{no}
$$
where $nr$ is the number of robots, $no$ is the number of objects, $\CS_{r_i},\ i\in\{1,...,nr\}$ is the configuration space of robot $r_i$.
\begin{definition}\textbf{Gripper.}\label{def:gripper}
  A \emph{gripper} $\gripper$ is defined as a frame attached to the link of a robot. $\gripper(\conf),\ \conf\in\CS$ denotes the pose of the frame when the system is in configuration $\conf$.
\end{definition}
\begin{definition}\textbf{Handle.}\label{def:handle}
  A \emph{handle} is composed of
  \begin{itemize}
  \item a frame $\handle$ attached to the root joint of an object,
  \item a list $flags=(x,y,z,rx,ry,rz)$ of 6 Boolean values.
  \end{itemize}
  $\handle(\conf),\ \conf\in\CS$ denotes the pose of the frame when the system is in configuration $\conf$.
\end{definition}
\begin{definition}\textbf{Grasp.}\label{def:grasp}
  A \emph{grasp} is a numerical constraint $h$ over $\CS$, defined by
  \begin{itemize}
  \item a gripper $\gripper$,
  \item a handle $\handle$.
  \end{itemize}
  Let $\bar{h}$ be the smooth mapping from $\CS$ to $\reals^6$ that maps to any configuration $\conf$ of the system, the expression
  \begin{equation}\label{eq:barh}
    \bar{h}(\conf) = \log_{\reals^3\times SO(3)}\left(\gripper^{-1}(\conf) \handle(\conf)\right).
  \end{equation}
  $h(\conf)$ is obtained by extracting from $\bar{h}$ the components the values of which are \true in the handle flag.
\end{definition}
\begin{definition}\textbf{Grasp complement.}
  Given a grasp constraint defined by gripper $\gripper$, handle $\handle$ and some flag vector, the grasp complement is a parameterized non-linear constraint defined by
  $$
  h_{comp}(\conf) = \mathbf{h}_0
  $$
  where $h_{comp}$ is composed of the components of $\bar{h}$ that are not in $h$ and $\mathbf{h}_0$ is a vector with the same size as $h_{comp}$ output.
\end{definition}

\paragraph{Geometric interpretation and examples}
The 3 first components of $\bar{h}(\conf)$ in equation~(\ref{eq:barh}) correspond to the position of the center of $\handle(\conf)$ in the frame of $\gripper(\conf)$. The 3 last components of $\bar{h}(\conf)$ is a vector representing the relative orientation of $\handle(\conf)$ with respect to $\gripper(\conf)$. The direction of the vector represents the axis of rotation, the norm of the vector represents the angle of rotation.
\begin{itemize}
\item If $flags=(\true,\true,\true,\true,\true,\true)$ the grasp is satisfied if $\gripper(\conf)$ and $\handle(\conf)$ coincide:
  $$ h =\bar{h},$$
  $h_{comp}$ is an empty constraint;
\item if $flags=(\true,\true,\true,\true,\true,\false)$ the grasp is satisfied if the centers and $z$ axes of $\gripper(\conf)$ and $\handle(\conf)$ coincide (free rotation around $z$). This is useful for cylindical objects:
  \begin{align*}
    h &= (\bar{h}_1,\bar{h}_2,\bar{h}_3,\bar{h}_4,\bar{h}_5),\\
    h_{comp} &= (\bar{h}_6);
  \end{align*}
\item if $flags=(\true,\true,\true,\false,\false,$ $\false)$ the grasp is satisfied if the centers of $\gripper(\conf)$ and $\handle(\conf)$ coincide (free rotation). This is useful for spherical objects.
  \begin{align*}
    h &= (\bar{h}_1,\bar{h}_2,\bar{h}_3),\\
    h_{comp} &= (\bar{h}_4,\bar{h}_5,\bar{h}_6).
  \end{align*}
\end{itemize}
If $\conf_0$ is a configuration satisfying the grasp constraint: $h(\conf_0)$=0, then the submanifold defined by
$$
\left\{\conf\in\CS,\ \ h(\conf)=0\ \ h_{comp}(\conf) =  h_{comp}(\conf_0)\right\}
$$
contains all the configurations that are reachable from $\conf_0$ while keeping the grasp.

\subsection{Stable Contact pose}\label{subsec:placement}

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figures/convex-shape-contact.pdf}
  \end{center}
  \caption{Distance defined by two convex polygons.}
  \label{fig:convex-shape-contact}
\end{figure}

When an object is not grasped, it should lie in a stable pose. There are two simple methods to do that:
\begin{enumerate}
\item defining virtual grippers in the environment and virtual handles on the object, implicitely defines a discrete set of poses,
\item defining a virtual gripper on an horizontal plane and a virtual handle on the object, and using a grasp with flags $(\false,\false,\true,\true,\true,\false)$ constrains the object to move on a infinite horizontal plane.
\end{enumerate}

We propose here a third method that enables users to define contact surfaces in a more flexible way. In this aim, we denote by:
\begin{itemize}
\item $(o_i)_{i\in I}$ a set of convex polygons attached to an object,
\item $(f_j)_{j\in J}$ a set of convex polygons attached to the environment or to a mobile part of a robot that can receive objects (mobile robot for instance),
\item respectively $C_{o_i}$, $\mathbf{n}_{o_i}$ the barycenter of $o_i$ and the normal to the plane containing $o_i$,
\item $C_{f_j}$, $\mathbf{n}_{f_j}$ the barycenter of $f_j$ and the normal to the plane containing $f_j$,
\item $P(C_{o_i},f_j)$, the orthogonal projection of $C_{o_i}$ onto the plane containing $f_j$.
\end{itemize}
Then we define the distance between polygons $o_i$ and $f_j$ as the distance of $C_{o_i}$ to the cylindrical volume of generatrix $\mathbf{n}_{f_j}$ and of directrix $f_j$:
\begin{align}\label{eq:contact-distance}
  d(f_j,o_i) = \sqrt{d_{\parallel}^2 + d_{\perp}^2}
\end{align}
where
\begin{align*}
  i,j & \mbox{ are the indices that minimize the above distance,}\\
d_{\parallel} &= \left\{\begin{array}{lll} d(f_j,P(C_{o_i},f_j))&\mbox{if}& P(C_{o_i},f_j)\mbox{ outside } f_j\\
0 & & \mbox{otherwise}
\end{array}\right. \\
d_{\perp} &= \mathbf{n}_{f_j}.\vec{C_{f_j}C_{o_i}}.
\end{align*}
Figure~\ref{fig:convex-shape-contact} illustrates this notation.

We denote by $o_i(\conf)$ and $f_j(\conf)$ the poses in configuration $\conf$ of frames with respective centers $C_{o_i}$ and $C_{f_j}$ and with $x$-axis normal to each polygon. In an analogous way as in Definition~\ref{def:grasp}, we define
\begin{equation}\label{eq:convex-shape-contact-hbar}
\bar{h}(\conf) = \log_{\reals^3\times SO(3)} \left(f_{j}(\conf)^{-1}o_i(\conf)\right)
\end{equation}
The contact constraint is defined by the following piecewise differentiable function:
\begin{equation}\label{eq:convex-shape-contact}
h(\conf) = \left\{\begin{array}{lll}(\bar{h}_1,0,0,\bar{h}_4,\bar{h}_5)&
\mbox{if} & P(C_{o_i},f_j)\mbox{ inside } f_j\\
(\bar{h}_1,\bar{h}_2,\bar{h}_3,\bar{h}_4,\bar{h}_5)&
\mbox{if} & P(C_{o_i},f_j)\mbox{ outside } f_j\end{array}\right.
\end{equation}
It is straightforward that the above function vanishes if and only if two convex polygons are in contact.

As for grasps, we need to define a parameterized complement constraint for the contact constraint in order to specify the submanifold of configurations that are reachable from one configuration while keeping the object in a constant stable pose. The naive way consists in defining
$$
h_{comp}(\conf) = \left\{\begin{array}{lll}(\bar{h}_2,\bar{h}_3,\bar{h}_6)&
\mbox{if} & P(C_{o_i},f_j)\mbox{ inside } f_j\\
(0,0,\bar{h}_6)&
\mbox{if} & P(C_{o_i},f_j)\mbox{ outside } f_j\end{array}\right.
$$
$\bar{h}_2$, $\bar{h}_3$, $\bar{h}_6$ respectively represent the translation in $x-y$ plane and the rotation around $z$-axis of frame $o_i$ with respect to frame $f_j$. Let $\conf_0$ be a configuration such that $h(\conf_0)=0$. The submanifold defined by
\begin{equation}\label{eq:stable-pose}
\left\{\conf\in\CS,\ \ h(\conf)=0\ \ h_{comp}(\conf) = h_{comp}(\conf_0) \right\}
\end{equation}
contains one object pose for each pair of polygons $(o_i,f_j)$, and there are
$|I| . |J|$ possible combinations. This contraint is therefore not suitable to
enforce object immobility along a path since the object may jump from one pose to another one.

To disambiguate the various combination of convex polygons that can be in contact, we define
\begin{equation}\label{eq:convex-shape-contact-complement}
h_{comp}(\conf) = \left\{\begin{array}{l}(\bar{h}_2 + 2jM,\bar{h}_3 + 2iM,\bar{h}_6)\\
\mbox{if } P(C_{o_i},f_j)\mbox{ inside } f_j\\
(2jM,2iM,\bar{h}_6)\\
\mbox{if } P(C_{o_i},f_j)\mbox{ outside } f_j\end{array}\right.
\end{equation}
where
\begin{itemize}
\item $i$ and $j$ are the indices that minimize distance~(\ref{eq:contact-distance}),
\item $M$ is a positive real number such that for any $\kappa\in J$, all vertices of $f_\kappa$ are included in the disk of center $C_{f_\kappa}$ and of radius $M$.
\end{itemize}
With this definition, the submanifold defined by~(\ref{eq:stable-pose}), (\ref{eq:convex-shape-contact}), (\ref{eq:convex-shape-contact-complement}) contains configurations where the object is in a unique stable pose. The polygon indices $i$ and $j$, as well as their relative position can indeed be recovered from~(\ref{eq:convex-shape-contact-complement}):
\begin{align*}
  i &= \left\lfloor\frac{\bar{h}_3}{2M} + \frac{1}{2}\right\rfloor\\
  j &= \left\lfloor\frac{\bar{h}_2}{2M} + \frac{1}{2}\right\rfloor\\
  \bar{h}_1 &= \bar{h}_4 = \bar{h}_5 = 0\\
  \bar{h}_2 &= h_{comp\ 1} - 2jM\\
  \bar{h}_3 &= h_{comp\ 2} - 2iM\\
  \bar{h}_6 &= h_{comp\ 3}
\end{align*}
and from (\ref{eq:convex-shape-contact-hbar}),
$$
f_{j}(\conf)^{-1}o_i(\conf) = \exp_{\reals^3\times SO(3)}\left(\bar{h}\right).
$$
Uniqueness comes from the fact that when two convex polygons are in contact,
necessarily, $|\bar{h}_2| \leq M$,$|\bar{h}_3| \leq M$.
% Remark about convergence of the solver and about explicit constraint.

\subsection{constraint graph}

According to Definition~\ref{def:manipulation-problem}, the set of admissible configurations is the union of sub-manifolds of the configuration space of the system. Each sub-manifold is defined by grasp and/or stable contact constraints. We call each sub-manifold a \textit{state} of the problem. A state is defined by a set of grasps and a set of stable contact poses.

A state can be defined by a subset of active grasps, any object not grasped being in a stable contact pose. Let $\ngrippers$ and $\nhandles$ repsectively denote the number of grippers and handles. A state $\state$ is denoted by a vector of size $\ngrippers$:
\begin{equation}\label{eq:state}
\state = \left(h_1,\cdots,h_{\ngrippers}\right)
\end{equation}
where $h_i\in\{\emptyset,1,\cdots,\nhandles\}$ denotes the id of the handle grasped by gripper $i$; $h_i=\emptyset$ means that gripper $i$ does not grasp any handle.

\paragraph{Number of states:} note that for $i\in\{1,\cdots,\nhandles\}$ the number of occurences of $i$ in $\state$ is at most 1: a handle cannot be grasped by several grippers. Note also that the number of occurences of $\emptyset$ is not limited: several grippers may hold nothing. Let $m$ be a non-negative integer not greater than $\ngrippers$ and than $\nhandles$ and let us count the number of states with $m$ handles grasped. The number of subset of $m$ handles among $\nhandles$ is equal to $\frac{\nhandles!}{(\nhandles-m)!m!}$. And the number of ways of dispatching them among the $\ngrippers$ grippers is equal to $\frac{\ngrippers!}{(\ngrippers-m)!}$. Thus, the total number of states is equal to
$$
\sum_{m=0}^{\min(\ngrippers,\nhandles)} \frac{\nhandles!}{(\nhandles-m)!m!} \frac{\ngrippers!}{(\ngrippers-m)!}
$$
\begin{definition}\label{def:neighboring-states}\textbf{Neighboring states}\\
  Two states $\state_1=(h_{1 1},\cdots,h_{\ngrippers 1})$ and $\state_2=(h_{1 2},\cdots,h_{\ngrippers 2})$ are neighboring if they differ by only one grasp and the grasp is empty in one of the states:
  \begin{align*}
    &\exists i\in\{1,\cdots,\ngrippers\},\,h_{i 1} \not= h_{i 2} \mbox{ and } (h_{i 1}=\emptyset \mbox{ or } h_{i 2}=\emptyset), \mbox{ and }\\
    &\forall j\in\{1,\cdots,\ngrippers\},\, j\not=i,h_{j 1} = h_{j 2}.
  \end{align*}
\end{definition}
\begin{definition}\label{def:constraint-graph}\textbf{Constraint graph}\\
  The constraint graph related to a manipulation problem as defined in Definition~\ref{def:manipulation-problem} is a graph
  \begin{itemize}
  \item the nodes of which are states defined by subsets of grasps~(\ref{eq:state}),
  \item an edge connects two states if and only if they are neighboring or they are equal. Edges are also called \textit{transitions}.
  \end{itemize}
  Nodes contain
  \begin{itemize}
  \item the grasp constraints that are active in the corresponding state,
  \item a placement constraint for each object that is not grasped by any handle.
  \end{itemize}
  Transitions contain
  \begin{itemize}
  \item the constraints of the node they connect with the least active grasps,
  \item the parameterized complement constraint of each of the latter.
  \end{itemize}
\end{definition}

\subsection{Example}

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figures/example.png}
    \def\svgwidth{\columnwidth}
    \graphicspath{{./figures/}}
    \input{figures/example-graph.pdf_tex}
  \end{center}
  \caption{Top: two robot UR3 with one gripper each (X=red, Y=gree, Z=blue) manipulating a cylinder with two handles. The environment contains one rectangular contact surface (in red). The cylinder has two rectangular contact surfaces (in green).
  Bottom: the corresponding constraint graph. Names of states follow Expression~(\ref{eq:state}): for instance, $(\emptyset,1)$ means that gripper of robot 2 grasps handle 1 of the cylinder. In this state, there is no placement constraint.}
  \label{fig:2ur3-cylinder}
\end{figure}
To illustrate the notions exposed in the previous sections, let us consider an example of two UR3 robots manipulating a cylinder illustrated in Figure~\ref{fig:2ur3-cylinder}. The robot is equipped with one gripper attached to the end-effector. The cylinder is equipped with two handles and with two square contact surfaces corresponding to the top and bottom faces of the cylinder.

Grasp constraints are denoted by
$$
grasp_{ij} \mbox{ where } i,j\in\{1,2\}.
$$
Grasp complement constraints are denoted by
$$
grasp_{ij}/comp \mbox{ where } i,j\in\{1,2\}.
$$

$i$ is the index of the gripper $j$ is the index of the handle. The flag of the handles is
$$(\true,\true,\true,\false,\true,\true).$$
Therefore grasp constraints are of dimension 5 and keep the rotation of the gripper around the cylinder axis free.
The object placement constraint is denoted by $place$ and the complement by $place/comp$.
\begin{table}
  \begin{center}
    \begin{tabular}{|p{.44\linewidth}|p{.44\linewidth}|}
      \hline
      state & active constraints\\
      \hline
      $(\emptyset,\emptyset)$ & $place$ \\
      $(j,\emptyset),\,j\in\{1,2\}$ & $grasp_{1j}$\\
      $(\emptyset,j),\,j\in\{1,2\}$ & $grasp_{2j}$\\
      $(i,j)\,i,j\in\{1,2\}$ & $grasp_{1i},grasp_{2j}$\\
      \hline
    \end{tabular}
  \end{center}
  \caption{State constraints.}
  \label{tab:state-constraints}
\end{table}
Table~\ref{tab:state-constraints} indicates which constraints are active for each state.

\begin{table}
  \begin{center}
    \begin{tabular}{|p{.23\linewidth}|p{.18\linewidth}|p{.43\linewidth}|}
      \hline
      transition & belongs to & additional constraints\\
      \hline
      $(\emptyset,\emptyset)\rightarrow (i,\emptyset)$ & $(\emptyset,\emptyset)$ & $place/comp$\\
      $(\emptyset,\emptyset)\rightarrow (\emptyset,i)$ & $(\emptyset,\emptyset)$ & $place/comp$\\
      $(i,\emptyset)\rightarrow(i,j)$ & $(i,\emptyset)$ & $grasp_{1i}/comp$\\
      $(\emptyset,j)\rightarrow(i,j)$ & $(\emptyset,j)$ & $grasp_{2j}/comp$\\
      \hline
    \end{tabular}
  \end{center}
  \caption{Transition constraints: $i,j$ are either 1 or 2. column "belongs to" means that paths along the transition belong to the state.}
  \label{tab:transition-constraints}
\end{table}

\subsection{Automatic construction}

\begin{algorithm}
  \algblock[globalVariable]{GlobalVariable}{EndGlobalVariable}
  \begin{algorithmic}[1]
    \GlobalVariable
    \State{$\nobjects$}\Comment{number of objects}
    \State{$\ngrippers$}\Comment{number of grippers}
    \State{$\nhandles$}\Comment{number of handles}
    \EndGlobalVariable
    \Function{buildConstraintGraph}{}
    \State{$\mathcal{G}\gets\left[0,\cdots,\ngrippers-1\right]$}
    \Comment{gripper indices}
    \State{$\mathcal{H}\gets\left[0,\cdots,\nhandles-1\right]$}
    \Comment{handle indices}
    \State {$Gr\gets\left[\emptyset,\cdots,\emptyset\right]$}\Comment{list of size $\ngrippers$}
    \State \Call{recurse}{$\mathcal{G}$,$\mathcal{H}$,$Gr$}
    \EndFunction
    \Function{recurse}{$\mathcal{G}$,$\mathcal{H}$,$Gr$}
    \State\Call{createState}{$Gr$}
    \If{$\mathcal{G}=\emptyset$ or $\mathcal{H}=\emptyset$}
    \State{\Return}
    \EndIf
    \For{$g$ in $\mathcal{G}$}
    \State{$\mathcal{G}^{\prime}\gets \mathcal{G}\setminus\{g\}$}
    \For{$h$ in $\mathcal{H}$}
    \State{$\mathcal{H}^{\prime}\gets\mathcal{H}\setminus\{h\}$}
    \State{$Gr^{\prime}\gets Gr$}
    \State{$Gr^{\prime}[g]\gets h$}
    \State{\Call{createState}{$Gr^{\prime}$}}
    \State{\Call{createTransition}{$Gr$, $Gr^{\prime}$}}
    \State{\Call{recurse}{$\mathcal{G}^{\prime}$,$\mathcal{H}^{\prime}$,$Gr^{\prime}$}}
    \EndFor
    \EndFor
    \EndFunction
    \Function{createState}{$Gr$}
    \If{\Call{existState}{$Gr$}}\State{\Return}\EndIf
    \State{$\state\gets\mbox{ new state}$}
    \State{$\state.Pl\gets\left[\true,...,\true\right]$}\Comment{list of size $\nobjects$}
    \For{$g\mbox{ in } \left[0,\cdots,\ngrippers-1\right]$}
    \State{$h\gets Gr[g]$}
    \State{$\state.Pl[\Call{objectIndex}{h}]\gets\false$}
    \State{$\state$.add(\Call{graspConstraint}{g,h})}
    \EndFor
    \For{$o\mbox{ in }\left[0,\cdots,\nobjects\right]$}
    \If{$\state.Pl[o]$}
    \State{$\state$.add(\Call{placeConstraint}{$o$}})
    \EndIf
    \EndFor
    \EndFunction
    \Function{createTransition}{$Gr_1,Gr_2$}
    \State{$\transition\gets\mbox{ new transition}(Gr_1,Gr_2)$}
    \State{$\state_1\gets$\Call{state}{$Gr_1$}}\Comment{Recover state for this set of grasps}
    \For{$g\mbox{ in } \left[0,\cdots,\ngrippers-1\right]$}
    \State{$h\gets Gr_1[g]$}
    \State{$\transition$.add(\Call{graspConstraint}{g,h})}
    \State{$\transition$.add(\Call{graspConstraintComp}{g,h})}
    \EndFor
    \For{$o\mbox{ in }\left[0,\cdots,\nobjects\right]$}
    \If{$\state_1.Pl[o]$}
    \State{$\transition$.add(\Call{placeConstraint}{$o$}})
    \State{$\transition$.add(\Call{placeConstraintComp}{$o$}})
    \EndIf
    \EndFor
    \EndFunction
    \Function{existState}{$Gr$}
    \Comment{Return $\true$ if a state has already been created for the set of
    grasps given as input.}
    \EndFunction
    \Function{state}{$Gr$}
    \Comment{Return the state created with the set of grasps given as input.}
    \EndFunction
    \Function{objectIndex}{$h$}
    \Comment{Return the index of the object handle $h$ belongs to.}
    \EndFunction
  \end{algorithmic}
  \caption{Recursive Construction of the constraint graph. The
    construction starts by the state with no grasp. Call to recurse
    function loops over the available grippers and handles and create
    states with one more grasp, and a transition to these new
    states. In each state, a placement constraint is added for each
    object of which no handle is grasped. Variables $\mathcal{G}$ and
    $\mathcal{H}$ contain the indices of the free grippers and
    handles. Variable $Gr$ stores the current set of grasps following
    Expression~(\ref{eq:state}). Functions {\scriptsize GRASP}{\small C}{\scriptsize ONSTRAINT}, {\scriptsize GRASP}{\small C}{\scriptsize ONSTRAINT}{\small C}{\scriptsize OMP},{\scriptsize PLACE}{\small C}{\scriptsize ONSTRAINT}, {\scriptsize PLACE}{\small C}{\scriptsize ONSTRAINT}{\small C}{\scriptsize OMP} build grasp, placement constraints and their complements as defined in sections~\ref{subsec:grasp} and ~\ref{subsec:placement}.}
  \label{algo:constraintGraphFactory}
\end{algorithm}

Given a set of grippers, handles and objects, the constraint graph can be constructed automatically. \href{https://github.com/humanoid-path-planner/hpp-manipulation-corba/blob/5af1b3bad68e8c339d5f42eb72173d7356504532/src/hpp/corbaserver/manipulation/constraint_graph_factory.py#L187}{Here} is an implementation in python. Algorithm~\ref{algo:constraintGraphFactory} describes this implementation.

\section{Manipulation planning}

In this section, we show how the constraint graph defined in the previous section is used to plan collision-free manipulation paths. Although we are working on an extension of Algorithm RMR*~\cite{schmitt17icra} to several grippers, objects and handles, the only manipulation planning algorithm available up to now in HPP is an extension of RRT algorithm described in the next section.

\subsection{M-RRT}

Manipulation Randomly exploring Random Tree is an extension of RRT algorithm~\cite{LavKuf01b} that grows trees in the free configuration space, exploring the different states of the manipulation problem. Algorithm~\ref{algo:M-RRT} describes the algorithm implemented \href{https://github.com/humanoid-path-planner/hpp-manipulation/blob/16369aa291ab1b17ef6176ae8b8b2512b5e6fff7/src/manipulation-planner.cc#L159}{here}.
\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{oneStep}{}
    $q_{rand}\gets$\Call{randomConfig}{}
    \EndFunction
  \end{algorithmic}
  \caption{}
  \label{algo:M-RRT}
\end{algorithm}

\subsection{Waypoint edges}

\subsection{Crossed foliation issue}

\subsection{Prototyping dedicated algorithms}

### 2020-07-10 script/example.py

@@ -1,0 +1,97 @@
+from math import pi, sqrt
+from hpp.corbaserver.manipulation.ur5 import Robot
+from hpp.corbaserver.manipulation import Client, ConstraintGraph,\
+    ConstraintGraphFactory, ProblemSolver
+from hpp.gepetto.manipulation import ViewerFactory
+from hpp_idl.hpp import Equality, EqualToZero
+
+class Ground (object):
+  rootJointType = 'anchor'
+  urdfFilename = "package://example/urdf/ground.urdf"
+  srdfFilename = "package://example/srdf/ground.srdf"
+
+class Box (object):
+  rootJointType = 'freeflyer'
+  urdfFilename = "package://example/urdf/box.urdf"
+  srdfFilename = "package://example/srdf/box.srdf"
+
+class Cylinder (object):
+  rootJointType = 'freeflyer'
+  urdfFilename = "package://example/urdf/cylinder.urdf"
+  srdfFilename = "package://example/srdf/cylinder.srdf"
+
+Robot.urdfFilename = "package://example-robot-data/robots/ur_description/urdf/ur3_gripper.urdf"
+Robot.srdfFilename = "package://example-robot-data/robots/ur_description/srdf/ur3_gripper.srdf"
+
+from hpp.corbaserver import loadServerPlugin
+loadServerPlugin ("corbaserver", "manipulation-corba.so")
+Client ().problem.resetProblem ()
+
+robot = Robot('2ur3-box', 'r0')
+robot.setJointPosition ('r0/root_joint', [-.25, 0, 0, 0, 0, 0, 1])
+ps = ProblemSolver (robot)
+vf = ViewerFactory (ps)
+
+vf.loadRobotModel (Robot, "r1")
+robot.setJointPosition ('r1/root_joint', [.25, 0, 0, 0, 0, 1, 0])
+
+# Change bounds of robots to increase workspace and avoid some collisions
+robot.setJointBounds ('r0/shoulder_pan_joint', [-pi, 4])
+robot.setJointBounds ('r1/shoulder_pan_joint', [-pi, 4])
+robot.setJointBounds ('r0/shoulder_lift_joint', [-pi, 0])
+robot.setJointBounds ('r1/shoulder_lift_joint', [-pi, 0])
+robot.setJointBounds ('r0/elbow_joint', [-2.6, 2.6])
[…]

### 2020-07-10 script/gepetto-gui.sh

@@ -1,0 +1,2 @@
+#!/bin/sh
+ROS_PACKAGE_PATH=${ROS_PACKAGE_PATH}:${PWD}/share gepetto-gui

### 2020-07-10 script/hpp-corbaserver.sh

@@ -1,0 +1,2 @@
+#!/bin/sh
+ROS_PACKAGE_PATH=${ROS_PACKAGE_PATH}:${PWD}/share hppcorbaserver

### 2020-07-10 script/tiago.py

@@ -1,0 +1,175 @@
+#!/usr/bin/env python
+# Start hppcorbaserver before running this script
+# Note that an instance of omniNames should be running in background
+#
+# Import libraries and load robots. {{{1
+
+# Import. {{{2
+from argparse import ArgumentParser
+from math import sqrt
+from hpp.corbaserver.manipulation.tiago import Robot
+from hpp.corbaserver.manipulation import ProblemSolver, ConstraintGraph, \
+  ConstraintGraphFactory, Constraints, Rule, Client
+from hpp.gepetto.manipulation import ViewerFactory
+from math import pi
+from hpp.corbaserver import loadServerPlugin
+loadServerPlugin ("corbaserver", "manipulation-corba.so")
+Client ().problem.resetProblem ()
+# 2}}}
+
+# parse arguments {{{2
+parser = ArgumentParser()
+parser.add_argument('-N', default=20, type=int)
+parser.add_argument('--display', action='store_true')
+parser.add_argument('--run', action='store_true')
+args = parser.parse_args()
+# 2}}}
+
+# Load PR2 and a box to be manipulated. {{{2
+class Box (object):
+  rootJointType = 'freeflyer'
+  packageName = 'hpp_tutorial'
+  meshPackageName = 'hpp_tutorial'
+  urdfName = 'box'
+  urdfSuffix = ""
+  srdfSuffix = ""
+
+class Environment (object):
+  packageName = 'iai_maps'
+  meshPackageName = 'iai_maps'
+  urdfName = 'kitchen_area'
+  urdfSuffix = ""
+  srdfSuffix = ""
+
+robot = Robot ('pr2-box', 'pr2')
+ps = ProblemSolver (robot)
+## ViewerFactory is a class that generates Viewer on the go. It means you can
+## restart the server and regenerate a new windows.
+## To generate a window:
+## vf.createRealClient ()
+vf = ViewerFactory (ps)
+
+vf.loadObjectModel (Box, 'box')
+vf.loadEnvironmentModel (Environment, "kitchen_area")
+
+robot.setJointBounds ("pr2/root_joint" , [-5,-2,-5.2,-2.7,-2,2,-2,2] )
+robot.setJointBounds ("box/root_joint", [-5.1,-2,-5.2,-2.7,0,1.5,-2,2,-2,2,-2,2,-2,2])
+# 2}}}
+
+# 2}}}
+
[…]

### 2020-09-30 hpp.tex
Write related work.

@@
 \section{Introduction}
-This \paper describes a software platform taylored for manipulation planning.
+This \paper describes a software platform taylored for manipulation planning in
+robotics.
 The main contributions are:
@@
 The \paper is organized as follows.
-\partie~\ref{partie:preliminary} introduces some preliminary notions like kinematic chains and Lie groups that are used to model the configuration space of each joint.
-
-Each \partie is implemented by one or several software packages. Instead of providing formulas for some values need to be computed, we provide a link to the C++ implementation.
+\Partie~\ref{partie:preliminary} introduces some preliminary notions like kinematic chains and Lie groups that are used to model the configuration space of each joint.
+\Partie~\ref{sec:constraints} introduces non linear constraints and solvers that are at the core of manipulation problem definition.
+\Partie~\ref{sec:manipulation-problem} defines the problem of prehensile manipulation in the general setting.
+\Partie~\ref{sec:manipulation-planning} provides a general algorithm that solves manipulation planning problems.
+Finally, \partie~\ref{sec:humanoid-path-planner} is devoted to a software platform implementing the notions introduced in the previous sections.
+
+Each \partie is implemented by one or several software packages. Instead of providing formulas for some values that need to be computed, we provide a link to the C++ implementation.
+
+\section{Related work}\label{sec:related-work}
+
+Motion planning has given rise to a lot of research work for the past decades.
+The problem consists in finding a collision-free path for a given system in
+an environment populated with obstacles. The field covers a large variety of
+different applications ranging from navigation for autonomous vehicles in
+partially known environments~\cite{DarpaUrbanChallenge} to path planning for
+deformable objects~\cite{LamKav2001},\\
+\cite{RouFerTai2020} and many other applications like coverage path planning~\cite{coverage2001},\\ \cite{coverage2013}, or pursuit evasion planning~\cite{pursuit-evasion1999}.
+
+Planning motions for high dimensional robots like humanoid robots or multi-arm
+system has been proven to be of a very high complexity~\cite{SchSha1983},\\ \cite{Canny1983}. Starting in the 1990's random sampling methods have been proposed
+to solve the problem, trading the completeness property against efficiency in
+solving problem in high dimensional configuration spaces\\ \cite{KSLO1996}, \cite{HsuLatMot99}, \cite{KufLav00}. Since then, asymptotically optimal random sampling algorithms have
+been proposed~\cite{KarFra2011}.
+
+\subsection{Path planning with non-linear constraints}
+
+Some systems are subject to non-linear constraints. These non-linear constraints
+define a sub-manifold of the configuration the robot must stay on. For instance,
+legged robots that must keep contact with the ground and enforce quasi-static
+equilibrium, or multi-arm systems grasping a same object are subject to this
+type of constraints. As the volume of the constrained manifold is usually equal
+to zero, sampling random configurations satisfying the constraints is an event
+of zero probability. To sample configurations on the constrained manifold,
+\cite{DalNakLamLau2009} and \cite{berenson2011} projects random configurations using a generalization of
+Newton-Raphson algorithm. Another way to sample configurations on a constrained
+manifold consist in expressing some configuration variables with respect to
+others~\cite{Cortes2002},\\ \cite{MirLam2018} when this is possible.\\ \cite{atlasRRT} propose another method based on non-linear projection. They cover the
+constrained manifold by growing an atlas composed of local charts. This
+approximation provides a probability distribution that is closer to the uniform
+distribution over the manifold than the projection of a uniform distribution
+over the configuration space.
+
+\subsection{Manipulation planning}
+
+Manipulation planning is a particular instance of path planning were
+some objects are moved by some robots. The configuration space of the
+whole system is constrained by non-linear constrained due to the fact
+that objects cannot move by themselves and should stay in a stable
+pose when not grasped by a robot.  The accessible configuration space
+is thus a union of sub-manifold as defined in the previous
+section. Each of these manifold is moreover a foliation where each
+leaf corresponds to a stable position of an object or to a grasp of an
+object by a gripper. The geometrical structure of the problem is now
+well understood~\cite{AlaSimLau89}. Although several instances of
+manipulation problem exist like manipulation by
+pushing~\cite{BenRiv1998}, or by throwing~\cite{WooZacLyn2017}, as well as
+multicontact planning \cite{Bretl2006}, \cite{LenVaiYosKhe2013},
+\cite{TonDelPetParManMan2018}, in this paper, we are only concerned with
+prehensile manipulation planning.
+
+Papers about manipulation planning are commonly divided into several categories.
+
+Navigation Among Movable Obstacles (NAMO) \cite{wilfong1988motion}, \cite{stilman2008planning} consists in finding
+a path for a robot that needs to displace objects in order to move to a goal
+configuration. The final position of the objects does not matter in this case.
+
+Rearrangement planning \cite{ota2004rearrangement}, \cite{LIS210}, consists in finding a sequence of manipulation paths that
+move some objects from an initial pose to a final pose. The final configuration
+of the robot is not specified. A simplifying asumption is the existence of a
+monotone solution, that is a solution where each object is grasped once only and
+is moved from its initial pose to its final pose \cite{stilman2008planning}, \cite{stilman2007manipulation}, \cite{srivastavaetal2014icra}, \cite{nieuwenhuisen2008effective}, \cite{ota2004rearrangement}.
+They mostly use two-level methods
+composed of a symbolic or task planner and of a motion planner~\cite{lavalle2006planning}.
+
+\subsection{Open-source software platforms}
 
@@
 \paragraph{Vector representations}
-Each Liegroup element is represented by a vector. Rotations are represented by
+Each Lie group element is represented by a vector. Rotations are represented by
 unit quaternions.
@@
 
-\section{Non-linear constraints and solvers}
+\section{Non-linear constraints and solvers}\label{sec:constraints}
 
@@
 To optimize constraint resolution, we perform variable substitution when possible in order to reduce both the number of variables and the dimension of the resulting implicit constraint.
-We describe here the method that has been first published in~\cite{mirabel:hal-01804774}.
+We describe here the method that has been first published in~\cite{MirLam2018}.
 Unlike in the former paper, the description we give in Algorithm~\ref{alg:ExplicitConstraintSet} is closer to the real implementation.
@@
 
-\section{Manipulation Problem}
+\section{Manipulation Problem}\label{sec:manipulation-problem}
 
@@
 \end{equation}
-It is straightforward that the above function vanishes if and only if two convex polygons are in contact.
+It is straightforward that the above function vanishes if and only if two convex polygons $o_i$ and $f_j$ are in contact and $C_{o_i}$ is inside $f_j$.
 
@@
 contains one object pose for each pair of polygons $(o_i,f_j)$, and there are
-$|I| . |J|$ possible combinations. This contraint is therefore not suitable to
+$|I| . |J|$ possible combinations. This constraint is therefore not suitable to
 enforce object immobility along a path since the object may jump from one pose to another one.
@@
 
-\subsection{constraint graph}
+\subsection{Merging constraint and complement into an explicit constraint}
+\label{subsec:explicit}
+
+Note that when a grasp constraint and its complement are combined, they
+constitute an explicit constraint since the pose of the object grasped
+uniquely depends on the configuration of the robot that grasps.
+
+Similarly, when a placement constraint and its complement are combined, they
+constitute an explicit constraint since the pose of the object placed uniquely
+depends on the pose of the contact surface on which the object is placed. This
+latter pose
+\begin{itemize}
+\item either depends on the configuration of the robot the contact surface
+  belongs to,
+\item or is constant if the contact surface belongs to the environment.
+\end{itemize}
+In any case, the explicit expression of the object pose depends on the right
+hand side of the complement constraint that is constant along transition paths.
+
+During the construction of the constraint graph (described in Section~\ref{subsec:constraint graph}), grasp and placement
+constraints, their complements and the associated explicit constraints
+are created together and registered using method\\
+\texttt{registerConstraint} of Class
+\href{https://gepettoweb.laas.fr/hpp/hpp-manipulation/doxygen-html/classhpp_1_1manipulation_1_1graph_1_1Graph.html}{\texttt{ConstraintGraph}}.
+
+\subsection{Constraint Graph}\label{subsec:constraint graph}
 
@@
 We call each sub-manifold a \textit{state} of the problem.
-A state is defined by a set of grasps and a set of stable contact poses.
 
 A state can be defined by a subset of active grasps, any object not grasped being in a stable contact pose.
-Let $\ngrippers$ and $\nhandles$ repsectively denote the number of grippers and handles.
+Let $\ngrippers$, $\nhandles$, and $\nobjects$ respectively denote the number of grippers, handles, and objects.
+
+We denote by
+\begin{itemize}
+\item $grasp_{ij}\ i\in\{1,\cdots,\ngrippers\}\ j\in\{1,\cdots,\nhandles\}$ the grasp constraint of handle $j$ by gripper $i$,
+\item $grasp_{ij}/comp\ i\in\{1,\cdots,\ngrippers\}\ j\in\{1,\cdots,\nhandles\}$ the complement constraint of the latter,
+\item $place_{i}\ i\in\{1,\cdots\nobjects\}$ the placement constraint of object $i$,
+\item $place_{i}/comp\ i\in\{1,\cdots\nobjects\}$ the complement constraint of the latter.
+\end{itemize}
+
 A state $\state$ is denoted by a vector of size $\ngrippers$:
@@
 Note also that the number of occurences of $\emptyset$ is not limited: several grippers may hold nothing.
-Let $m$ be a non-negative integer not greater than $\ngrippers$ and than $\nhandles$ and let us count the number of states with $m$ handles grasped.
+Let $m$ be a non-negative integer not greater than $\ngrippers$ nor $\nhandles$ and let us count the number of states with $m$ handles grasped.
 The number of subset of $m$ handles among $\nhandles$ is equal to $\frac{\nhandles!}{(\nhandles-m)!m!}$.
@@
 
-\subsection{Example}
+\subsection{Example}\label{subsec:ur3-cylinder}
 
@@
 The cylinder is equipped with two handles and with two square contact surfaces corresponding to the top and bottom faces of the cylinder.
-
-Grasp constraints are denoted by
-$$
-grasp_{ij} \mbox{ where } i,j\in\{1,2\}.
-$$
-Grasp complement constraints are denoted by
-$$
-grasp_{ij}/comp \mbox{ where } i,j\in\{1,2\}.
-$$
-
-$i$ is the index of the gripper $j$ is the index of the handle. The flag of the handles is
+$\ngrippers=2,\nhandles=2,\nobjects=1$.
+The flag of the handles are
 $$(\true,\true,\true,\false,\true,\true).$$
 Therefore grasp constraints are of dimension 5 and keep the rotation of the gripper around the cylinder axis free.
-The object placement constraint is denoted by $place$ and the complement by $place/comp$.
 \begin{table}
@@
       \hline
-      $(\emptyset,\emptyset)$ & $place$ \\
+      $(\emptyset,\emptyset)$ & $place_1$ \\
       $(j,\emptyset),\,j\in\{1,2\}$ & $grasp_{1j}$\\
@@
 \end{table}
-Table~\ref{tab:state-constraints} indicates which constraints are active for each state.
+Table~\ref{tab:state-constraints} indicates which constraints are active for each state. Table~\ref{tab:transition-constraints} indicates which constraints are active for each transition.
 
@@
       \hline
-      $(\emptyset,\emptyset)\rightarrow (i,\emptyset)$ & $(\emptyset,\emptyset)$ & $place/comp$\\
-      $(\emptyset,\emptyset)\rightarrow (\emptyset,i)$ & $(\emptyset,\emptyset)$ & $place/comp$\\
+      $(\emptyset,\emptyset)\rightarrow (i,\emptyset)$ & $(\emptyset,\emptyset)$ & $place_1/comp$\\
+      $(\emptyset,\emptyset)\rightarrow (\emptyset,i)$ & $(\emptyset,\emptyset)$ & $place_1/comp$\\
       $(i,\emptyset)\rightarrow(i,j)$ & $(i,\emptyset)$ & $grasp_{1i}/comp$\\
@@
   \end{center}
-  \caption{Transition constraints: $i,j$ are either 1 or 2. column "belongs to" means that paths along the transition belong to the state.}
+  \caption{Transition constraints: $i,j$ are either 1 or 2. column
+    "belongs to" means that paths along the transition belong to the
+    state, \textit{i.e.} the transition contains the state
+    constraints.}
   \label{tab:transition-constraints}
@@
 \begin{algorithm}
-  \algblock[globalVariable]{GlobalVariable}{EndGlobalVariable}
+  \algblockdefx[globalVariable]{GlobalVariable}{EndGlobalVariable}{\textbf{global variables}}{}
   \begin{algorithmic}[1]
@@
     \State{$\mathcal{G}\gets\left[0,\cdots,\ngrippers-1\right]$}
-    \Comment{gripper indices}
+    \Comment{list of gripper indices}
     \State{$\mathcal{H}\gets\left[0,\cdots,\nhandles-1\right]$}
-    \Comment{handle indices}
+    \Comment{list of handle indices}
     \State {$Gr\gets\left[\emptyset,\cdots,\emptyset\right]$}\Comment{list of size $\ngrippers$}
@@
     \State{$Gr^{\prime}[g]\gets h$}
+    \State{$isNewState\gets\mathbf{not}\ $\Call{existState}{$Gr^{\prime}$}}
     \State{\Call{createState}{$Gr^{\prime}$}}
     \State{\Call{createTransition}{$Gr$, $Gr^{\prime}$}}
-    \State{\Call{recurse}{$\mathcal{G}^{\prime}$,$\mathcal{H}^{\prime}$,$Gr^{\prime}$}}
+    \If{$isNewState$}{\Call{recurse}{$\mathcal{G}^{\prime}$,$\mathcal{H}^{\prime}$,$Gr^{\prime}$}}\label{cond:recurse}
+    \EndIf
     \EndFor
@@
     \State{$\state\gets\mbox{ new state}$}
-    \State{$\state.Pl\gets\left[\true,...,\true\right]$}\Comment{list of size $\nobjects$}
+    \State{$\state.Pl\gets\left[\true,...,\true\right]$}\Comment{list of size $\nobjects$}\label{beg:comp-placement-constraints}
     \For{$g\mbox{ in } \left[0,\cdots,\ngrippers-1\right]$}
@@
     \State{$\state$.add(\Call{graspConstraint}{g,h})}
-    \EndFor
+    \EndFor\label{end:comp-placement-constraints}
     \For{$o\mbox{ in }\left[0,\cdots,\nobjects\right]$}
@@
     \For{$o\mbox{ in }\left[0,\cdots,\nobjects\right]$}
+    \label{beg:set-placement-constraints}
     \If{$\state_1.Pl[o]$}
@@
     \EndIf
-    \EndFor
-    \EndFunction
-    \Function{existState}{$Gr$}
-    \Comment{Return $\true$ if a state has already been created for the set of
-    grasps given as input.}
-    \EndFunction
-    \Function{state}{$Gr$}
-    \Comment{Return the state created with the set of grasps given as input.}
-    \EndFunction
-    \Function{objectIndex}{$h$}
-    \Comment{Return the index of the object handle $h$ belongs to.}
+    \EndFor\label{end:set-placement-constraints}
     \EndFunction
@@
   \caption{Recursive Construction of the constraint graph. The
-    construction starts by the state with no grasp. Call to recurse
-    function loops over the available grippers and handles and create
-    states with one more grasp, and a transition to these new
-    states. In each state, a placement constraint is added for each
-    object of which no handle is grasped. Variables $\mathcal{G}$ and
-    $\mathcal{H}$ contain the indices of the free grippers and
-    handles. Variable $Gr$ stores the current set of grasps following
-    Expression~(\ref{eq:state}).
-Functions {\scriptsize GRASP}{\small C}{\scriptsize ONSTRAINT}, {\scriptsize GRASP}{\small C}{\scriptsize ONSTRAINT}{\small C}{\scriptsize OMP},{\scriptsize PLACE}{\small C}{\scriptsize ONSTRAINT}, {\scriptsize PLACE}{\small C}{\scriptsize ONSTRAINT}{\small C}{\scriptsize OMP} build grasp, placement constraints and their complements as defined in sections~\ref{subsec:grasp} and ~\ref{subsec:placement}.}
+    construction starts by the state with no grasp. Call to
+    {\scriptsize RECURSE} function loops over the available grippers
+    and handles and creates states with one more grasp, and a
+    transition to these new states. In each state, a placement
+    constraint is added for each object of which no handle is
+    grasped. Variables $\mathcal{G}$ and $\mathcal{H}$ contain the
+    indices of the free grippers and handles. Variable $Gr$ stores the
+    current set of grasps following Expression~(\ref{eq:state}).
+    Lines~\ref{beg:comp-placement-constraints} to
+    \ref{end:comp-placement-constraints}
+    compute which objects are not grasped.
+    Lines~\ref{beg:set-placement-constraints} to
+    \ref{end:set-placement-constraints} insert placement constraints in the
+    state for those objects. Line~\ref{cond:recurse} recurses only if the
+    lastest node reached is new.}
   \label{algo:constraintGraphFactory}
@@
 Algorithm~\ref{algo:constraintGraphFactory} describes this implementation.
-
-\section{Manipulation planning}
+Functions
+\begin{itemize}
+\item {\scriptsize GRASP}{\small C}{\scriptsize ONSTRAINT},
+\item {\scriptsize GRASP}{\small C}{\scriptsize ONSTRAINT}{\small C}{\scriptsize OMP}, build grasp constraint and complement constraint as defined in Section~\ref{subsec:grasp},
+\item {\scriptsize PLACE}{\small C}{\scriptsize ONSTRAINT},
+\item {\scriptsize PLACE}{\small C}{\scriptsize ONSTRAINT}{\small C}{\scriptsize OMP} build placement constraints and complement as defined in Section~\ref{subsec:placement},
+\item {\scriptsize EXIST}{\small S}{\scriptsize TATE}($Gr$) returns $\true$ if a state has already been created for the set of grasps given as input,
+\item{\scriptsize STATE}($Gr$) returns the state created with the set of grasps given as input,
+\item {\scriptsize OBJECT}{\small I}{\scriptsize NDEX}($h$) returns the index of the object handle $h$ belongs to.
+\end{itemize}
+
+\section{Manipulation planning}\label{sec:manipulation-planning}
 
@@
 
-\subsection{M-RRT}
+\subsection{Manipulation-RRT}
 
 Manipulation Randomly exploring Random Tree is an extension of RRT algorithm~\cite{LavKuf01b} that grows trees in the free configuration space, exploring the different states of the manipulation problem.
-Algorithm~\ref{algo:M-RRT} describes the algorithm implemented \href{https://github.com/humanoid-path-planner/hpp-manipulation/blob/16369aa291ab1b17ef6176ae8b8b2512b5e6fff7/src/manipulation-planner.cc#L159}{here}.
+Algorithm~\ref{algo:M-RRT} describes the algorithm implemented in C++ \href{https://github.com/humanoid-path-planner/hpp-manipulation/blob/16369aa291ab1b17ef6176ae8b8b2512b5e6fff7/src/manipulation-planner.cc#L159}{here}.
+
+After initializing the roadmap with the initial and goal configurations, the algorithm iteratively calls method {\scriptsize ONE}{\small S}{\scriptsize TEP} until a solution path is found or the maximum number of iterations is reached.
+This latter method shoots a random configuration (line 6) and for each connected component of the roadmap and each state of the constraint graph, extends the nearest node in the direction of the random configuration (lines 7--10).
+For each successful extension, the end of the extension path is stored for later connection (line 11).
+After the extension step, the algorithm tries to connect new nodes to other connected components using two strategies:
+\begin{enumerate}
+\item function \TRYCONNECTNEWNODES calls method \CONNECT between all pairs of new nodes,
+\item function \TRYCONNECTTOROADMAP tries to connect each new node to the nearest nodes in other connected components of the roadmap also using function \CONNECT.
+\end{enumerate}
+Function \CONNECT attempts to connect two configurations in two states.
+First, it checks whether there exists a transition between the states.
+If so, it checks that the right hand side of the transition parameterized constraints is the same for both configurations (up to the error threshold).
+Then it returns the linear interpolation between the configurations, projected onto the submanifold defined by the transition constraints.
+If the path is in collision, only the collision-free part at the beginning of the path is returned.
+
+Function \EXTEND attempts to generate a path from a configuration in a state to another state following a random transition.
+Similarly as for function \CONNECT, the path is projected onto the sub-manifold defined by the transition constraints.
+The end configuration is obtained by applying to the random configuration the constraints of the transition and of the goal state.
 \begin{algorithm}
   \begin{algorithmic}[1]
-    \Function{oneStep}{}
-    $q_{rand}\gets$\Call{randomConfig}{}
+    \Function{initializeRoadmap}{$\conf_{init}$,$\conf_{goal}$}
+    \State$\roadmap\gets\mbox{ new roadmap}$
+    \State{$\roadmap$.\Call{addNode}{$\conf_{init}$};
+      $\roadmap$.\Call{addNode}{$\conf_{goal}$}}
+    \EndFunction
+    \Function{oneStep}{$\roadmap$}
+    \State{$newNodes\gets$ empty list}
+    \State{$\conf_{rand}\gets$\Call{shootRandomConfig}{ }}
+    \For{$cc$ in connected components of $\roadmap$}
+    \For{$s$ in constraint graph states}
+    \State{$\conf_{near}\gets$\Call{nearestNode}{$cc$,$s$,$\conf_{rand}$}}
+    \State{$p\gets$\Call{extend}{$s$, $\conf_{near}$, $\conf_{rand}$}}
+    \If{$p$}$\,newNodes\gets\,newNodes\cup\{\mbox{end of}\,p\}$\EndIf
+    \EndFor
+    \EndFor
+    \State{$nc\gets$\Call{tryConnectNewNodes}{$\roadmap$, $newNodes$}}
+    \If{$nc=0$}
+    \State{\Call{tryConnectToRoadmap}{$\roadmap$,$newNodes$}}
+    \EndIf
+    \EndFunction
+    \Function{tryConnectNewNodes}{$\roadmap$, $nodes$}
+    \For{$\conf_1,\conf_2\mbox{ in } nodes,\ \conf_1\not=\conf_2$}
+    \State{$s_1\gets\Call{state}{\conf_1};s_2\gets\Call{state}{\conf_2}$}
+    \State{$p\gets\Call{Connect}{\conf_1,s_1,\conf_2,s_2}$}
+    \If{$p$}
+    \State{$\roadmap$.\Call{addEdge}{$\conf_1$,$\conf_2$,$p$}}
+    \EndIf
+    \EndFor
+    \EndFunction
+    \Function{tryConnectToRoadmap}{$\roadmap$,$nodes$}
+    \For{$\conf_1\mbox{ in } nodes$}
+    \State{$s_1\gets\Call{state}{\conf_1}$}
+    \For{$cc$ in connected components of $\roadmap$}
+    \If{$\conf_1\notin cc$}
+    \State{$near\gets K\mbox{ nearest neighbors of }\conf_1\mbox{ in }cc$}
+    \For{$\conf_2\mbox{ in }near$}
+    \State{$s_2\gets\Call{state}{\conf_2}$}
+    \State{$p\gets\Call{Connect}{\conf_1,s_1,\conf_2,s_2}$}
+    \If{$p$}{$\roadmap$.\Call{addEdge}{$\conf_1,\conf_2$}}\EndIf
+    \EndFor
+    \EndIf
+    \EndFor
+    \EndFor
+    \EndFunction
+    \Function{extend}{$s$, $\conf_{near}$, $\conf_{rand}$}
+    \State{$solver\gets$\Call{SolverBySubstitution}{}}
+    \State{$\transition\gets$ random edge getting out of $s$}
+    \State{$g\gets$ state $\transition$ points to}
+    \For{$c$ in $g$.\Call{constraints}{ }}
+    \State{$solver$.\Call{addConstraint}{$c(\conf)=0$}}
+    \EndFor
+    \For{$c$ in $\transition$.\Call{constraints}{ }}
+    \State{$solver$.\Call{addConstraint}{$c(\conf) = c(\conf_{near})$}}
+    \EndFor
+    \State{$\conf_{target}\gets solver$.\Call{solve}{$\conf_{rand}$}}
+    \If{$\conf_{target}$}
+    \State{$p\gets$ linear interpolation from $\conf_{near}$ to $\conf_{target}$}
+    \State{$p$.\Call{addConstraints}{$\transition$.{\scriptsize CONSTRAINTS}()}}
+    \EndIf
+    \If {$p$ collision free} {\Return $p$}
+    \Else { \Return collision-free portion of $p$ starting at $\conf_{near}$}
+    \EndIf
+    \EndFunction
+    \Function{connect}{$\conf_1,s_1,\conf_2,s_2$}
+    \State{parameter $\epsilon>0$}
+    \State{$p\gets$ linear interpolation from $\conf_1$ to $\conf_2$}
+    \State{$\transition\gets$\Call{transition}{$s_1$,$s_2$}}
+    \If{not $\transition$}{ \Return $\emptyset$}
+    \EndIf
+    \For{$c$ in $\transition$.\Call{constraints}{ }}
+    \If{$\|c(\conf_2) - c(\conf_1)\|\geq\epsilon$}{ \Return $\emptyset$}
+    \Else
+    \State{$p$.\Call{addConstraint}{$c(\conf)=c(\conf_1)$}}
+    \EndIf
+    \EndFor
+    \If {$p$ in collision} { \Return $\emptyset$}
+    \EndIf
+    \State{\Return $p$}
     \EndFunction
   \end{algorithmic}
-  \caption{}
+  \caption{Manipulation RRT algorithm iteratively calls method {\scriptsize ONE}{\small S}{\scriptsize TEP} until a solution path is found or the maximum number of iterations is reached.}
   \label{algo:M-RRT}
@@
 
-\subsection{Waypoint edges}
-
-\subsection{Crossed foliation issue}
-
-\subsection{Prototyping dedicated algorithms}
+\subsection{Examples}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{figures/example-qrand-2.png}
+    \includegraphics[width=\linewidth]{figures/example-q0-2.png}
+    \includegraphics[width=\linewidth]{figures/example-q-wp-2.png}
+  \end{center}
+  \caption{Example of extension along a transition of the constraint graph. Top
+    $\conf_{rand}$, middle $\conf_{near}$, bottom $\conf_{target}$.
+  }
+  \label{fig:MRRT-extension}
+\end{figure}
+
+In this section, we illustrate the algorithm described in the previous section with two examples.
+Figure~\ref{fig:MRRT-extension} illustrates function \EXTEND defined in the previous section on the example of Figure~\ref{fig:2ur3-cylinder}.
+The system considered is
+composed of two robots and a cylinder with two handles. The top picture displays
+$\conf_{rand}$. The middle picture displays $\conf_{near}$ that lies in state
+$(\emptyset,\emptyset)$. The transition that is randomly selected
+(Algorithm~\ref{algo:M-RRT}, line 33) is
+$(\emptyset,\emptyset) \rightarrow (1,\emptyset)$, meaning that robot 1 will
+try to grasp handle 1. According to tables~\ref{tab:state-constraints} and
+\ref{tab:transition-constraints}, the constraints of the transition are
+$(place_1, place_1/comp)$.
+The first one is of type~(\ref{eq:convex-shape-contact}), the second one is
+of type~(\ref{eq:convex-shape-contact-complement}) and is parameterized: the
+right hand side uniquely defines the contact surfaces (object and robot or environment) and the position of the
+object on the contact surface. $\conf_{target}$ is obtained by projecting $\conf_{rand}$
+onto the manifold defined by the following constraints
+(Algorithm~\ref{algo:M-RRT}, lines 35--39):
+\begin{itemize}
+\item $place_1$, $place_1/comp$ that belong to the transition,
+\item $grasp_{11}$ that belongs to the goal state.
+\end{itemize}
+According to Section~\ref{subsec:explicit}, the two first constraints can be
+replaced by an explicit constraint: the position of the object can be deduced
+from the right hand side of $place_1/comp$ that is initialized with configuration
+$\conf_{near}$.
+
+After substitution, the set of constraints is reduced to an implicit constraint
+on the configuration variables of robot 1 (6 variables). The solution found by the solver, $\conf_{target}$ (line 39) is displayed in
+Figure~\ref{fig:MRRT-extension} bottom. Notice that as expected, the position of the object is the same in $\conf_{target}$ as in $\conf_{near}$.
+
+The path returned by function \EXTEND is the linear interpolation between
+$\conf_{near}$ and $\conf_{target}$ constrained with $place_1$, $place_1/comp$ with
+right hand side initialized with $\conf_{near}$. As explained earlier, this
+constraint is replaced by an explicit constraint. Let us notice that the linear interpolation already satisfies the constraint, but this is not always the case.
+
+If the latter path is in collision, the collision-free part of the path starting
+at $\conf_{near}$ is returned.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.75\linewidth]{figures/example-qtarget-1.png}
+    \includegraphics[width=.75\linewidth]{figures/example-connect-q1.png}
+  \end{center}
+  \caption{Method \CONNECT applied to two configurations.}
+  \label{fig:connect}
+\end{figure}
+
+Figure~\ref{fig:connect} illustrates method \CONNECT applied to two configurations $\conf_1$ (top) and $\conf_2$ (bottom).
+Both configurations lie in state $(1,\emptyset)$\footnote{Note that $\conf_1$ lies at the intersection between states
+  $(\emptyset,\emptyset)$ and $(1,\emptyset)$.}.
+The transition between those states $(1,\emptyset)\rightarrow(1,\emptyset)$ contains the following constraints (tables~\ref{tab:state-constraints} and \ref{tab:transition-constraints}):
+\begin{itemize}
+\item $grasp_{11}/comp$, $grasp_{11}$.
+\end{itemize}
+Method \CONNECT checks that the right hand side of $grasp_{11}/comp$ is the same
+for $\conf_1$ and $\conf_2$, up to the error threhold (Algorithm~\ref{algo:M-RRT}, line~51).
+From a geometrical point of view, this means that the orientation of the cylinder along its axis, with respect to the gripper is the same in both configurations.
+Let us recall that the right hand side of $grasp_{11}$ is 0.
+If the condition is satisfied, the method builds the linear interpolation between $\conf_1$ and $\conf_2$ with the explicit constraint equivalent to $\{grasp_{11}/comp$, $grasp_{11}\}$ and returns this path if it is collision-free.
+
+\subsection{Waypoint transitions}
+
+\begin{figure}
+  \begin{center}
+    \def\svgwidth {\linewidth}
+    {
+      \graphicspath{{./figures/}}
+      \input {figures/grasp-waypoint.pdf_tex}
+    }
+    \includegraphics[width=.8\linewidth]{figures/dual-grasp-pregrasp.png}
+  \end{center}
+  \caption{Along a transition where an object already grasped is grasped by
+    an additional gripper, an intermediate waypoint state called
+    \textit{pregrasp} (pg) is added. This intermediate state is represented by
+    an hexagonal box. $\gripper{2} > \handle{2}|(1,\emptyset)$ means that
+    gripper 2 is going to grasp handle 2 from the state where gripper 1 grasps
+    handle 1. The constraints associated to this waypoint state are the
+    constraints of the state with the least active grasps (here $(1,\emptyset)$)
+    and the pregrasp constraint corresponding to the grasp that is added in
+    the node with the most active grasps (here $pregrasp_{22}$). The transition
+    constraints are the same for all transitions (in red) and the same as
+    the loop transition of the state with the least active grasps (in blue: here
+    $grasp_{11}$ and $grasp_{11}/comp$).}
+  \label{fig:pregrasp-grasp}
+\end{figure}
+\begin{figure}
+  \centerline {
+    \parbox {.69\linewidth} {
+      \def\svgwidth {\linewidth}
+      {
+        \graphicspath{{./figures/}}
+        \input {figures/placement-waypoints.pdf_tex}
+      }
+    }
+    \parbox {.3\linewidth} {
+      \begin {tabular}{c}
+      \includegraphics [width=.9\linewidth] {figures/placement-pregrasp-ur3.png}\\
+      \includegraphics [width=.9\linewidth] {figures/placement-grasp-ur3.png}\\
+      \includegraphics [width=.9\linewidth] {figures/placement-preplace-ur3.png}\\
+      \end{tabular}
+    }
+  }
+  \caption{Along a transition where an object in placement is grasped
+    by a gripper, we add three waypoint states called
+    \textit{pregrasp} (pg) where the gripper is above the object,
+    \textit{graps-placement} (gp) where the object is grasped but
+    still in placement and \textit{preplacement} (pp) where the object
+    is grasped above the contact surface. All transitions between the
+    state with the least active grasps and the waypoint $gp$ have the
+    same constraints as the loop transition of the state with the
+    least active grasps (here: $place_1$ and $place_1/comp$ in
+    blue). All transitions between the waypoint state $gp$ and the
+    state with the most active grasps have the same constraints as the
+    loop transition of the state with the most active grasps (here: $grasp_{11}$
+    and $grasp_{11}/comp$ in red).}
+  \label{fig:pregrasp-placement}
+\end{figure}
+
+A prehensile manipulation motion contains by definition configurations that are in contact:
+\begin{itemize}
+\item between gripper and object during grasp,
+\item between object and contact surface when the object lies in a stable pose.
+\end{itemize}
+Contacts are difficult to handle by classical collision detection libraries and are often considered as collisions.
+To overcome this issue, we keep the gripper open during grasp, and objects slightly above contact surfaces in stable poses.
+
+Even with these simple tricks however, solution paths to a manipulation problem need to come close to collision, raising the well-known issue of narrow passages.
+
+To cope with this issue, we define intermediate states in the
+constraint graph called waypoint states. These states are inserted
+between the regular states of the constraint graph. They require some prior
+definitions.
+
+\begin{definition}\label{def:pregrasp}\textbf{pregrasp}
+  A \emph{pregrasp} is a numerical constraint $h$ over $\CS$, defined by
+  \begin{itemize}
+  \item a gripper $\gripper$,
+  \item a handle $\handle$.
+  \item a non-negative real number $\Delta$.
+  \end{itemize}
+  Let $\bar{h}$ be the smooth mapping from $\CS$ to $\reals^6$ that maps to any configuration $\conf$ of the system, the expression
+  \begin{equation}\label{eq:barh-pregrasp}
+    \bar{h}(\conf) = \log_{\reals^3\times SO(3)}\left(\gripper^{-1}(\conf) \handle(\conf)\right) - \left(\Delta\, 0\, 0\, 0\, 0\, 0\right)^T.
+  \end{equation}
+  $h(\conf)$ is obtained by extracting from $\bar{h}$ the components the values of which are \true in the handle flag.
+\end{definition}
+Note that when this constraint is satisfied, the handle is translated
+along $x$ axis by distance $\Delta$ compared to a configuration
+satisfying the grasp constraint. The value of $\Delta$ depends on the geometry
+of the gripper and object. Clearance values are associated to the handle: $cl_{o}$ and to the gripper: $cl_{g}$. $\Delta$ is defined as $cl_{o}+cl_{g}$. The
+clearance parameters are part of the definition of the gripper and handle and
+are stored in the SRDF files.
+
+\begin{definition}\label{def:preplacement}\textbf{preplacement}
+  A preplacement is a numerical constraint $h$ over $\CS$, defined by
+  \begin{itemize}
+  \item $(o_i)_{i\in I}$ a set of convex polygons attached to an object,
+  \item $(f_j)_{j\in J}$ a set of convex polygons attached to the environment or to a mobile part of a robot that can receive objects (mobile robot for instance),
+  \item a non-negative real number $\Delta$.
+  \end{itemize}
+  with the same notation as in Section~\ref{subsec:placement}, we define $i$ and
+  $j$ as the indices that minimize $d(f_j,o_i)$~
+  (Equation~(\ref{eq:contact-distance})), and
+  \begin{equation}\label{eq:convex-shape-contact-hbar-preplacement}
+    \bar{h}(\conf) = \log_{\reals^3\times SO(3)} \left(f_{j}(\conf)^{-1}o_i(\conf)\right) + \left(\Delta\,0\,0\,0\,0\,0\right)^T
+  \end{equation}
+  The left hand side of the preplacement constraint is defined by Equation~
+  (\ref{eq:convex-shape-contact}).
+\end{definition}
+Note that when this constraint is satisfied, the object is translated of a
+distance $\Delta$ along the normal to the contact surface.
+
+We denote by
+\begin{itemize}
+\item $pregrasp_{ij}\ i\in\{1,\cdots,\ngrippers\}\ j\in\{1,\cdots,\nhandles\}$ the pregrasp constraint of handle $j$ by gripper $i$,
+\item $preplace_{i}\ i\in\{1,\cdots\nobjects\}$ the preplacement constraint of object $i$.
+\end{itemize}
+
+We replace the transitions of the constraint graph defined in
+Section~\ref{subsec:constraint graph} by a sequence of intermediate
+states and transitions as follows.
+
+Following definition~\ref{def:neighboring-states}, if two states $\state_1$
+and $\state_2$ are neighbors, one of them contains an additional grasp with
+respect to the other one. Without loss of generality, let us consider that
+$\state_2$ contains the additional grasp $\grasp{\gripper_i}{\handle_j}$,
+$i\in\{1,\cdots\ngrippers\}$, $j\in\{1,\cdots\nhandles\}$. Then there are two
+cases:
+\begin{enumerate}
+\item the object handle $\handle_j$ belongs to is already grasped in state
+  $\state_1$, or
+\item the object handle $\handle_j$ belongs to is in placement is
+  state $\state_1$.
+\end{enumerate}
+In case 1, we replace the transitions between $\state_1$ and $\state_2$ by an additional waypoint state and four waypoint transitions as explained in Figure~\ref{fig:pregrasp-grasp}.
+
+In case 2, we replace the transitions between $\state_1$ and $\state_2$ by three additional waypoint states and eight waypoint transitions as explain in Figure~\ref{fig:pregrasp-placement}.
+
+\begin{figure}
+  \centerline{
+    \def\svgwidth {\linewidth}
+      {\tiny
+       \graphicspath{{./figures/}}
+       \input {figures/waypoint-graph.pdf_tex}
+    }
+  }
+  \caption{Structure of the constraint graph corresponding to the
+    system in Section~\ref{subsec:ur3-cylinder} after inserting
+    waypoint transitions.  Waypoint transitions starting from/going to
+    $(\emptyset,\emptyset)$ contain three waypoint states. All other
+    waypoint transitions contain one waypoint state.}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.75\linewidth]{figures/example-q-wp-1.png}
+    \includegraphics[width=.75\linewidth]{figures/example-q-wp-2.png}
+    \includegraphics[width=.75\linewidth]{figures/example-q-wp-3.png}
+  \end{center}
+  \caption{Example of extension along the waypoint transition between states
+    $(\emptyset,\emptyset)$ and $(1,\emptyset)$. Each picture represents a
+    waypoint. The last waypoint lies in state $(1,\emptyset)$. $\conf_{near}$
+    and $\conf_{rand}$ are the same as in Figure~\ref{fig:MRRT-extension}.}
+  \label{fig:MRRT-extension-wp}
+\end{figure}
+
+\paragraph{Construction of a path along a waypoint transition} Function
+{\scriptsize EXTEND} in Algorithm~\ref{algo:M-RRT} builds a path along a
+transition from an initial configuration by projecting the configuration
+onto the sub-manifold defined by the goal state constraints and by the
+transition constraints. The right hand side of the transition constraint is
+priorly initialized with the initial configuration.
+
+A waypoint transition builds a path by defining a sequence of configurations
+that lie in the intermediate waypoint states, each waypoint is obtained by
+projecting the previous waypoint onto the corresponding manifold.
+Figure~\ref{fig:MRRT-extension-wp} proposes an example of extension along
+edge $(\emptyset,\emptyset)\rightarrow(1,\emptyset)$ from configuration
+$\conf_{near}$ (Figure~\ref{fig:MRRT-extension} middle). The edge contains three
+waypoints. The random configuration $\conf_{rand=}$ is displayed in
+Figure~\ref{fig:MRRT-extension} top. Table~\ref{table:waypoints} shows the
+waypoint configurations that are produced when extending $\conf_{near}$ toward
+$\conf_{rand}$, and the constraints applied to compute these configurations.
+\begin{table}
+{
+\begin{tabular} {|l|l|}
+  \hline
+  $\conf_{near}$ &
+  \begin{tabular}{l} \textbf{state} $(\emptyset,\emptyset)$ \\
+  \end{tabular}\\
+  \hline
+  $\conf_1$ &
+  \begin{tabular}{l} \textbf{waypoint state} $\gripper{1}>\handle{1}|(\emptyset,\emptyset)|pg$ \\
+    \textbf{constraints} \\
+    {\color{blue}$place_1(\conf)=0$} \\
+    {\color{blue}$place_1/comp(\conf) = place_1/comp(\conf_{near})$} \\
+    $pregrasp_{11}(\conf) = 0$ \\
+    \textbf{solver initialized with} $\conf_{rand}$
+  \end{tabular}\\
+  \hline
+  $\conf_2$ & 
+  \begin{tabular}{l}
+    \textbf{waypoint state} $\gripper{1}>\handle{1}|(\emptyset,\emptyset)|gp$ \\
+    \textbf{state} \\
+    $(1,\emptyset)$ \\
+    \textbf{constraints} \\
+    {\color{blue}$place_1(\conf)=0$} \\
+    {\color{blue}$place_1/comp(\conf) = place_1/comp(\conf_{near})$} \\
+    $grasp_{11}(\conf) = 0$ \\
+    \textbf{solver initialized with} $\conf_{1}$
+  \end{tabular} \\
+  \hline
+  $\conf_{target}$ &
+  \begin{tabular}{l} \textbf{state} $(1,\emptyset)$ \\
+    \textbf{waypoint state} $\gripper{1}>\handle{1}|(\emptyset,\emptyset)|pp$ \\
+    \textbf{constraints} \\
+    {\color{blue}$grasp_{11}(\conf)=0$} \\
+    {\color{blue}$grasp_{11}/comp(\conf) = grasp_{11}/comp(\conf_{2})$} \\
+    $preplace_{1}(\conf) = 0$\\
+    \textbf{solver initialized with} $\conf_{2}$
+  \end{tabular}\\
+  \hline
+\end{tabular}
+\caption{Waypoint configurations computed along edge $(\emptyset,\emptyset)$.
+  The resulting path between $\conf_{near}$ and $\conf_{target}$ is a concatenation
+  of constrained linear interpolation. Constraints applied between a waypoint
+  and its predecessor are shown in blue.}
+\label{table:waypoints}
+}
+\end{table}
+\paragraph{Implementation} From an implementation point of view, class
+\href{https://gepettoweb.laas.fr/hpp/hpp-manipulation/doxygen-html/classhpp_1_1manipulation_1_1graph_1_1WaypointEdge.html}{\texttt{WaypointEdge}} derives from class \href{https://gepettoweb.laas.fr/hpp/hpp-manipulation/doxygen-html/classhpp_1_1manipulation_1_1graph_1_1Edge.html}{\texttt{Edge}}.
+The waypoint configurations are computed by method\\ \texttt{generateTargetConfig} that is specialized in
+class\\ \texttt{WaypointEdge}.
+
+Note that waypoint states are internal to a waypoint edges and are not known by
+the constraint graph when determining in which state a configuration lies
+(Algorithm~\ref{algo:M-RRT} lines~17 and 23) and when visiting the states of the
+constraint graph (Algorithm~\ref{algo:M-RRT} line~8).
+
+\section{Humanoid Path Planner}\label{sec:humanoid-path-planner}
+
+In this section, we describe in more details the software platform
+\href{https://humanoid-path-planner.github.io/hpp-doc}{Humanoid Path
+  Planner} that implements the concepts and algorithms of the previous
+sections.
+
+Humanoid Path Planner is a collection of standard software packages that depend
+on each other. The main packages are the following:
+\begin{itemize}
+\item \texttt{hpp-fcl} a modified version of \href{https://github.com/flexible-collision-library/fcl}{\texttt{fcl}}. The main additional features are the following:
+  \begin{itemize}
+  \item computation of a lower bound of the distance when testing collision
+    between two objects. This is required for continuous collision detection,
+  \item security margins in collision checking,
+  \end{itemize}
+\item \href{https://stack-of-tasks.github.io/pinocchio}{\texttt{pinocchio}}~\cite{pinocchio} a library computing forward kinematics and dynamics for multi-body kinematic chains,
+\item \texttt{hpp-constraints} a library that implements numerical constraints
+  and solvers,
+\item \texttt{hpp-core} a library that implements most of the conceps relative
+  to motion planning. The main features are
+  \begin{itemize}
+  \item abstraction of paths in configuration spaces and some implementations,
+  \item abstaction of path planning and path optimzation and some implementations,
+  \item abstraction of steering methods and some implementations,
+  \item roadmaps,
+  \item validation of configurations and paths,
+  \end{itemize}
+\item \texttt{hpp-manipulation} a library that implements the manipulation
+  layer with
+  \begin{itemize}
+    \item composite kinematic chains composed of the robots and objects,
+    \item the constraint graph,
+    \item M-RRT algorithm,
+  \end{itemize}
+\item \texttt{hpp-manipulation-urdf} an extension of the SRDF parser to retrieve
+  information relative to objects, like the definition of grippers, handles, and
+  contact surfaces.
+\end{itemize}
+An HPP session consists of a standalone executable \texttt{hppcorbaserver}
+that implements CORBA services. These services can be extended via a plugin
+system. The application can then be controlled via python script or C++ code.
+CORBA clients are provided in python and C++. The packages implementing CORBA
+clients and servers are
+\begin{itemize}
+\item \texttt{hpp-corbaserver} for canonical path planning problems, and
+\item \texttt{hpp-manipulation-corba} for manipulation problems. This
+latter package also provides an implementation of the automatic constraint graph
+construction in python.
+\end{itemize}
+The environment used for path planning as well as the paths computed can be
+displayed using \texttt{gepetto-gui} through packages
+\begin{itemize}
+  \item\texttt{gepetto-viewer},
+  \item\texttt{gepetto-viewer-corba}, and
+  \item\texttt{hpp-gepetto-viewer}.
+\end{itemize}
+
+\subsection{Virtual machine}
+
+A virtual docker image can be downloaded to run, test and replicate the
+examples described in the next sections. After installing docker on a Linux machine, type
+
+\noindent
+{\small\texttt{xhost +\\docker run --rm -it \\ --name hpp --volume="/tmp/.X11-unix:/tmp/.X11-unix"\\ gitlab.laas.fr:4567/humanoid-path-planner/hpp-doc/\\ journal2020}}
+
+and follow the instructions.
+
+\subsection{Tutorial}
+
+\subsection{Benchmark}

### 2020-10-06 hpp.tex
Write state of the art and start experimental results.

@@
 \section{Introduction}
-This \paper describes a software platform taylored for manipulation planning in
-robotics.
+
+Robots in industrial manufacturing are today mostly programmed by
+hand. They repeat the same motion thousands of times with a high
+accuracy. Automating a task with some variability is however very
+challenging since it requires more programming effort to integrate
+sensors and some kind of motion planning in the process. The work
+described in this paper is a small step to the simplification of
+industrial process automation in the presence of some variability,
+like the initial position of some object or the presence of unknown
+obstacles. Although the work only covers the motion planning -- and
+more accurately the manipulation planning -- issue, we think that it
+is important not only to develop algorithms, but also to provide them
+within an open-source software platform in order to make the
+evaluation and then the integration of those algorithms easier.
+
+This \paper therefore describes a software platform called \href{https://humanoid-path-planner.github.io/hpp-doc}{Humanoid Path
+  Planner} taylored for manipulation planning in robotics. It can handle many types of robots, from manipulator arms to legged humanoid robots.
 The main contributions are:
 \begin{itemize}
-\item an original and general implementation of prehensile manipulation based on
+\item an original and general modelling of prehensile manipulation based on
   non-linear constraints,
@@
 
-The \paper is organized as follows.
+This paper extends the work presented in \cite{mirabel-iros2016} and \cite{MirLam2018} with some new material.
+
+The \paper is organized as follows. \Partie~\ref{sec:related-work} presents some related work for constrained motion planning and manipulation planning.
 \Partie~\ref{partie:preliminary} introduces some preliminary notions like kinematic chains and Lie groups that are used to model the configuration space of each joint.
@@
 \Partie~\ref{sec:manipulation-planning} provides a general algorithm that solves manipulation planning problems.
-Finally, \partie~\ref{sec:humanoid-path-planner} is devoted to a software platform implementing the notions introduced in the previous sections.
-
-Each \partie is implemented by one or several software packages. Instead of providing formulas for some values that need to be computed, we provide a link to the C++ implementation.
+Finally, \partie~\ref{sec:humanoid-path-planner} is devoted to the software platform implementing the notions introduced in the previous sections.
+Experimental results are provided for a large variety of problems.
+
+Each \partie is implemented by one or several software packages.
+For some values that need to be computed, instead of providing formulas, we sometimes provide a link to the C++ or python implementation.
 
@@
 partially known environments~\cite{DarpaUrbanChallenge} to path planning for
-deformable objects~\cite{LamKav2001},\\
-\cite{RouFerTai2020} and many other applications like coverage path planning~\cite{coverage2001},\\ \cite{coverage2013}, or pursuit evasion planning~\cite{pursuit-evasion1999}.
+deformable objects~\cite{LamKav2001}, \cite{RouFerTai2020} and many other applications like coverage path planning~\cite{coverage2001}, \cite{coverage2013}, or pursuit evasion planning~\cite{pursuit-evasion1999}.
 
 Planning motions for high dimensional robots like humanoid robots or multi-arm
-system has been proven to be of a very high complexity~\cite{SchSha1983},\\ \cite{Canny1983}. Starting in the 1990's random sampling methods have been proposed
+systems has been proven to be of a very high complexity~\cite{SchSha1983}, \cite{Canny1983}. Starting in the 1990's random sampling methods have been proposed
 to solve the problem, trading the completeness property against efficiency in
-solving problem in high dimensional configuration spaces\\ \cite{KSLO1996}, \cite{HsuLatMot99}, \cite{KufLav00}. Since then, asymptotically optimal random sampling algorithms have
+solving problems in high dimensional configuration spaces \cite{KSLO1996}, \cite{HsuLatMot99}, \cite{KufLav00}.
+The latter methods are said probabistically complete since the probability to find a solution if one exists converges to 1 when the time of computation tends to infinity.
+Since then, asymptotically optimal random sampling algorithms have
 been proposed~\cite{KarFra2011}.
@@
 
-Some systems are subject to non-linear constraints. These non-linear constraints
-define a sub-manifold of the configuration the robot must stay on. For instance,
-legged robots that must keep contact with the ground and enforce quasi-static
-equilibrium, or multi-arm systems grasping a same object are subject to this
-type of constraints. As the volume of the constrained manifold is usually equal
-to zero, sampling random configurations satisfying the constraints is an event
-of zero probability. To sample configurations on the constrained manifold,
-\cite{DalNakLamLau2009} and \cite{berenson2011} projects random configurations using a generalization of
-Newton-Raphson algorithm. Another way to sample configurations on a constrained
-manifold consist in expressing some configuration variables with respect to
-others~\cite{Cortes2002},\\ \cite{MirLam2018} when this is possible.\\ \cite{atlasRRT} propose another method based on non-linear projection. They cover the
-constrained manifold by growing an atlas composed of local charts. This
-approximation provides a probability distribution that is closer to the uniform
-distribution over the manifold than the projection of a uniform distribution
-over the configuration space.
+Some systems are subject to non-linear constraints. These non-linear
+constraints define a sub-manifold of the configuration space the robot must
+stay on. For instance, legged robots that must keep contact with the
+ground and enforce quasi-static equilibrium, or multi-arm systems
+grasping a same object are subject to this type of constraints. As the
+volume of the constrained manifold is usually equal to zero, sampling
+random configurations satisfying the constraints is an event of zero
+probability. To sample configurations on the constrained manifold,
+\cite{DalNakLamLau2009} and \cite{berenson2011} project random
+configurations using a generalization of Newton-Raphson
+algorithm. Another way to sample configurations on a constrained
+manifold consists in expressing some configuration variables with
+respect to others~\cite{Cortes2002}, \cite{MirLam2018} when this is
+possible. \cite{atlasRRT} propose another method based on non-linear
+projection. They cover the constrained manifold by growing an atlas
+composed of local charts. This approximation provides a probability
+distribution that is closer to the uniform distribution over the
+manifold than the projection of a uniform distribution over the
+configuration space. \cite{tb-rrt2016} propose a variation of the latter paper.
+The main difference resides in the fact that the nodes
+built on the tangent space are not immediately projected on the manifold.
+\cite{KinMolKav2019} provide a rich overview
+of the various approaches to motion planning with non-linear
+constraints.
 
@@
 Manipulation planning is a particular instance of path planning were
-some objects are moved by some robots. The configuration space of the
-whole system is constrained by non-linear constrained due to the fact
-that objects cannot move by themselves and should stay in a stable
-pose when not grasped by a robot.  The accessible configuration space
-is thus a union of sub-manifold as defined in the previous
-section. Each of these manifold is moreover a foliation where each
-leaf corresponds to a stable position of an object or to a grasp of an
-object by a gripper. The geometrical structure of the problem is now
-well understood~\cite{AlaSimLau89}. Although several instances of
+some objects are moved by some robots.
+Although several instances of
 manipulation problem exist like manipulation by
@@
 prehensile manipulation planning.
+The configuration space of the whole system is constrained by
+non-linear constraints due to the fact that objects cannot move by
+themselves and should stay in a stable pose when not grasped by a
+robot.  The accessible configuration space is thus a union of
+submanifolds as defined in the previous section. Each of these
+manifolds is moreover a foliation where each leaf corresponds to a
+stable position of an object or to a grasp of an object by a
+gripper. The geometrical structure of the problem is now well
+understood~\cite{AlaSimLau89}.
+
+The first attempt to solve manipulation planning problems using random sampling
+was proposed by \cite{simeon2004ijrr} where a reduction property reduces the
+complexity of the problem.
 
@@
 
-Rearrangement planning \cite{ota2004rearrangement}, \cite{LIS210}, consists in finding a sequence of manipulation paths that
-move some objects from an initial pose to a final pose. The final configuration
-of the robot is not specified. A simplifying asumption is the existence of a
-monotone solution, that is a solution where each object is grasped once only and
-is moved from its initial pose to its final pose \cite{stilman2008planning}, \cite{stilman2007manipulation}, \cite{srivastavaetal2014icra}, \cite{nieuwenhuisen2008effective}, \cite{ota2004rearrangement}.
-They mostly use two-level methods
-composed of a symbolic or task planner and of a motion planner~\cite{lavalle2006planning}.
+Rearrangement planning \cite{ota2004rearrangement}, \cite{LIS210},
+\cite{KroBek2015}, \cite{LerPha2015} consists in finding a sequence of
+manipulation paths that move some objects from an initial pose to a
+final pose. The final configuration of the robot is not specified. A
+simplifying asumption is the existence of a monotone solution, that is
+a solution where each object is grasped once only and is moved from
+its initial pose to its final pose \cite{stilman2008planning},
+\cite{stilman2007manipulation}, \cite{srivastavaetal2014icra}, \cite{nieuwenhuisen2008effective},
+\cite{ota2004rearrangement}. They mostly use two-level methods
+composed of a symbolic or task planner and of a motion
+planner~\cite{cambon:hal-01976081}.
+
+Other contributions in manipulation planning explicitely address the problem of multi-arm manipulation \cite{GhaCorSim2009}, \cite{HarTsuLau2014}, \cite{DobBek2015}, \cite{XiaLerPha2017}.
+
+\cite{schmitt19icra} propose an approach where two robots manipulate an object in a dynamic environment. The output of the algorithm is a sequence of controllers rather than a sequence of paths.
 
 \subsection{Open-source software platforms}
+
+Open-source software platforms are an important tool to make possible fair
+comparison between algorithms.
+Several software platforms are available for motion planning and/or manipulation
+planning in the robotics community.
+The most popular one is undoubtedly OMPL \cite{ompl2012}.
+OMPL integrates many randomized path planning algorithms and is widely used for teaching.
+Recently, \cite{KinMolKav2019} proposed an extension for systems subject to non-linear constraints.
+
+OpenRave~\cite{diankov_thesis} is
+a software platform that addresses motion and manipulation planning. It includes
+computation of forward kinematics.
+
+One of the main differences between our solution and the previously cited ones resides in the way manipulation constraints are compiled into a graph.
+To our knowledge none of the previous solutions can handle a variety of problems as large as those exposed in Section~\ref{sec:benchmarks}
 
@@
 \noindent
-{\small\texttt{xhost +\\docker run --rm -it \\ --name hpp --volume="/tmp/.X11-unix:/tmp/.X11-unix"\\ gitlab.laas.fr:4567/humanoid-path-planner/hpp-doc/\\ journal2020}}
-
+{\scriptsize\texttt{xhost +\\docker run --rm -it --name hpp \\--volume="/tmp/.X11-unix:/tmp/.X11-unix"\\ gitlab.laas.fr:4567/humanoid-path-planner/hpp-doc/\\ journal2020}}
+
+\noindent
 and follow the instructions.
 
-\subsection{Tutorial}
-
-\subsection{Benchmark}
+\subsection{Experimental results}\label{sec:benchmarks}
+
+In this section, we report on several experimental results obtained
+with HPP software on constrained motion planning and on manipulation
+planning problems. The raw data can be found in
+\href{https://github.com/humanoid-path-planner/hpp_benchmark/tree/v4.10.0/2020-07-23}{\texttt{hpp\_benchmark}} package. We present here only a few test cases.
+The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with 32 Bbytes of RAM and 3 Gbytes of cache memory.
+For each test case, we report the minimum, maximum, mean and standard deviation of the time of computation on the one hand, and of the number of nodes in the roadmap built to solve the problem on the other hand.
+
+\subsubsection{Constrained motion planning}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.49\linewidth]{figures/hrp2-q-init.png}
+    \includegraphics[width=.49\linewidth]{figures/hrp2-q-goal.png}
+  \end{center}
+  \caption{Constrained motion planning for HRP-2 humanoid robot sliding on the ground in quasi-static equilibrium: the feet should stay horizontal with a fixed relative position and the center of mass should project between the feet.
+The initial configuration is shown on the left.
+The goal configuration is shown on the right.
+The algorithm a constrained RRT close to the one described in~\cite{DalNakLamLau2009}}.
+  \label{fig:benchmark-hrp2}
+\end{figure}
+One test case concerns constrained motion planning.
+The robot is an HRP-2 humanoid robot in quasi-static equilibrium that can slide on the ground.
+This type of motion can be post-processed into a walking motion using the method described in \cite{dalibard:hal-00654175}.
+The results are displayed in Table~\ref{tab:benchmark-hrp2}.
+\begin{table}
+  \begin{center}
+  \begin{tabular}{|l|l|l|l|l|}
+    \hline
+    & min & max & mean & std dev \\
+    \hline
+    time (s) & 0.03 & 11.64 & 1.32 & 2.55 \\
+    nodes & 4 &  136 & 32.40 & 30.72\\
+    \hline
+  \end{tabular}
+  \end{center}
+  \caption{Experimental results for HRP-2 sliding on the ground: time of computation and number of nodes.}
+  \label{tab:benchmark-hrp2}
+\end{table}
+
+\subsubsection{Manipulation planning}
+
+In this section, we present some experimental results of manipulation planning
+problems obtained with M-RRT algorithm described in Section~\ref{sec:manipulation-planning}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{figures/baxter-manipulation-boxes}
+  \end{center}
+  \caption{Manipulation problem with Baxter robot manipulating two small boxes. The robot is requested to swap the boxes.}
+  \label{fig:baxter-manipulation-boxes}
+\end{figure}
+
+\begin{table}
+  \begin{center}
+  \begin{tabular}{|l|l|l|l|l|}
+    \hline
+    & min & max & mean & std dev \\
+    \hline
+    time (s) & 0.84 & 15.60 & 7.60 & 4.48 \\
+    nodes & 23 &  375 & 176.15 & 108.54 \\
+    \hline
+  \end{tabular}
+  \end{center}
+  \caption{Experimental results for Baxter robot manipulating two boxes on a table.}
+  \label{tab:baxter-manipulation-boxes}
+\end{table}
+
+The first test case features robot Baxter manipulating two boxes on a table (see Figure~\ref{fig:baxter-manipulation-boxes}).
+The boxes are swaped between the initial and final configurations.
+The robot has two grippers and each box is equipped with a handle.
+The constraint graph thus contains 7 nodes.
+The experimental results are displayed in Table~\ref{tab:baxter-manipulation-boxes}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{figures/pr2-manipulation-two-hands-init.png}
+    \includegraphics[width=\linewidth]{figures/pr2-manipulation-two-hands-goal.png}    
+  \end{center}
+  \caption{Manipulation planning problem with PR-2 robot manipulating a box. The robot needs to flip the box upside down from an initial pose (top) to a goal pose (bottom).}
+  \label{fig:pr2-manipulation-two-hands}
+\end{figure}
+
+\begin{table}
+  \begin{center}
+  \begin{tabular}{|l|l|l|l|l|}
+    \hline
+    & min & max & mean & std dev \\
+    \hline
+    time (s) & 0.92 & 9.62 & 3.30 & 2.47 \\
+    nodes & 6 &  111 & 32.90 & 31.63 \\
+    \hline
+  \end{tabular}
+  \end{center}
+  \caption{Experimental results for PR-2 robot manipulating a box on a table.}
+  \label{tab:pr2-manipulation-two-hands}
+\end{table}
+
+The second test case features robot PR-2 manipulating a box on a table.
+The robot is resquested to flip the box upside down from an initial pose to a goal pose as represented in Figure~\ref{fig:pr2-manipulation-two-hands}.
+The robot is equipped with two grippers, the box is equipped with two handles.
+The constraint graph contains 7 nodes.
+Table~\ref{tab:pr2-manipulation-two-hands} shows the experimental results.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.49\linewidth]{figures/romeo-placard-init.png}
+    \includegraphics[width=.49\linewidth]{figures/romeo-placard-goal.png}    
+  \end{center}
+  \caption{Manipulation planning problem with Romeo robot manipulating a placard. The robot needs to flip the placard by 180 degrees from an initial pose (left) to a goal pose (right).}
+  \label{fig:romeo-placard}
+\end{figure}
+
+\begin{table}
+  \begin{center}
+  \begin{tabular}{|l|l|l|l|l|}
+    \hline
+    & min & max & mean & std dev \\
+    \hline
+    time (s) & 4.64 & 554.18 & 151.49 & 158.64 \\
+    nodes & 27 &  2448 & 610.45 & 662.83 \\
+    \hline
+  \end{tabular}
+  \end{center}
+  \caption{Experimental results for Romeo robot manipulating a placard.}
+  \label{tab:romeo-placard}
+\end{table}
+
+The third test case features humanoid robot Romeo manipulating a placard. The
+robot is requested to rotate the placard by 180 degrees. The robot is equipped
+with two grippers and the placard with two handles. Each handle is associated
+to a unique gripper. The number of states of the contraint graph is thus 3.

### 2020-10-07 hpp.tex
Finish section on experimental results.

@@
 to a unique gripper. The number of states of the contraint graph is thus 3.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{figures/construction-set-init.png}
+    \includegraphics[width=\linewidth]{figures/construction-set-goal.png}    
+  \end{center}
+  \caption{Construction set: two robots are requested to assemble magnetic
+    spheres on a cylinder from an initial configuration (top) to a goal state
+    (bottom).}
+  \label{fig:construction-set}
+\end{figure}
+
+\begin{table}
+  \begin{center}
+  \begin{tabular}{|l|l|l|l|l|}
+    \hline
+    & min & max & mean & std dev \\
+    \hline
+    time (s) & 0.20 & 214.22 & 17.11 & 45.84 \\
+    nodes & 10 &  39 & 14.70 & 6.48 \\
+    \hline
+  \end{tabular}
+  \end{center}
+  \caption{Experimental results for construction set assembly.}
+  \label{tab:construction-set}
+\end{table}
+
+In the three previous test cases, the constraint graph was automatically built
+following Algorithm~\ref{algo:constraintGraphFactory}. If the number of grippers
+and handles increases, the number of states in the constraint may increase very
+quickly. Using python bindings, it is possible however to define constraint
+graphs with only the necessary states. We now present a test case that
+illustrates this possibility.
+
+In this example, an operator provides the sequence of actions (transitions)
+the system needs to go through:
+\begin{enumerate}
+\item robot 1 grasps sphere 1,
+\item robot 2 grasps cylinder 1,
+\item robot 1 sticks sphere 1 on cylinder 1,
+\item robot 1 releases sphere 1,
+\item robot 1 grasps sphere 2,
+\item robot 1 sticks sphere 2 on cylinder 1,
+\item robot 1 releases sphere 2,
+\item robot 2 put cylinder 1 on the ground.
+\end{enumerate}
+From this sequence of actions, the sequence of states visited is
+computed and only those states (9 in total) are built in the
+constraint graph. Then, a sequence of subgoals in the successive
+states is computed, in such a way that each sub-goal is accessible by
+the previous one (on the same leaf of the corresponding transition
+foliation). The subgoals are then linked by running a constrained visibility
+PRM algorithm (\cite{SimLauNis00}) on each leaf. The python code can be found
+on \href{https://github.com/humanoid-path-planner/hpp_benchmark/blob/master/2020-07-23/construction-set/script.py}{github.com}.
+
+Figure~\ref{fig:construction-set} displays the initial configuration and the
+goal state. Table~\ref{tab:construction-set} shows the experimental results.
+
+\subsubsection{Analysis}
+
+The experimental results show that M-RRT is able to solve a variety of
+manipulation problems including with a legged robots in quasi-static
+equilibrium. No parameter tuning is required between the different problems. All
+parameters are set to a default value for all test cases.
+
+As for any random motion planning method, we observe a large standard
+deviation within the 20 runs of each test case, both for the number of nodes
+and for the time of computation.
+
+We have also observed experimentally that the efficiency of M-RRT decreases
+when
+\begin{enumerate}
+\item the number of states to visit to solve a problem increases,
+\item the number of foliated states increases.
+\end{enumerate}
+For this reason, M-RRT is not able to solve the construction set
+problem in a reasonable amount of time. It remains to our knowledge
+the only algorithm in the literature able to solve a variety of
+problems as large as those exposed in this section.

### 2020-10-09 ijrr.tex
Add figures and scripts for benchmarks.

% sage_latex_guidelines.tex V1.20, 14 January 2017

\documentclass[Afour,sageh,times]{sagej}

%\usepackage{tabularx}
\usepackage{moreverb,url}
\usepackage{xspace}
%\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{yfonts}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,
  urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\input{new-command}
\newtheorem{definition}{Definition}

\def\volumeyear{2020}

\begin{document}

\runninghead{F.~Lamiraux et al}

\title{Prehensile Manipulation Planning: Modelling, Algorithms and Implementation}

\author{Florent Lamiraux\affilnum{1}, Joseph Mirabel\affilnum{1}}

\affiliation{
  \affilnum{1}LAAS-CNRS, University of Toulouse, France\\
}

\corrauth{Florent Lamiraux, LAAS-CNRS,
  7 avenue du Colonel Roche,
  Toulouse, France
}

\email{florent.lamiraux@laas.fr}

\begin{abstract}
\end{abstract}

\keywords{robotics, manipulation planning, constrained path planning,
  path planning}

\maketitle

\input{hpp}

\begin{acks}
This work has been partially supported by Airbus S.A.S. in the framework of the common laboratory Rob4Fam.
\end{acks}

%\bibliographystyle{authordate1}
\bibliographystyle{abbrvnat}

\bibliography{hpp}

\end{document}

### 2020-10-09 python/grasp_no_wp.py

@@ -1,0 +1,123 @@
+from math import sqrt
+from hpp import Transform
+from hpp.corbaserver.manipulation import ConstraintGraph, Constraints
+from hpp.corbaserver import Client
+Client ().problem.resetProblem ()
+from manipulation import robot, vf, ps, Ground, Box, Pokeball, PathPlayer, gripperName, ballName
+
+vf.loadEnvironmentModel (Ground, 'ground')
+vf.loadObjectModel (Pokeball, 'pokeball')
+robot.setJointBounds ('pokeball/root_joint', [-.4,.4,-.4,.4,-.1,2.,
+                                              -1.0001, 1.0001,-1.0001, 1.0001,
+                                              -1.0001, 1.0001,-1.0001, 1.0001,])
+
+
+q1 = [0, -1.57, 1.57, 0, 0, 0, .3, 0, 0.025, 0, 0, 0, 1]
+
+## Create constraint graph
+graph = ConstraintGraph (robot, 'graph')
+
+## Create constraint of relative position of the ball in the gripper when ball
+## is grasped
+ballInGripper = [0, .137, 0, 0.5, 0.5, -0.5, 0.5]
+ps.createTransformationConstraint ('grasp', gripperName, ballName,
+                                   ballInGripper, 6*[True,])
+
+## Create nodes and edges
+#  Warning the order of the nodes is important. When checking in which node
+#  a configuration lies, node constraints will be checked in the order of node
+#  creation.
+graph.createNode (['grasp', 'placement'])
+graph.createEdge ('placement', 'placement', 'transit', 1, 'placement')
+graph.createEdge ('grasp', 'grasp', 'transfer', 1, 'grasp')
+graph.createEdge ('placement', 'grasp', 'grasp-ball', 1, 'placement')
+graph.createEdge ('grasp', 'placement', 'release-ball', 1, 'grasp')
+
+## Create transformation constraint : ball is in horizontal plane with free
+## rotation around z
+ps.createTransformationConstraint ('placement', '', ballName,
+                                   [0,0,0.025,0, 0, 0, 1],
+                                   [False, False, True, True, True, False,])
+#  Create complement constraint
+ps.createTransformationConstraint ('placement/complement', '', ballName,
+                                   [0,0,0.025,0, 0, 0, 1],
+                                   [True, True, False, False, False, True,])
+
+ps.setConstantRightHandSide ('placement', True)
[…]

### 2020-10-09 python/grasp_with_wp.py

@@ -1,0 +1,91 @@
+from math import sqrt
+from hpp import Transform
+from hpp.corbaserver.manipulation import ConstraintGraph, \
+    ConstraintGraphFactory, Constraints
+from hpp.corbaserver import Client
+Client ().problem.resetProblem ()
+from manipulation import robot, vf, ps, Ground, Box, Pokeball, PathPlayer, gripperName, ballName
+
+vf.loadEnvironmentModel (Ground, 'ground')
+vf.loadObjectModel (Pokeball, 'pokeball')
+robot.setJointBounds ('pokeball/root_joint', [-.4,.4,-.4,.4,-.1,2.,
+                                              -1.0001, 1.0001,-1.0001, 1.0001,
+                                              -1.0001, 1.0001,-1.0001, 1.0001,])
+
+
+q1 = [0, -1.57, 1.57, 0, 0, 0, .3, 0, 0.025, 0, 0, 0, 1]
+
+## Create constraint graph
+graph = ConstraintGraph (robot, 'graph')
+
+# Required calls
+factory = ConstraintGraphFactory (graph)
+factory.setGrippers (['ur5/gripper'])
+factory.setObjects (["pokeball", ], [['pokeball/handle'],], [['pokeball/bottom'],])
+# Optionally
+factory.environmentContacts (['ground/surface'])
+factory.generate ()
+# graph is initialized
+
+ps.selectPathValidation ("Dichotomy", 0)
+ps.selectPathProjector ("Progressive", 0.1)
+graph.initialize ()
+
+## Project initial configuration on state 'placement'
+res, q_init, error = graph.applyNodeConstraints ('free', q1)
+q2 = q1 [::]
+q2 [7] = .2
+
+## Project goal configuration on state 'free'
+res, q_goal, error = graph.applyNodeConstraints ('free', q2)
+
+## Define manipulation planning problem
+ps.setInitialConfig (q_init)
+ps.addGoalConfig (q_goal)
+
+# v = vf.createViewer ()
+# pp = PathPlayer (v)
+# v (q1)
+
+## Build relative position of the ball with respect to the gripper
+# for i in range (100):
+#   q = robot.shootRandomConfig ()
+#   res,q3,err = graph.generateTargetConfig ('grasp-ball', q_init, q)
+#   if res and robot.isConfigValid (q3): break;
+
+# if res:
+#   robot.setCurrentConfig (q3)
[…]

### 2020-10-09 python/parse_benchmark.py

@@ -1,0 +1,30 @@
+import numpy as np
+import datetime as dt
+def parse(filename):
+    times = list()
+    nnodes = list()
+    with open(filename, 'r') as f:
+        state = "time"
+        while state != "finished":
+            line = f.readline()
+            if line[:7] == "Average":
+                break
+            if state == "time":
+                time = line.split(":")
+                assert (len(time) == 3)
+                times.append(dt.timedelta(hours=int(time[0]),
+                                          minutes=int(time[1]),
+                                          seconds=float(time[2])).total_seconds())
+                state="number nodes"
+            elif state == "number nodes":
+                nnodes.append(int(line[14:].rstrip('\n')))
+                state = "time"
+    return np.array(times), np.array(nnodes)
+
+def print_stat(times, nnodes):
+    print("time (s) & {:.2f} & {:.2f} & {:.2f} & {:.2f} \\\\".format(min(times), max(times),
+                                                     np.mean(times),
+                                                     np.std(times)))
+    print("nodes & {} &  {} & {:.2f} & {:.2f}".format(min(nnodes), max(nnodes),
+                                                np.mean(nnodes),
+                                                np.std(nnodes)))

### 2020-10-09 script/generate_pictures.py

@@ -1,0 +1,16 @@
+e = 'r0/gripper > object/handle1 | f'
+for i in range(10000):
+    qrand = robot.shootRandomConfig()
+    res,q,err = cg.generateTargetConfig(e, q0, qrand)
+    q[6:12] = q_init[6:12]
+    if not res: continue
+    res, pid, msg = ps.directPath(q_init, q, False)
+    if res: break
+
+e = 'r1/gripper > object/handle2 | 0-0'
+for i in range(10000):
+    qrand = robot.shootRandomConfig()
+    res,q,err = cg.generateTargetConfig(e, q1, qrand)
+    if not res: continue
+    res, pid, msg = ps.directPath(q1, q, False)
+    if res: break

### 2020-10-09 hpp.tex
Update hpp.tex.

@@
 built on the tangent space are not immediately projected on the manifold.
-\cite{KinMolKav2019} provide a rich overview
-of the various approaches to motion planning with non-linear
-constraints.
+\cite{CefOri2019} propose a general framework to plan task constrained motions
+in the presence of moving obstacle.
+\cite{KinMolKav2019} provide an in-depth review of the various approaches to
+motion planning with non-linear constraints.
 
@@
 composed of a symbolic or task planner and of a motion
-planner~\cite{cambon:hal-01976081}.
+planner~\cite{cambon:hal-01976081}, \cite{KaeLoz2013}, \cite{18-toussaint-RSS}.
 
@@
 \cite{schmitt19icra} propose an approach where two robots manipulate an object in a dynamic environment. The output of the algorithm is a sequence of controllers rather than a sequence of paths.
+
+Our work shares many ideas with \cite{HauNgt2011} where the notion of constraint
+graph is present, although not as clearly expressed as in this paper. The main
+contribution of our work with respect to this latter paper is that the
+constraint graph is built automatically at the cost of a more restricted range
+of applications. We address indeed only prehensile manipulation.
 
@@
 
-\subsection{Waypoint transitions}
+\subsection{Waypoint transitions}\label{subsec:waypoint}
 
@@
 
+\subsubsection{Influence of waypoint transitions}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{figures/ur5-grasps-pokeball.png}
+  \end{center}
+  \caption{UR-5 robot manipulating a ball lying on a plane. The robot is requested to pick the ball and place it a few centimeters aside.}
+  \label{fig:ur5-pokeball}
+\end{figure}
+
+\begin{table}
+  \begin{center}
+  \begin{tabular}{|l|l|l|l|l|}
+    \hline
+    & min & max & mean & std dev \\
+    \hline
+    with waypoints&&&&\\
+    time (s) & 0.01 & 0.49 & 0.12 & 0.14 \\
+    nodes & 4 &  30 & 9.75 & 7.26\\
+    \hline
+    without waypoints&&&&\\
+    time (s) & 4.15 & 73.53 & 26.24 & 17.07 \\
+    nodes & 97 &  1609 & 711.55 & 407.98\\
+    \hline
+  \end{tabular}
+  \end{center}
+  \caption{UR-5 manipulating a box with and without waypoint transitions}
+  \label{tab:waypoint}
+\end{table}
+
+All the previous experimental results have been obtained using
+waypoint transitions as described in Section~\ref{subsec:waypoint}. We
+now empirically show the positive effect of waypoints on the the
+efficiency of manipulation planning. For that, we run 20 times
+Algorithm M-RRT on the same problem with and without waypoint
+transitions.  The problem is defined by a UR-5 robot manipulating a
+ball as shown in Figure~\ref{fig:ur5-pokeball}. The results are
+reported in Table~\ref{tab:waypoint}. We can notice that for this
+example, the waypoint transitions make the time of computation and the
+number of nodes decrease by 2 orders of magnitude. This can be
+explained by the fact that in grasp configurations, the gripper is
+very close to the object and only a small part of the approaching
+directions of the gripper toward the object leads to collision-free
+paths. Waypoint states on the contrary are away from obstacles and
+easier to reach. The transition between the pregrasp waypoint and the
+grasp $\cap$ placement waypoint is always collision-free.
+
 \subsubsection{Analysis}

### 2020-10-14 hpp.tex
Add docker directory.

@@
 A virtual docker image can be downloaded to run, test and replicate the
-examples described in the next sections. After installing docker on a Linux machine, type
-
-\noindent
-{\scriptsize\texttt{xhost +\\docker run --rm -it --name hpp \\--volume="/tmp/.X11-unix:/tmp/.X11-unix"\\ gitlab.laas.fr:4567/humanoid-path-planner/hpp-doc/\\ journal2020}}
-
-\noindent
-and follow the instructions.
+examples described in the next sections. An archive is provided with this paper
+inflate this archive and follow instructions in the README file.
 

### 2020-10-14 docker/config.sh

@@ -1,0 +1,8 @@
+export PATH=/opt/openrobots/bin:$PATH
+export LD_LIBRARY_PATH=/opt/openrobots/lib
+export PYTHONPATH=/opt/openrobots/lib/python3.6/site-packages
+export ROS_PACKAGE_PATH=/opt/openrobots/share:/opt/ros/melodic/share
+export CMAKE_PREFIX_PATH=/opt/openrobots:/opt/ros/melodic
+export PKG_CONFIG_PATH=/opt/openrobots/lib/pkgconfig
+
+alias python='python3'

### 2020-10-14 hpp.tex
Write conclusion and set doublespace mode.

@@
 problems as large as those exposed in this section.
+
+\section{Conclusion and future work}
+
+This paper presents a software platform aimed at prototyping and solving
+prehensile manipulation planning problems. The platform provide an original
+algorithm M-RRT that is an extension of RRT exploring the leaves of the
+foliation defined by the manipulation constraints. The automatic insertion
+of way point states make the resolution more efficient and the resulting paths
+more natural.
+
+We think that this platform is a perfect basis for researchers who want to
+develop and benchmark new manpulation planning algorithms.
+
+To show the maturity of the project, we provide a docker image embarking the
+software.
+
+As a future work, we aim at working on general manipulation planning algorithms
+that can handle use cases as diverse as those proposed in the benchmark section.
+A good candidate is a generalization of RMR*~\cite{schmitt17icra}. We also aim at working on manipulation path optimization since paths computed by
+random algorithms are too long to be applied on real robots as such.

### 2020-10-19 hpp.tex
[biblio] Add reference to SoT and Hauser.

@@
 solving problems in high dimensional configuration spaces \cite{KSLO1996}, \cite{HsuLatMot99}, \cite{KufLav00}.
-The latter methods are said probabistically complete since the probability to find a solution if one exists converges to 1 when the time of computation tends to infinity.
+The latter methods are said probabilistically complete since the probability to find a solution if one exists converges to 1 when the time of computation tends to infinity.
 Since then, asymptotically optimal random sampling algorithms have

### 2020-10-19 hpp.tex
Small improvements.

@@
 challenging since it requires more programming effort to integrate
-sensors and some kind of motion planning in the process. The work
-described in this paper is a small step to the simplification of
-industrial process automation in the presence of some variability,
-like the initial position of some object or the presence of unknown
-obstacles. Although the work only covers the motion planning -- and
-more accurately the manipulation planning -- issue, we think that it
+sensors and motion planning in the process. The work described in this
+paper is a small step to the simplification of industrial process
+automation in the presence of some variability, like the variation of
+the initial position of some object or the presence of unknown
+obstacles. The work only covers the motion planning -- and
+more accurately the manipulation planning -- issue. The integration in a whole
+process is still under development. We think that it
 is important not only to develop algorithms, but also to provide them
@@
 \item an open-source software suite that implements all the above following state of the art development tools and methods.
+\item a docker image of the above software with installation instructions provided with this paper.
 \end{itemize}
@@
 \cite{TonDelPetParManMan2018}, in this paper, we are only concerned with
-prehensile manipulation planning.
+prehensile manipulation.
 The configuration space of the whole system is constrained by
@@
 submanifolds as defined in the previous section. Each of these
-manifolds is moreover a foliation where each leaf corresponds to a
+manifolds may moreover be a foliation where each leaf corresponds to a
 stable position of an object or to a grasp of an object by a
-gripper. The geometrical structure of the problem is now well
-understood~\cite{AlaSimLau89}.
+gripper. The geometrical structure of the problem has been well
+understood for a long time~\cite{AlaSimLau89}.
 
@@
 Navigation Among Movable Obstacles (NAMO) \cite{wilfong1988motion}, \cite{stilman2008planning} consists in finding
-a path for a robot that needs to displace objects in order to move to a goal
+a path for a robot that needs to move objects in order to reach a goal
 configuration. The final position of the objects does not matter in this case.
@@
 simplifying asumption is the existence of a monotone solution, that is
-a solution where each object is grasped once only and is moved from
+a solution where each object is grasped at most once and is moved from
 its initial pose to its final pose \cite{stilman2008planning},
@@
 One of the main differences between our solution and the previously cited ones resides in the way manipulation constraints are compiled into a graph.
-To our knowledge none of the previous solutions can handle a variety of problems as large as those exposed in Section~\ref{sec:benchmarks}
+To our knowledge none of the previous solutions can handle a variety of problems as large as those exposed in Section~\ref{sec:benchmarks}.
 
@@
 Representing a freeflying object by three virtual translations followed by
-three virtual rotations called roll, pitch and yaw is indeed a poor workaround.
+three virtual rotations called roll, pitch and yaw is indeed a poor workaround
+due to the presence of singularities. A good illustration of this is the gimball
+lock issue that arose during Apollo~13 flight. To avoid singularities, we
+propose the following defintion.
 
@@
 \item unbounded \textit{rotation} with configuration space  $SO(2)$,
-\item \textit{planar} with configuration space $SE(2)$,
-\item \textit{freeflyer} with configuration space $SE(3)$.
+\item \textit{planar} joint with configuration space $SE(2)$,
+\item \textit{freeflyer} joint with configuration space $SE(3)$.
 \end{itemize}
 $SO(n)$ and $SE(n)$ respectively stand for \textit{special orthogonal group} and \textit{special Euclidean group}.
-They represents the group of rigid-body transformations and the group of rotations in $\reals^n$.
+They represent the group of rotations and the group of rigid-body transformations in $\reals^n$.
 
@@
 
-The joint configuration spaces enumerated in the previous paragraph: $\mathbb{R}^n$, $SE(n)$ are all Lie groups.
+The joint configuration spaces enumerated in the previous paragraph: $\mathbb{R}^n$, $SO(n)$, and $SE(n)$ are all Lie groups.
 The group operation is $+$ for $\mathbb{R}^n$, and composition denoted as $"."$ for $SE(n)$.
 We refer to~\cite{LMS94} -- Appendix A for a thorough definition of Lie groups.
-We expose here only properties that are useful for the sequel of the \paper.
+We expose here only properties that are useful for the following developments.
 
@@
 Item 1 above is called the \textit{transport} of velocity $\mathbf{v}$ to $g$.
-Item 3 above is called the exponential map of $\liegroup$ and is denoted by $\exp$.
+Item 3 is called the exponential map of $\liegroup$ and is denoted by $\exp$.
 
@@
 Note that $\mathbf{x}$ is also the coordinate vector of $g(\mathbf{x})$ in the moving frame $g$.
-If $\mathbf{v}$, $\omega$ are a linear and an angular velocities at the origin, $(\mathbf{v},\omega)$ is transported to $g$
+If $\mathbf{v}$, $\omega$ are a linear and an angular velocities at the origin, $(\mathbf{v},\omega)$ is \textit{transported} to $g$
   as the same linear and angular velocities expressed in the moving frame. In other words, if
@@
   with $R\in SO(3)$ and $\mathbf{t}\in\reals^3$  is the homogeneous matrix
-  representing $g$, and $(\mathbf{v},\mathbf{\omega})$ is a velocity in $T_{I_3}SE(3)$, the velocity transported to $g$ corresponds to the linear velocity $R\mathbf{v}$ and $R\omega$ of the moving frame.
+  representing $g$, and $(\mathbf{v},\mathbf{\omega})$ is a velocity in $T_{I_3}SE(3)$, the velocity \textit{transported} to $g$ corresponds to linear and angular velocities $R\mathbf{v}$ and $R\omega$ of the moving frame.
   Integral curves of the vector field mentioned in item 2 above correspond to helicoidal motions of constant velocity expressed in the moving frame.
@@
 
-Therefore elements of $SE(3)$ are represented by a vector in $\reals^7$ where the 3 first components represent the image of the origin (vector $\mathbf{t}$ in Equation~\ref{eq:homogeneous matrix}), the 4 last components represent a unit quaternion.
-
-Elements of $SO(3)$ are represented by a unit vector of dimension 4.
+Therefore elements of $SE(3)$ are represented by a vector in $\reals^7$ where the 3 first components represent the image of the origin (vector $\mathbf{t}$ in Equation~\ref{eq:homogeneous matrix}), the 4 last components $(x,y,z,w)$ represent unit quaternion $w + x {i} + y {j} + z {k}$.
+
+Elements of $SO(3)$ are likewise represented by a unit vector of dimension 4.
 
@@
 
-As expressed earlier, following a constant velocity $\dot{\conf}$ from the neutral element of a joint configuration space leads to another configuration denoted as
+As expressed earlier, following a constant velocity\footnote{More precisely, following the vector field generated by $\dot{\conf}\in T_{\neutral}\liegroup$ according to the Lie group structure} $\dot{\conf}$ from the neutral element of a joint configuration space leads to another configuration denoted as
 $$
@@
 since the Lie group operator of $\reals$ is $+$ and $\exp_{\reals}$ is the identity.
-In order to homogeneize notation, we define the following operators:
+In order to homogeneize notation, we define the following operators. For any
+$\conf_0,\conf_1\in\liegroup$ and $\dconf\in T_{\neutral}\liegroup$:
 \begin{eqnarray}\label{eq:plus}
-  \conf_0 \oplus \dconf &\triangleq& \conf_0.\exp(\dconf),\\
+  \conf_0 \oplus \dconf &\triangleq& \conf_0.\exp(\dconf) \in \liegroup,\\
   \label{eq:minus}
-  \conf_1 \ominus \conf_0 &\triangleq& \log(\conf_0^{-1}.\conf_1).
+  \conf_1 \ominus \conf_0 &\triangleq& \log(\conf_0^{-1}.\conf_1) \in T_{\neutral}\liegroup.
 \end{eqnarray}
@@
 %% begin impl
-Cartesian products of Lie groups are represented by class \href{https://gepettoweb.laas.fr/hpp/hpp-pinocchio/doxygen-html/classhpp_1_1pinocchio_1_1LiegroupSpace.html}{\texttt{LiegroupSpace}}.
+Cartesian products of Lie groups are represented by Class \href{https://gepettoweb.laas.fr/hpp/hpp-pinocchio/doxygen-html/classhpp_1_1pinocchio_1_1LiegroupSpace.html}{\texttt{LiegroupSpace}}.
 Elements of these spaces are represented by classes
@@
 It may be useful to use a non zero right hand side for the same function $h$. We define for that parameterized non-linear constraints.
-\begin{definition}\textbf{Parameterized non-linear constraints.}
+\begin{definition}\textbf{Parameterized non-linear constraint.}
   A parameterized non-linear constraint is defined by a differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and by a vector $\mathbf{h}_0$ of $\reals^m$ and is written
@@
 %% begin impl
-Differentiable functions are represented by class\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1DifferentiableFunction.html}{\texttt{DifferentiableFunction}}.
+Differentiable mappings are represented by abstract Class\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1DifferentiableFunction.html}{\texttt{DifferentiableFunction}}.
 %% end impl
@@
 In this \paper, we will make use of the term Jacobian in a generalized way.
-If $h$ is a differentiable function from a Lie group $\liegroup_1$ to a Lie group $\liegroup_2$, and $\conf_1$ an element of $\liegroup_1$, we will denote by $\frac{\partial h}{\partial\conf}(\conf_1)$ the operator that maps velocities in $T_{\conf_1}\liegroup_1$ to the corresponding velocity of the image by $h$ in $T_{h(\conf_1)}\liegroup_2$.
+If $h$ is a differentiable function from a Lie group $\liegroup_1$ to a Lie group $\liegroup_2$, and $\conf_1$ an element of $\liegroup_1$, we will denote by $\frac{\partial h}{\partial\conf}(\conf_1)$ the operator that maps velocities in $T_{\conf_1}\liegroup_1$ to the velocity in $T_{h(\conf_1)}\liegroup_2$ \textit{transported by} $h$\footnote{If $\dconf\in T_{\conf_1}\liegroup_1$ is a velocity along a time parameteriezd curve $\gamma$, $\frac{\partial h}{\partial\conf}(\conf_1)\dconf$ is the velocity at
+  the same time along Curve $h(\gamma)$.}.
 
@@
 \end{equation}
-where $.^{+}$ denotes the Moore Penrose pseudo inverse, and $\alpha_i$ is a positive real number called the step size.
+where $.^{+}$ denotes the Moore Penrose\footnote{He has just been awarded the Nobel Prize.} pseudo inverse, and $\alpha_i$ is a positive real number called the step size.
 Taking $\alpha_i=1$ exactly solves the linear approximation, but it may not be the best choice in general.
@@
 In manipulation planning applications, where robots manipulate objects, once an object is grasped, the position of the object can be computed explicitely from the configuration of the robot.
-In this case, some configuration variables of the system depends on other:
+In this case, some configuration variables of the system depend on other:
 $$
@@
 $nc$ is the size of the latter.
-$args$ is an array that stores for each configuration variable, the index in $explicit$ of the constraint that computes this configuration variable, -1 if no constraints computes the index.
+$args$ is an array that stores for each configuration variable, the index in $explicit$ of the constraint that computes this configuration variable, -1 if no constraint computes the index.
 Procedure \href{https://github.com/humanoid-path-planner/hpp-constraints/blob/2555dfec945575f824bd973c30c5c13ec3b67645/src/explicit-constraint-set.cc\#L144}{ADD} tests whether explicit constraint $E$ is compatible with the previously inserted constraints.
 Line 6 checks whether any output variable of $E$ is already computed by a previous explicit constraint.
-If so the procedure returns failure and the $E$ is not inserted.
+If so the procedure returns failure and $E$ is not inserted.
 The Loop at line 9 recursively checks that any element of $out$ is not an input variable of a previously inserted constraint.
-If the loop ends without returning failure,  line 18 stores that elements of $out$ are computed by $E$ and $E$ is inserted in vector of constraints.
+If the loop ends without returning failure,  line 18 stores that elements of $out$ are computed by $E$ and $E$ is inserted in the vector of constraints.
 Function \href{https://github.com/humanoid-path-planner/hpp-constraints/blob/2555dfec945575f824bd973c30c5c13ec3b67645/src/explicit-constraint-set.cc\#L301}{computeOrder} at line 20 recursively computes the order in which the explicit constraints are evaluated, following the rule that the input of a constraint should be evaluated before the output.}
@@
   \label{eq:substitution-3}
-  \conf &=& (\conf_{in},\conf_{out}).
+  {in}\cap{out}=\emptyset.
 \end{eqnarray}
@@
 % begin impl
-The solver by substitution described in this section is implemented by class \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}, that store an instance of \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1ExplicitConstraintSet.html}{\texttt{ExplicitConstraintSet}}.
+The solver by substitution described in this section is implemented by Class \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}, that store an instance of \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1ExplicitConstraintSet.html}{\texttt{ExplicitConstraintSet}}.
 
 \paragraph{Important remark} As mentioned in Table~\ref{tab:vector representation}, the configuration and velocity vectors may have different sizes.
-As a consequence, index sets $in$ and $out$ in Definition~\ref{def:Explicit} correspond to configuration vector indices, while in Expression~(\ref{eq:jacobian-explicit}) they correspond to velocity vector indices.
+As a consequence, index sets $in$ and $out$ in Definition~\ref{def:Explicit} correspond to configuration vector indices, while in Expression~(\ref{eq:jacobian-explicit}), they correspond to velocity vector indices.
 To keep notation simple, we use the same notation for different sets.
@@
 $$
-Paths are implemented by class \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1Path.html}{\texttt{Path}}. Several implementations of unconstrained paths are provided:\\
+Paths are implemented by Class \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1Path.html}{\texttt{Path}}. Several implementations of unconstrained paths are provided:\\
 \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1StraightPath.html}{\texttt{StraightPath}} for linear interpolation generalized to Lie groups, \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1ReedsSheppPath.html}{\texttt{ReedsSheppPath}}, \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1DubinsPath.html}{\texttt{DubinsPath}} for non\-holo\-nomic mobile robots.
@@
 Before inserting a projected path in a roadmap for instance, it is therefore necessary to detect possible discontinuities.
-\cite{mirabel:hal-01360409} proposes two algorithms to check whether a projected path is continuous.
+\cite{Hauser-RSS-13} proposes a solution to this problem.
+\cite{mirabel:hal-01360409} describe two algorithms to check whether a projected path is continuous.
+These algorithms are implemented by
+classes \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1pathProjector_1_1Dichotomy.html}{\texttt{pathProjector::Dichotomy}} and \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1pathProjector_1_1Progressive.html}{\texttt{pathProjector::Progressive}}.
 
@@
   \item one or several objects,
+  \item a set of possible grasps,
   \item environment contact surfaces,
@@
 \end{definition}
-We will now provide precise definitions for gripper, grasp and stable contact pose.
+We will now provide precise definitions for grippers, grasps and stable contact poses.
 
@@
 \begin{itemize}
-\item If $flags=(\true,\true,\true,\true,\true,\true)$ the grasp is satisfied if $\gripper(\conf)$ and $\handle(\conf)$ coincide:
+\item If $flags=(\true,\true,\true,\true,\true,\true)$ the grasp is satisfied iff $\gripper(\conf)$ and $\handle(\conf)$ coincide:
   $$ h =\bar{h},$$
   $h_{comp}$ is an empty constraint;
-\item if $flags=(\true,\true,\true,\true,\true,\false)$ the grasp is satisfied if the centers and $z$ axes of $\gripper(\conf)$ and $\handle(\conf)$ coincide (free rotation around $z$).
+\item if $flags=(\true,\true,\true,\true,\true,\false)$ the grasp is satisfied iff the centers and $z$ axes of $\gripper(\conf)$ and $\handle(\conf)$ coincide (free rotation around $z$).
 This is useful for cylindical objects:
@@
   \end{align*}
-\item if $flags=(\true,\true,\true,\false,\false,$ $\false)$ the grasp is satisfied if the centers of $\gripper(\conf)$ and $\handle(\conf)$ coincide (free rotation).
+\item if $flags=(\true,\true,\true,\false,\false,$ $\false)$ the grasp is satisfied iff the centers of $\gripper(\conf)$ and $\handle(\conf)$ coincide (free rotation).
 This is useful for spherical objects.
@@
 contains all the configurations that are reachable from $\conf_0$ while keeping the grasp.
+Note that this representation of relative pose constraints has been used in the Stack of Task software, although it is not described in the corresponding paper~\cite{SoT}.
+It is different from Task Space Regions \cite{berenson2011} where open domains of $SE(3)$ are defined.
 
@@
 
-When an object is not grasped, it should lie in a stable pose. There are two simple methods to do that:
+When an object is not grasped, it should lie in a stable pose. There are two simple methods to enforce that:
 \begin{enumerate}
@@
 \end{align*}
-Figure~\ref{fig:convex-shape-contact} illustrates this notation.
+Figure~\ref{fig:convex-shape-contact} illustrates this definition.
 
@@
 constitute an explicit constraint since the pose of the object grasped
-uniquely depends on the configuration of the robot that grasps.
+uniquely depends on the configuration of the robot that grasps the object.
 
@@
 
-According to Definition~\ref{def:manipulation-problem}, the set of admissible configurations is the union of sub-manifolds of the configuration space of the system.
+According to Definition~\ref{def:manipulation-problem}, the set of admissible configurations of a manipulation problem is the union of sub-manifolds of the configuration space of the system.
 Each sub-manifold is defined by grasp and/or stable contact constraints.
@@
   \item the nodes of which are states defined by subsets of grasps~(\ref{eq:state}),
-  \item an edge connects two states if and only if they are neighboring or they are equal. Edges are also called \textit{transitions}.
-  \end{itemize}
+  \item two edges (back and forth) connect two states if they are neighboring,
+  \item one edge connect each state to itself.
+  \end{itemize}
+  Edges are also called \textit{transitions}.
   Nodes contain
@@
   \end{center}
-  \caption{Top: two robot UR3 with one gripper each (X=red, Y=gree, Z=blue) manipulating a cylinder with two handles.
+  \caption{Top: two UR3 robots with one gripper each (X=red, Y=green, Z=blue) manipulating a cylinder with two handles.
 The environment contains one rectangular contact surface (in red).
@@
   \begin{center}
-    \begin{tabular}{|p{.23\linewidth}|p{.18\linewidth}|p{.43\linewidth}|}
+    \begin{tabular}{|p{.25\linewidth}|p{.18\linewidth}|p{.41\linewidth}|}
       \hline
@@
     \State{$\state.Pl[\Call{objectIndex}{h}]\gets\false$}
-    \State{$\state$.add(\Call{graspConstraint}{g,h})}
+    \State{$\state$.\ADD(\Call{graspConstraint}{g,h})}
     \EndFor\label{end:comp-placement-constraints}
@@
     \If{$\state.Pl[o]$}
-    \State{$\state$.add(\Call{placeConstraint}{$o$}})
+    \State{$\state$.\ADD(\Call{placeConstraint}{$o$}})
     \EndIf
@@
     \State{$h\gets Gr_1[g]$}
-    \State{$\transition$.add(\Call{graspConstraint}{g,h})}
-    \State{$\transition$.add(\Call{graspConstraintComp}{g,h})}
+    \State{$\transition$.\ADD(\Call{graspConstraint}{g,h})}
+    \State{$\transition$.\ADD(\Call{graspConstraintComp}{g,h})}
     \EndFor
@@
     \If{$\state_1.Pl[o]$}
-    \State{$\transition$.add(\Call{placeConstraint}{$o$}})
-    \State{$\transition$.add(\Call{placeConstraintComp}{$o$}})
+    \State{$\transition$.\ADD(\Call{placeConstraint}{$o$}})
+    \State{$\transition$.\ADD(\Call{placeConstraintComp}{$o$}})
     \EndIf
     \EndFor\label{end:set-placement-constraints}
+    \State{$\transition_1 \gets\mbox{ new transition}(Gr_2,Gr_1)$}
+    \State{$\transition_1$.\SETCONSTRAINTS(\Call{$\transition$.{\scriptsize CONSTRAINTS}})}
     \EndFunction
@@
 of type~(\ref{eq:convex-shape-contact-complement}) and is parameterized: the
-right hand side uniquely defines the contact surfaces (object and robot or environment) and the position of the
+right hand side uniquely defines the contact surfaces and the position of the
 object on the contact surface. $\conf_{target}$ is obtained by projecting $\conf_{rand}$
@@
 clearance parameters are part of the definition of the gripper and handle and
-are stored in the SRDF files.
+are stored in SRDF files.
 
@@
 $\conf_{near}$ (Figure~\ref{fig:MRRT-extension} middle). The edge contains three
-waypoints. The random configuration $\conf_{rand=}$ is displayed in
+waypoints. The random configuration $\conf_{rand}$ is displayed in
 Figure~\ref{fig:MRRT-extension} top. Table~\ref{table:waypoints} shows the
@@
   $\conf_{near}$ &
-  \begin{tabular}{l} \textbf{state} $(\emptyset,\emptyset)$ \\
+  \begin{tabular}{l} \textbf{in state} $(\emptyset,\emptyset)$ \\
   \end{tabular}\\
@@
   \begin{tabular}{l} \textbf{waypoint state} $\gripper{1}>\handle{1}|(\emptyset,\emptyset)|pg$ \\
+    \textbf{in state} $(\emptyset,\emptyset)$ \\
     \textbf{constraints} \\
@@
     \textbf{waypoint state} $\gripper{1}>\handle{1}|(\emptyset,\emptyset)|gp$ \\
-    \textbf{state} \\
-    $(1,\emptyset)$ \\
+    \textbf{in states} $(\emptyset,\emptyset)$ and $(1,\emptyset)$ \\
     \textbf{constraints} \\
@@
   $\conf_{target}$ &
-  \begin{tabular}{l} \textbf{state} $(1,\emptyset)$ \\
+  \begin{tabular}{l}
     \textbf{waypoint state} $\gripper{1}>\handle{1}|(\emptyset,\emptyset)|pp$ \\
+    \textbf{in state} $(1,\emptyset)$ \\
     \textbf{constraints} \\
@@
 \end{tabular}
-\caption{Waypoint configurations computed along edge $(\emptyset,\emptyset)$.
+\caption{Waypoint configurations computed along edge $(\emptyset,\emptyset)\rightarrow(1,\emptyset)$.
   The resulting path between $\conf_{near}$ and $\conf_{target}$ is a concatenation
@@
 \end{table}
-\paragraph{Implementation} From an implementation point of view, class
+\paragraph{Implementation} From an implementation point of view, Class
 \href{https://gepettoweb.laas.fr/hpp/hpp-manipulation/doxygen-html/classhpp_1_1manipulation_1_1graph_1_1WaypointEdge.html}{\texttt{WaypointEdge}} derives from class \href{https://gepettoweb.laas.fr/hpp/hpp-manipulation/doxygen-html/classhpp_1_1manipulation_1_1graph_1_1Edge.html}{\texttt{Edge}}.
 The waypoint configurations are computed by method\\ \texttt{generateTargetConfig} that is specialized in
-class\\ \texttt{WaypointEdge}.
-
-Note that waypoint states are internal to a waypoint edges and are not known by
+Class\\ \texttt{WaypointEdge}.
+
+Note that waypoint states are internal to waypoint edges and are not known by
 the constraint graph when determining in which state a configuration lies
@@
   \end{itemize}
-\item \texttt{hpp-manipulation} a library that implements the manipulation
-  layer with
+\item \texttt{hpp-manipulation} a library that implements manipulation
+  problems and manipulation planning with
   \begin{itemize}
@@
 A virtual docker image can be downloaded to run, test and replicate the
-examples described in the next sections. An archive is provided with this paper
-inflate this archive and follow instructions in the README file.
+examples described in the next sections. An archive is provided with this paper.
+Inflate the archive and follow instructions in the README file.
 
@@
 \href{https://github.com/humanoid-path-planner/hpp_benchmark/tree/v4.10.0/2020-07-23}{\texttt{hpp\_benchmark}} package. We present here only a few test cases.
-The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with 32 Bbytes of RAM and 3 Gbytes of cache memory.
+The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with 32 Megabytes of RAM and 3 Gbytes of cache memory.
 For each test case, we report the minimum, maximum, mean and standard deviation of the time of computation on the one hand, and of the number of nodes in the roadmap built to solve the problem on the other hand.
@@
 The goal configuration is shown on the right.
-The algorithm a constrained RRT close to the one described in~\cite{DalNakLamLau2009}}.
+The algorithm is a constrained RRT close to the one described in~\cite{DalNakLamLau2009}}.
   \label{fig:benchmark-hrp2}
@@
   \end{center}
-  \caption{Manipulation planning problem with Romeo robot manipulating a placard. The robot needs to flip the placard by 180 degrees from an initial pose (left) to a goal pose (right).}
+  \caption{Manipulation planning problem with Romeo robot manipulating a placard.
+The robot needs to flip the placard by 180 degrees from an initial pose (left) to a goal pose (right), keeping balance.}
   \label{fig:romeo-placard}
@@
 
-The third test case features humanoid robot Romeo manipulating a placard. The
+The third test case features Humanoid Robot Romeo manipulating a placard. The
 robot is requested to rotate the placard by 180 degrees. The robot is equipped
@@
 graphs with only the necessary states. We now present a test case that
-illustrates this possibility.
+illustrates this possibility. The system is represented in Figure~\ref{fig:construction-set}.
 
@@
 waypoint transitions as described in Section~\ref{subsec:waypoint}. We
-now empirically show the positive effect of waypoints on the the
+now empirically show the positive effect of waypoints on the
 efficiency of manipulation planning. For that, we run 20 times
@@
 The experimental results show that M-RRT is able to solve a variety of
-manipulation problems including with a legged robots in quasi-static
+manipulation problems including with a legged robot in quasi-static
 equilibrium. No parameter tuning is required between the different problems. All
@@
 For this reason, M-RRT is not able to solve the construction set
-problem in a reasonable amount of time. It remains to our knowledge
+problem in a reasonable amount of time. It remains to our knowledge however
 the only algorithm in the literature able to solve a variety of
@@
 
-This paper presents a software platform aimed at prototyping and solving
-prehensile manipulation planning problems. The platform provide an original
-algorithm M-RRT that is an extension of RRT exploring the leaves of the
-foliation defined by the manipulation constraints. The automatic insertion
-of way point states make the resolution more efficient and the resulting paths
-more natural.
+This paper presents a software platform aimed at prototyping and
+solving a great variety of prehensile manipulation planning
+problems. The platform provides an original algorithm M-RRT that is an
+extension of RRT exploring the leaves of the foliations defined by the
+manipulation constraints. The automatic insertion of waypoint states
+makes the resolution more efficient and the resulting paths more
+natural.
 

### 2020-10-20 hpp.tex
Minor changes.

@@
 It is sometimes useful to produce a configuration $\conf$ that satisfies a constraint (or a set of constraints) of type~(\ref{eq:constraint}) from a configuration $\conf_0$ that does not.
-This action is called the \textit{projection of $\conf_0$ onto the submanifold defined by the constraint} and is performed by a numerical solver that iteratively linearizes the constraint as follows:
+This action is called the \textit{projection of $\conf_0$ onto the submanifold defined by the constraint} and is performed by a Gauss-Newton solver~\cite[Chapter~10]{NoceWrig06} that iteratively linearizes the constraint as follows:
 $$
@@
 $$
-\conf=(\conf_{rob},\conf_{obj})\in\CS,\ \ \ \ \conf_{obj} = g_{grasp}(\conf{rob}).
+\conf=(\conf_{rob},\conf_{obj})\in\CS,\ \ \ \ \conf_{obj} = g_{grasp}(\conf_{rob}).
 $$
@@
 solving this constraint possibly with other constraints using an iterative scheme~(\ref{eq:HierarchicalIterative}) is obviously sub-optimal.
-Notice that the right hand side of~(\ref{eq:explicit-grasp}) is the $\log$ of the pose of the object when grasped, with respect to the actual pose of the object.
 

### 2020-11-05 ijrr.tex
Write conclusion and set doublespace mode.
Small improvements.
Add abstract.

@@
 
-\documentclass[Afour,sageh,times]{sagej}
+\documentclass[Afour,sageh,times,doublespace]{sagej}
 
@@
 \begin{abstract}
+  This paper presents a software platform taylored for prehensile manipulation
+  planning named Humanoid Path Planner. The platform implements an original way
+  of modelling manipulation planning through a constraint graph that represents
+  the numerical constraints that define the manipulation problem. We propose an
+  extension of RRT algorithm to manipulation planning that is able to solve a
+  large variety of problems. We provide replicable experimental results via a
+  docker image that readers may download to run the experimental results by
+  themselves.
 \end{abstract}
@@
 \keywords{robotics, manipulation planning, constrained path planning,
-  path planning}
+  path planning, software platform}
 

### 2021-05-21 hpp.tex
Cite RBPRM.
Add references to Brock and Vendittelli.
Put paper in format IEEE-TRO.

@@
 challenging since it requires more programming effort to integrate
-sensors and motion planning in the process. The work described in this
+sensors and motion planning in the process. A good example of this difficulty is
+the Amazon picking challenge \cite{brock-2018}. The work described in this
 paper is a small step to the simplification of industrial process
@@
 \begin{itemize}
-\item an original and general modelling of prehensile manipulation based on
+\item an original and general modeling of prehensile manipulation based on
   non-linear constraints,
@@
 \item an open-source software suite that implements all the above following state of the art development tools and methods.
-\item a docker image of the above software with installation instructions provided with this paper.
-\end{itemize}
-
-This paper extends the work presented in \cite{mirabel-iros2016} and \cite{MirLam2018} with some new material.
+\item a docker image of the above software with installation instructions provided with this paper. This image makes the experimental results replicable.
+\end{itemize}
+
+This paper extends the work presented in \cite{mirabel-iros2016} and \cite{MirLam2018} with the following new material:
+\begin{itemize}
+\item description of the configuration space as a Cartesian product of Lie groups (Section~\ref{partie:preliminary}),
+\item unified and detailed definition of the grasp and placement constraints that are only evoked in \cite{mirabel-iros2016} (Section~\ref{sec:manipulation-problem}),
+\item automatic construction of the constraint graph (Section~\ref{sec:manipulation-problem}),
+\item the docker image of the software,
+\item a description of the software platform (Section~\ref{sec:humanoid-path-planner}),
+\item experimental results on several different problems.
+\end{itemize}
 
@@
 pushing~\cite{BenRiv1998}, or by throwing~\cite{WooZacLyn2017}, as well as
-multicontact planning \cite{Bretl2006}, \cite{LenVaiYosKhe2013},
+multi-contact planning \cite{Bretl2006}, \cite{LenVaiYosKhe2013},
 \cite{TonDelPetParManMan2018}, in this paper, we are only concerned with
@@
 robot.  The accessible configuration space is thus a union of
-submanifolds as defined in the previous section. Each of these
+sub-manifolds as defined in the previous section. Each of these
 manifolds may moreover be a foliation where each leaf corresponds to a
@@
 gripper. The geometrical structure of the problem has been well
-understood for a long time~\cite{AlaSimLau89}.
+understood for a long time~\cite{AlaSimLau89}. Some specific instances of the
+problem have even been addressed recently \cite{vendittelli-2018}.
 
@@
 final pose. The final configuration of the robot is not specified. A
-simplifying asumption is the existence of a monotone solution, that is
+simplifying assumption is the existence of a monotone solution, that is
 a solution where each object is grasped at most once and is moved from
@@
 
-Other contributions in manipulation planning explicitely address the problem of multi-arm manipulation \cite{GhaCorSim2009}, \cite{HarTsuLau2014}, \cite{DobBek2015}, \cite{XiaLerPha2017}.
+Other contributions in manipulation planning explicitly address the problem of multi-arm manipulation \cite{GhaCorSim2009}, \cite{HarTsuLau2014}, \cite{DobBek2015}, \cite{XiaLerPha2017}.
 
@@
 
-\section{Preliminaries}\label{partie:preliminary}
-\subsection{Kinematic chains and Lie groups}
+\section{Preliminaries: kinematic chains and Lie groups}\label{partie:preliminary}
 
@@
 to each other by joints. Each joint has one degree of freedom either in
-rotation or in translation. A configuration of the kinematic chain is represented by a vector. Each component of the vector represents the angular or linear value of the correspondig joint.
+rotation or in translation. A configuration of the kinematic chain is represented by a vector. Each component of the vector represents the angular or linear value of the corresponding joint.
 
@@
 cannot be correctly represented by translation or rotation joints.
-Representing a freeflying object by three virtual translations followed by
+Representing a free-flying object by three virtual translations followed by
 three virtual rotations called roll, pitch and yaw is indeed a poor workaround
@@
 lock issue that arose during Apollo~13 flight. To avoid singularities, we
-propose the following defintion.
+propose the following definition.
 
@@
   representing $g$, and $(\mathbf{v},\mathbf{\omega})$ is a velocity in $T_{I_3}SE(3)$, the velocity \textit{transported} to $g$ corresponds to linear and angular velocities $R\mathbf{v}$ and $R\omega$ of the moving frame.
-  Integral curves of the vector field mentioned in item 2 above correspond to helicoidal motions of constant velocity expressed in the moving frame.
+  Integral curves of the vector field mentioned in item 2 above correspond to screw motions of constant velocity expressed in the moving frame.
 \end{itemize}
@@
 since the Lie group operator of $\reals$ is $+$ and $\exp_{\reals}$ is the identity.
-In order to homogeneize notation, we define the following operators. For any
+In order to homogenize notation, we define the following operators. For any
 $\conf_0,\conf_1\in\liegroup$ and $\dconf\in T_{\neutral}\liegroup$:
@@
 Given a kinematic chain with joints $(J_1, \cdots, J_{njoints})$, ordered in such
-a way that each joint has an index bigger than its parent in the tree, the configuration space of the robot is the cartesian product of the joint configuration spaces.
+a way that each joint has an index bigger than its parent in the tree, the configuration space of the robot is the Cartesian product of the joint configuration spaces.
 $$
@@
 $$
-This formula generalizes the linear interpolation to robots with freeflying bases, getting rid of singularities of roll -- pitch -- yaw parameterization.
+This formula generalizes the linear interpolation to robots with free-flying bases, getting rid of singularities of roll -- pitch -- yaw parameterization.
 %% begin impl
@@
 In this \paper, we will make use of the term Jacobian in a generalized way.
-If $h$ is a differentiable function from a Lie group $\liegroup_1$ to a Lie group $\liegroup_2$, and $\conf_1$ an element of $\liegroup_1$, we will denote by $\frac{\partial h}{\partial\conf}(\conf_1)$ the operator that maps velocities in $T_{\conf_1}\liegroup_1$ to the velocity in $T_{h(\conf_1)}\liegroup_2$ \textit{transported by} $h$\footnote{If $\dconf\in T_{\conf_1}\liegroup_1$ is a velocity along a time parameteriezd curve $\gamma$, $\frac{\partial h}{\partial\conf}(\conf_1)\dconf$ is the velocity at
+If $h$ is a differentiable function from a Lie group $\liegroup_1$ to a Lie group $\liegroup_2$, and $\conf_1$ an element of $\liegroup_1$, we will denote by $\frac{\partial h}{\partial\conf}(\conf_1)$ the operator that maps velocities in $T_{\conf_1}\liegroup_1$ to the velocity in $T_{h(\conf_1)}\liegroup_2$ \textit{transported by} $h$\footnote{If $\dconf\in T_{\conf_1}\liegroup_1$ is a velocity along a time parameterized curve $\gamma$, $\frac{\partial h}{\partial\conf}(\conf_1)\dconf$ is the velocity at
   the same time along Curve $h(\gamma)$.}.
@@
 It is sometimes useful to produce a configuration $\conf$ that satisfies a constraint (or a set of constraints) of type~(\ref{eq:constraint}) from a configuration $\conf_0$ that does not.
-This action is called the \textit{projection of $\conf_0$ onto the submanifold defined by the constraint} and is performed by a Gauss-Newton solver~\cite[Chapter~10]{NoceWrig06} that iteratively linearizes the constraint as follows:
+This action is called the \textit{projection of $\conf_0$ onto the sub-manifold defined by the constraint} and is performed by a Gauss-Newton solver~\cite[Chapter~10]{NoceWrig06} that iteratively linearizes the constraint as follows:
 $$
@@
 
-In manipulation planning applications, where robots manipulate objects, once an object is grasped, the position of the object can be computed explicitely from the configuration of the robot.
+In manipulation planning applications, where robots manipulate objects, once an object is grasped, the position of the object can be computed explicitly from the configuration of the robot.
 In this case, some configuration variables of the system depend on other:
@@
 
-Now that we are able to project configurations onto submanifolds defined by numerical constraints -- up to some numerical threshold, we need now to define paths on such submanifolds.
+Now that we are able to project configurations onto sub-manifolds defined by numerical constraints -- up to some numerical threshold, we need now to define paths on such sub-manifolds.
 The usual way of doing so is to discretize the path and to project each sample configuration.
@@
 Instead, we propose an alternative architecture where paths store the constraints they are subject to and apply the constraints at path evaluation.
-Let $P\in C^1([0,T],\CS)$ be a path without constraint defined on an interval $[0,T]$, and $\mathbf{proj}$ a projector onto a submanifold defined by numerical constraints (\textit{i.e.} an instance of\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}).
+Let $P\in C^1([0,T],\CS)$ be a path without constraint defined on an interval $[0,T]$, and $\mathbf{proj}$ a projector onto a sub-manifold defined by numerical constraints (\textit{i.e.} an instance of\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}).
 
@@
 
-The previous sections have presented how we model kinematic chains, configurations and velocities for a given robotic system, how configurations and paths can be projected onto a submanifold of the configuration space defined by numerical constraints.
+The previous sections have presented how we model kinematic chains, configurations and velocities for a given robotic system, how configurations and paths can be projected onto a sub-manifold of the configuration space defined by numerical constraints.
 
@@
 \item if $flags=(\true,\true,\true,\true,\true,\false)$ the grasp is satisfied iff the centers and $z$ axes of $\gripper(\conf)$ and $\handle(\conf)$ coincide (free rotation around $z$).
-This is useful for cylindical objects:
+This is useful for cylindrical objects:
   \begin{align*}
@@
 \end{itemize}
-If $\conf_0$ is a configuration satisfying the grasp constraint: $h(\conf_0)$=0, then the submanifold defined by
+If $\conf_0$ is a configuration satisfying the grasp constraint: $h(\conf_0)$=0, then the sub-manifold defined by
 $$
@@
 \begin{enumerate}
-\item defining virtual grippers in the environment and virtual handles on the object, implicitely defines a discrete set of poses,
+\item defining virtual grippers in the environment and virtual handles on the object, implicitly defines a discrete set of poses,
 \item defining a virtual gripper on an horizontal plane and a virtual handle on the object, and using a grasp with flags $(\false,\false,\true,\true,\true,\false)$ constrains the object to move on a infinite horizontal plane.
@@
 \end{equation}
-The contact constraint is defined by the following piecewise differentiable function:
+The contact constraint is defined by the following piece-wise differentiable function:
 \begin{equation}\label{eq:convex-shape-contact}
@@
 
-As for grasps, we need to define a parameterized complement constraint for the contact constraint in order to specify the submanifold of configurations that are reachable from one configuration while keeping the object in a constant stable pose.
+As for grasps, we need to define a parameterized complement constraint for the contact constraint in order to specify the sub-manifold of configurations that are reachable from one configuration while keeping the object in a constant stable pose.
 The naive way consists in defining
@@
 Let $\conf_0$ be a configuration such that $h(\conf_0)=0$.
-The submanifold defined by
+The sub-manifold defined by
 \begin{equation}\label{eq:stable-pose}
@@
 \end{itemize}
-With this definition, the submanifold defined by~(\ref{eq:stable-pose}), (\ref{eq:convex-shape-contact}), (\ref{eq:convex-shape-contact-complement}) contains configurations where the object is in a unique stable pose.
+With this definition, the sub-manifold defined by~(\ref{eq:stable-pose}), (\ref{eq:convex-shape-contact}), (\ref{eq:convex-shape-contact-complement}) contains configurations where the object is in a unique stable pose.
 The polygon indices $i$ and $j$, as well as their relative position can indeed be recovered from~(\ref{eq:convex-shape-contact-complement}):
@@
 
-\paragraph{Number of states:} note that for $i\in\{1,\cdots,\nhandles\}$ the number of occurences of $i$ in $\state$ is at most 1: a handle cannot be grasped by several grippers.
-Note also that the number of occurences of $\emptyset$ is not limited: several grippers may hold nothing.
+\paragraph{Number of states:} note that for $i\in\{1,\cdots,\nhandles\}$ the number of occurrences of $i$ in $\state$ is at most 1: a handle cannot be grasped by several grippers.
+Note also that the number of occurrences of $\emptyset$ is not limited: several grippers may hold nothing.
 Let $m$ be a non-negative integer not greater than $\ngrippers$ nor $\nhandles$ and let us count the number of states with $m$ handles grasped.
@@
     \EndFunction
+  \end{algorithmic}
+  \caption{Recursive Construction of the constraint graph. The
+    construction starts by the state with no grasp. Call to
+    {\scriptsize RECURSE} function loops over the available grippers
+    and handles and creates states with one more grasp, and a
+    transition to these new states. In each state, a placement
+    constraint is added for each object of which no handle is
+    grasped. Variables $\mathcal{G}$ and $\mathcal{H}$ contain the
+    indices of the free grippers and handles. Variable $Gr$ stores the
+    current set of grasps following Expression~(\ref{eq:state}).
+    Lines~\ref{beg:comp-placement-constraints} to
+    \ref{end:comp-placement-constraints}
+    compute which objects are not grasped.
+    Lines~\ref{beg:set-placement-constraints} to
+    \ref{end:set-placement-constraints} insert placement constraints in the
+    state for those objects. Line~\ref{cond:recurse} recurses only if the
+    latest node reached is new. Functions \CREATESTATE and \CREATETRANSITION are given in Algorithm~\ref{algo:createStateAndTransition}.}
+  \label{algo:constraintGraphFactory}
+\end{algorithm}
+\begin{algorithm}
+  \begin{algorithmic}[1]
     \Function{createState}{$Gr$}
@@
   \end{algorithmic}
-  \caption{Recursive Construction of the constraint graph. The
-    construction starts by the state with no grasp. Call to
-    {\scriptsize RECURSE} function loops over the available grippers
-    and handles and creates states with one more grasp, and a
-    transition to these new states. In each state, a placement
-    constraint is added for each object of which no handle is
-    grasped. Variables $\mathcal{G}$ and $\mathcal{H}$ contain the
-    indices of the free grippers and handles. Variable $Gr$ stores the
-    current set of grasps following Expression~(\ref{eq:state}).
-    Lines~\ref{beg:comp-placement-constraints} to
-    \ref{end:comp-placement-constraints}
-    compute which objects are not grasped.
-    Lines~\ref{beg:set-placement-constraints} to
-    \ref{end:set-placement-constraints} insert placement constraints in the
-    state for those objects. Line~\ref{cond:recurse} recurses only if the
-    lastest node reached is new.}
-  \label{algo:constraintGraphFactory}
+  \caption{Method \CREATESTATE builds the constraints relative to a state: one grasp constraint for each grasp, and one placement constraint for each object non grasped.
+\CREATETRANSITION builds the constraints relative to a transition: the constraints of the initial state (with the fewest grasps) and their complements.}
+  \label{algo:createStateAndTransition}
 \end{algorithm}
@@
 If so, it checks that the right hand side of the transition parameterized constraints is the same for both configurations (up to the error threshold).
-Then it returns the linear interpolation between the configurations, projected onto the submanifold defined by the transition constraints.
+Then it returns the linear interpolation between the configurations, projected onto the sub-manifold defined by the transition constraints.
 If the path is in collision, only the collision-free part at the beginning of the path is returned.
@@
     \EndFunction
+  \end{algorithmic}
+  \caption{Manipulation RRT algorithm iteratively calls method {\scriptsize ONE}{\small S}{\scriptsize TEP} until a solution path is found or the maximum number of iterations is reached.
+Function \CONNECT is described in Algorithm~\ref{algo:M-RRT-connect}}
+  \label{algo:M-RRT}
+\end{algorithm}
+\begin{algorithm}
+  \begin{algorithmic}
     \Function{connect}{$\conf_1,s_1,\conf_2,s_2$}
@@
   \end{algorithmic}
-  \caption{Manipulation RRT algorithm iteratively calls method {\scriptsize ONE}{\small S}{\scriptsize TEP} until a solution path is found or the maximum number of iterations is reached.}
-  \label{algo:M-RRT}
+  \caption{Function \CONNECT of M-RRT algorithm}
+  \label{algo:M-RRT-connect}
 \end{algorithm}
@@
 Method \CONNECT checks that the right hand side of $grasp_{11}/comp$ is the same
-for $\conf_1$ and $\conf_2$, up to the error threhold (Algorithm~\ref{algo:M-RRT}, line~51).
+for $\conf_1$ and $\conf_2$, up to the error threshold (Algorithm~\ref{algo:M-RRT}, line~51).
 From a geometrical point of view, this means that the orientation of the cylinder along its axis, with respect to the gripper is the same in both configurations.
@@
     \textit{pregrasp} (pg) where the gripper is above the object,
-    \textit{graps-placement} (gp) where the object is grasped but
+    \textit{grasp-placement} (gp) where the object is grasped but
     still in placement and \textit{preplacement} (pp) where the object
@@
   and solvers,
-\item \texttt{hpp-core} a library that implements most of the conceps relative
+\item \texttt{hpp-core} a library that implements most of the concept relative
   to motion planning. The main features are
@@
   \item abstraction of paths in configuration spaces and some implementations,
-  \item abstaction of path planning and path optimzation and some implementations,
+  \item abstraction of path planning and path optimization and some implementations,
   \item abstraction of steering methods and some implementations,
@@
 The first test case features robot Baxter manipulating two boxes on a table (see Figure~\ref{fig:baxter-manipulation-boxes}).
-The boxes are swaped between the initial and final configurations.
+The boxes are swapped between the initial and final configurations.
 The robot has two grippers and each box is equipped with a handle.
@@
 The second test case features robot PR-2 manipulating a box on a table.
-The robot is resquested to flip the box upside down from an initial pose to a goal pose as represented in Figure~\ref{fig:pr2-manipulation-two-hands}.
+The robot is requested to flip the box upside down from an initial pose to a goal pose as represented in Figure~\ref{fig:pr2-manipulation-two-hands}.
 The robot is equipped with two grippers, the box is equipped with two handles.
@@
 with two grippers and the placard with two handles. Each handle is associated
-to a unique gripper. The number of states of the contraint graph is thus 3.
+to a unique gripper. The number of states of the constraint graph is thus 3.
 
@@
 computed and only those states (9 in total) are built in the
-constraint graph. Then, a sequence of subgoals in the successive
+constraint graph. Then, a sequence of sub-goals in the successive
 states is computed, in such a way that each sub-goal is accessible by
 the previous one (on the same leaf of the corresponding transition
-foliation). The subgoals are then linked by running a constrained visibility
+foliation). The sub-goals are then linked by running a constrained visibility
 PRM algorithm (\cite{SimLauNis00}) on each leaf. The python code can be found
@@
 We think that this platform is a perfect basis for researchers who want to
-develop and benchmark new manpulation planning algorithms.
+develop and benchmark new manipulation planning algorithms. Note that some
+on-going work in humanoid locomotion  \cite{tonneau-2018} is based on HPP.
 

### 2021-05-21 hpp.tex
Fix equations of ConvexShapeContact.

@@
 \begin{equation}\label{eq:convex-shape-contact}
-h(\conf) = \left\{\begin{array}{lll}(\bar{h}_1,0,0,\bar{h}_4,\bar{h}_5)&
+h(\conf) = \left\{\begin{array}{lll}(\bar{h}_1,0,0,\bar{h}_5,\bar{h}_6)&
 \mbox{if} & P(C_{o_i},f_j)\mbox{ inside } f_j\\
-(\bar{h}_1,\bar{h}_2,\bar{h}_3,\bar{h}_4,\bar{h}_5)&
+(\bar{h}_1,\bar{h}_2,\bar{h}_3,\bar{h}_5,\bar{h}_6)&
 \mbox{if} & P(C_{o_i},f_j)\mbox{ outside } f_j\end{array}\right.
@@
 $$
-h_{comp}(\conf) = \left\{\begin{array}{lll}(\bar{h}_2,\bar{h}_3,\bar{h}_6)&
+h_{comp}(\conf) = \left\{\begin{array}{lll}(\bar{h}_2,\bar{h}_3,\bar{h}_4)&
 \mbox{if} & P(C_{o_i},f_j)\mbox{ inside } f_j\\
-(0,0,\bar{h}_6)&
+(0,0,\bar{h}_4)&
 \mbox{if} & P(C_{o_i},f_j)\mbox{ outside } f_j\end{array}\right.
 $$
-$\bar{h}_2$, $\bar{h}_3$, $\bar{h}_6$ respectively represent the translation in $x-y$ plane and the rotation around $z$-axis of frame $o_i$ with respect to frame $f_j$.
+$\bar{h}_2$, $\bar{h}_3$, $\bar{h}_4$ respectively represent the translation in $y-z$ plane and the rotation around $x$-axis of frame $o_i$ with respect to frame $f_j$.
 Let $\conf_0$ be a configuration such that $h(\conf_0)=0$.
@@
 
-To disambiguate the various combination of convex polygons that can be in contact, we define
+To disambiguate the various combinations of convex polygons that can be in contact, we define
 \begin{equation}\label{eq:convex-shape-contact-complement}
-h_{comp}(\conf) = \left\{\begin{array}{l}(\bar{h}_2 + 2jM,\bar{h}_3 + 2iM,\bar{h}_6)\\
+h_{comp}(\conf) = \left\{\begin{array}{l}(\bar{h}_2 + 2jM,\bar{h}_3 + 2iM,\bar{h}_4)\\
 \mbox{if } P(C_{o_i},f_j)\mbox{ inside } f_j\\
-(2jM,2iM,\bar{h}_6)\\
+(2jM,2iM,\bar{h}_4)\\
 \mbox{if } P(C_{o_i},f_j)\mbox{ outside } f_j\end{array}\right.
@@
   j &= \left\lfloor\frac{\bar{h}_2}{2M} + \frac{1}{2}\right\rfloor\\
-  \bar{h}_1 &= \bar{h}_4 = \bar{h}_5 = 0\\
+  \bar{h}_1 &= \bar{h}_5 = \bar{h}_6 = 0\\
   \bar{h}_2 &= h_{comp\ 1} - 2jM\\
   \bar{h}_3 &= h_{comp\ 2} - 2iM\\
-  \bar{h}_6 &= h_{comp\ 3}
+  \bar{h}_4 &= h_{comp\ 3}
 \end{align*}

### 2021-09-01 hpp.tex
Take into account reviewers comments.

@@
 \section{Introduction}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{figures/example.png}
+    \def\svgwidth{\columnwidth}
+    \graphicspath{{./figures/}}
+    \input{figures/example-graph.pdf_tex}
+  \end{center}
+  \caption{{\color{blue} Example of manipulation planning problem.} Top: two UR3 robots with one gripper each (X=red, Y=green, Z=blue) manipulating a cylinder with two handles.
+The environment contains one rectangular contact surface (in red).
+The cylinder has two rectangular contact surfaces (in green).
+  Bottom: the corresponding constraint graph.
+Names of states follow Expression~(\ref{eq:state}): for instance, $(\emptyset,1)$ means that gripper of robot 2 grasps handle 1 of the cylinder.
+In this state, there is no placement constraint.}
+  \label{fig:2ur3-cylinder}
+\end{figure}
 
@@
 
-This \paper therefore describes a software platform called \href{https://humanoid-path-planner.github.io/hpp-doc}{Humanoid Path
+This \paper therefore describes a software platform called {Humanoid Path
   Planner} taylored for manipulation planning in robotics. It can handle many types of robots, from manipulator arms to legged humanoid robots.
@@
 \item a manipulation planning algorithm that is able to tackle a great variety of manipulation planning problems,
-\item an open-source software suite that implements all the above following state of the art development tools and methods.
+\item an open-source software suite that implements all the above following state-of-the-art development tools and methods.
 \item a docker image of the above software with installation instructions provided with this paper. This image makes the experimental results replicable.
 \end{itemize}
-
-This paper extends the work presented in \cite{mirabel-iros2016} and \cite{MirLam2018} with the following new material:
+{\color{blue} Installation instructions can be found by following this \href{https://humanoid-path-planner.github.io/hpp-doc}{link}.}
+This paper extends the work presented in {\color{blue} previous papers }\cite{mirabel-iros2016},\cite{MirLam2018} with the following new material:
 \begin{itemize}
 \item description of the configuration space as a Cartesian product of Lie groups (Section~\ref{partie:preliminary}),
-\item unified and detailed definition of the grasp and placement constraints that are only evoked in \cite{mirabel-iros2016} (Section~\ref{sec:manipulation-problem}),
+\item unified and detailed definition of the grasp and placement constraints that are only evoked in {\color{blue}Mirabel \textit{et al}}~\cite{mirabel-iros2016} (Section~\ref{sec:manipulation-problem}),
 \item automatic construction of the constraint graph (Section~\ref{sec:manipulation-problem}),
@@
 \Partie~\ref{partie:preliminary} introduces some preliminary notions like kinematic chains and Lie groups that are used to model the configuration space of each joint.
-\Partie~\ref{sec:constraints} introduces non linear constraints and solvers that are at the core of manipulation problem definition.
+\Partie~\ref{sec:constraints} introduces non-linear constraints and solvers that are at the core of manipulation problem definition.
 \Partie~\ref{sec:manipulation-problem} defines the problem of prehensile manipulation in the general setting.
@@
 Some systems are subject to non-linear constraints. These non-linear
-constraints define a sub-manifold of the configuration space the robot must
-stay on. For instance, legged robots that must keep contact with the
-ground and enforce quasi-static equilibrium, or multi-arm systems
+constraints define sub-manifolds of the configuration space the robot
+must stay on. For instance, legged robots that must keep contact with
+the ground and enforce quasi-static equilibrium, or multi-arm systems
 grasping a same object are subject to this type of constraints. As the
@@
 probability. To sample configurations on the constrained manifold,
-\cite{DalNakLamLau2009} and \cite{berenson2011} project random
-configurations using a generalization of Newton-Raphson
-algorithm. Another way to sample configurations on a constrained
-manifold consists in expressing some configuration variables with
-respect to others~\cite{Cortes2002}, \cite{MirLam2018} when this is
-possible. \cite{atlasRRT} propose another method based on non-linear
-projection. They cover the constrained manifold by growing an atlas
-composed of local charts. This approximation provides a probability
-distribution that is closer to the uniform distribution over the
-manifold than the projection of a uniform distribution over the
-configuration space. \cite{tb-rrt2016} propose a variation of the latter paper.
-The main difference resides in the fact that the nodes
-built on the tangent space are not immediately projected on the manifold.
-\cite{CefOri2019} propose a general framework to plan task constrained motions
-in the presence of moving obstacle.
-\cite{KinMolKav2019} provide an in-depth review of the various approaches to
-motion planning with non-linear constraints.
+{\color{blue}Dalibard \textit{et al}}~\cite{DalNakLamLau2009} and {\color{blue}Benrenson
+\textit{et al}}~\cite{berenson2011} project random configurations using
+a generalization of Newton-Raphson algorithm. Another way to sample
+configurations on a constrained manifold consists in expressing some
+configuration variables with respect to others~\cite{Cortes2002},
+\cite{MirLam2018} when this is possible. {\color{blue}Jaillet \textit{et al}}~\cite{atlasRRT} propose
+another method based on non-linear projection. They cover the
+constrained manifold by growing an atlas composed of local
+charts. This approximation provides a probability distribution that is
+closer to the uniform distribution over the manifold than the
+projection of a uniform distribution over the configuration
+space. {\color{blue}Beobkyoon \textit{et al}}~\cite{tb-rrt2016} propose a variation of the latter paper.  The
+main difference resides in the fact that the nodes built on the
+tangent space are not immediately projected on the manifold.
+{\color{blue}Cefalo \textit{et al}}~\cite{CefOri2019} propose a general framework to plan task constrained
+motions in the presence of moving obstacle. {\color{blue}Kingston \textit{et al}}~\cite{KinMolKav2019}
+provide an in-depth review of the various approaches to motion
+planning with non-linear constraints.
 
@@
 some objects are moved by some robots.
-Although several instances of
+Although several instances of the
 manipulation problem exist like manipulation by
@@
 manifolds may moreover be a foliation where each leaf corresponds to a
-stable position of an object or to a grasp of an object by a
+stable pose of an object or to a grasp of an object by a
 gripper. The geometrical structure of the problem has been well
@@
 The first attempt to solve manipulation planning problems using random sampling
-was proposed by \cite{simeon2004ijrr} where a reduction property reduces the
+was proposed by {\color{blue}Siméon \textit{et al}}~\cite{simeon2004ijrr} where a reduction property reduces the
 complexity of the problem.
@@
 a path for a robot that needs to move objects in order to reach a goal
-configuration. The final position of the objects does not matter in this case.
+configuration. The final poses of the objects do not matter in this case.
 
@@
 composed of a symbolic or task planner and of a motion
-planner~\cite{cambon:hal-01976081}, \cite{KaeLoz2013}, \cite{18-toussaint-RSS}.
+planner~\cite{cambon:hal-01976081}, \cite{KaeLoz2013}, \cite{LozKae2014}, \cite{18-toussaint-RSS}.
 
@@
 
-\cite{schmitt19icra} propose an approach where two robots manipulate an object in a dynamic environment. The output of the algorithm is a sequence of controllers rather than a sequence of paths.
-
-Our work shares many ideas with \cite{HauNgt2011} where the notion of constraint
+{\color{blue}Schmitt \textit{et al}}~\cite{schmitt19icra} propose an approach where two robots manipulate an object in a dynamic environment.
+The output of the algorithm is a sequence of controllers rather than a sequence of paths.
+
+Our work shares many ideas with {\color{blue}Hauser and Ng-Thow-Hing} \cite{HauNgt2011} where the notion of constraint
 graph is present, although not as clearly expressed as in this paper. The main
@@
 OMPL integrates many randomized path planning algorithms and is widely used for teaching.
-Recently, \cite{KinMolKav2019} proposed an extension for systems subject to non-linear constraints.
+Recently, {\color{blue}Kingston \textit{et al}}~\cite{KinMolKav2019} proposed an extension for systems subject to non-linear constraints.
 
@@
 The group operation is $+$ for $\mathbb{R}^n$, and composition denoted as $"."$ for $SE(n)$.
-We refer to~\cite{LMS94} -- Appendix A for a thorough definition of Lie groups.
+We refer to~{\color{blue}Murray \textit{et al}}~\cite[Appendix A]{LMS94} for a thorough definition of Lie groups.
 We expose here only properties that are useful for the following developments.
@@
   \end{tabular}
-  \caption{Main Lie group types and their vector representations. Notice that the dimensions of the configuration representation and of the velocity representation may differ.}
+  \caption{Main Lie group types and their vector representations.
+Notice that the dimensions of the configuration representation and of the velocity representation may differ.
+{\color{blue} Using $(\cos\theta,\sin\theta)$ instead of $\theta$ for $SO(2)$ and $SE(2)$ makes the parameterization continuous when $\theta$ discontinuously switches from $-\pi$ to $\pi$}}
   \label{tab:vector representation}
@@
 $$
-In some cases, we may specify in subscript the Lie group that is used: \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/explog.hpp#L34}{$\exp_{SO(3)}$}, \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/explog.hpp#L236}{$\exp_{SE(3)}$}.
+In some cases, we may specify in subscript the Lie group that is used: \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/explog.hpp\#L34}{$\exp_{SO(3)}$}, \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/explog.hpp\#L236}{$\exp_{SE(3)}$}.
 
@@
 Although $\exp$ is not injective, choosing the smallest norm $\mathbf{v}$ uniquely defines function $\log$ from $\liegroup$ to $T_{\neutral}\liegroup$, up to some singularities where several candidates $\mathbf{v}$ are of equal norms.
-Again, we may specify the Lie group that is used: \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/log.hxx#L112}{$\log_{SE(3)}$}, \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/log.hxx#L15}{$\log_{SO(3)}$}
+Again, we may specify the Lie group that is used: \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/log.hxx\#L112}{$\log_{SE(3)}$}, \href{https://github.com/stack-of-tasks/pinocchio/blob/f8f3b9a24eab527df79650e3dc73410f9a46a2b2/src/spatial/log.hxx\#L15}{$\log_{SO(3)}$}
 
@@
 $$
-We denote by $iq_i$, and $iv_i$ the indices of joint $i$ in the robot configuration and velocity vectors.
+We denote by $iq_i$, and $iv_i$ the {\color{blue}starting} indices of joint $i$ in the robot configuration and velocity vectors.
 $$
@@
 $$
-\conf(t) = \conf_0 \oplus t (\conf_1 - \conf_0)
+\conf(t) = \conf_0 \oplus t (\conf_1 {\color{blue}\ominus} \conf_0)
 $$
@@
 
-\begin{definition}\textbf{Non-linear constraint.} A non linear constraint is defined by a differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and is written
+\begin{definition}\textbf{Non-linear constraint.} A non-linear constraint is defined by a {\color{blue} piece-wise} differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and is written
 \begin{equation}\label{eq:constraint}
@@
 \begin{definition}\textbf{Parameterized non-linear constraint.}
-  A parameterized non-linear constraint is defined by a differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and by a vector $\mathbf{h}_0$ of $\reals^m$ and is written
+  A parameterized non-linear constraint is defined by a {\color{blue} piece-wise} differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and by a vector $\mathbf{h}_0$ of $\reals^m$ and is written
   $$
@@
 %% begin impl
-Differentiable mappings are represented by abstract Class\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1DifferentiableFunction.html}{\texttt{DifferentiableFunction}}.
+{\color{blue} Piece-wise} differentiable mappings are represented by abstract Class\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1DifferentiableFunction.html}{\texttt{DifferentiableFunction}}.
 %% end impl
@@
 In this \paper, we will make use of the term Jacobian in a generalized way.
-If $h$ is a differentiable function from a Lie group $\liegroup_1$ to a Lie group $\liegroup_2$, and $\conf_1$ an element of $\liegroup_1$, we will denote by $\frac{\partial h}{\partial\conf}(\conf_1)$ the operator that maps velocities in $T_{\conf_1}\liegroup_1$ to the velocity in $T_{h(\conf_1)}\liegroup_2$ \textit{transported by} $h$\footnote{If $\dconf\in T_{\conf_1}\liegroup_1$ is a velocity along a time parameterized curve $\gamma$, $\frac{\partial h}{\partial\conf}(\conf_1)\dconf$ is the velocity at
+If $h$ is a {\color{blue} piece-wise} differentiable function from a Lie group $\liegroup_1$ to a Lie group $\liegroup_2$, and $\conf_1$ an element of $\liegroup_1$, we will denote by $\frac{\partial h}{\partial\conf}(\conf_1)$ the operator that maps velocities in $T_{\conf_1}\liegroup_1$ to the velocity in $T_{h(\conf_1)}\liegroup_2$ \textit{transported by} $h$\footnote{If $\dconf\in T_{\conf_1}\liegroup_1$ is a velocity along a time parameterized curve $\gamma$, $\frac{\partial h}{\partial\conf}(\conf_1)\dconf$ is the velocity at
   the same time along Curve $h(\gamma)$.}.
@@
 $$
-h(\conf_{i+1}) \approx h(\conf_i) + \frac{\partial h}{\partial \conf}(\conf_{i})(\conf_{i+1}-\conf_i) = 0
+h(\conf_{i+1}) \approx h(\conf_i) + \frac{\partial h}{\partial \conf}(\conf_{i})(\conf_{i+1}{\color{blue}\ominus}\conf_i) = 0
 $$
@@
 \begin{equation}\label{eq:HierarchicalIterative}
-\conf_{i+1} = \conf_i-\alpha_i\frac{\partial h}{\partial \conf}^{+}(\conf_{i}) h(\conf_i)
+\conf_{i+1} = \conf_i{\color{blue}\ominus}\alpha_i\frac{\partial h}{\partial \conf}^{+}(\conf_{i}) h(\conf_i)
 \end{equation}
@@
 Class\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1HierarchicalIterative.html}{\texttt{HierarchicalIterative}} implements the above Newton method.
-Several linesearch methods are implemented: \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1Backtracking.html}{\texttt{Backtracking}}, \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1ErrorNormBased.html}{\texttt{ErrorNormBased}}, \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1FixedSequence.html}{\texttt{FixedSequence}}, and \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1Constant.html}{\texttt{Constant}}.
+{\color{blue}
+Several line search methods are implemented:
+\begin{itemize}
+\item \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1Backtracking.html}{\texttt{Backtracking}} \cite{backtracking},
+\item \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1ErrorNormBased.html}{\texttt{ErrorNormBased}}:
+  $$\alpha_{i} = C - K  \text{tanh}(a \frac{\|f(\mathbf{q}_i)\|}{\epsilon^2} + b),$$
+  where $C$, $K$, $a$, and $b$ are constant values, and $\epsilon$ is the
+  error threshold,
+\item \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1FixedSequence.html}{\texttt{FixedSequence}} implements a fixed sequence of $\alpha_i$ that converges to 1,
+\item and \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/structhpp_1_1constraints_1_1solver_1_1lineSearch_1_1Constant.html}{\texttt{Constant}} sets $\alpha_i$ to 1.
+\end{itemize}
+}
 Note that to define a new constraint, the user needs to derive class \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1DifferentiableFunction.html}{\texttt{DifferentiableFunction}} and to implement methods \texttt{impl\_compute} and \texttt{impl\_jacobian}.
@@
 To optimize constraint resolution, we perform variable substitution when possible in order to reduce both the number of variables and the dimension of the resulting implicit constraint.
-We describe here the method that has been first published in~\cite{MirLam2018}.
+We describe here the method that has been first published in~{\color{blue}Mirabel \textit{et al}}~\cite{MirLam2018}.
 Unlike in the former paper, the description we give in Algorithm~\ref{alg:ExplicitConstraintSet} is closer to the real implementation.
@@
 Other coefficients of $J_{i}J$ are equal to the corresponding coefficients of $J$.
-An implementation of the above Jacobian product can be found \href{https://github.com/humanoid-path-planner/hpp-constraints/blob/e21490c8c713949bd3038dccb6fe02cf254a615f/src/explicit-constraint-set.cc#L267}{here}.
+An implementation of the above Jacobian product can be found \href{https://github.com/humanoid-path-planner/hpp-constraints/blob/e21490c8c713949bd3038dccb6fe02cf254a615f/src/explicit-constraint-set.cc\#L267}{here}.
 
 % begin impl
-The solver by substitution described in this section is implemented by Class \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}, that store an instance of \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1ExplicitConstraintSet.html}{\texttt{ExplicitConstraintSet}}.
+The solver by substitution described in this section is implemented by Class \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}, that stores an instance of \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1ExplicitConstraintSet.html}{\texttt{ExplicitConstraintSet}}.
 
@@
 The usual way of doing so is to discretize the path and to project each sample configuration.
-The shortcoming of this method is that is requires to choose a discretization step at path construction and to lose the continuous information of the path.
-
-Instead, we propose an alternative architecture where paths store the constraints they are subject to and apply the constraints at path evaluation.
+The shortcoming of this method is that it requires to choose a discretization step at path construction and to lose the continuous information of the path.
+
+Instead, we propose an alternative architecture where paths store the constraints they are subject to and apply the constraints at path evaluation {\color{blue}(\textit{i.e.} when computing the configuration at a given parameter)}.
 Let $P\in C^1([0,T],\CS)$ be a path without constraint defined on an interval $[0,T]$, and $\mathbf{proj}$ a projector onto a sub-manifold defined by numerical constraints (\textit{i.e.} an instance of\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}).
@@
 Before inserting a projected path in a roadmap for instance, it is therefore necessary to detect possible discontinuities.
-\cite{Hauser-RSS-13} proposes a solution to this problem.
-\cite{mirabel:hal-01360409} describe two algorithms to check whether a projected path is continuous.
+{\color{blue}Hauser} \cite{Hauser-RSS-13} proposes a solution to this problem.
+{\color{blue}In a previous paper} \cite{mirabel:hal-01360409}, {\color{blue}we} described two algorithms to check whether a projected path is continuous.
 These algorithms are implemented by
 classes \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1pathProjector_1_1Dichotomy.html}{\texttt{pathProjector::Dichotomy}} and \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1pathProjector_1_1Progressive.html}{\texttt{pathProjector::Progressive}}.
+{\color{blue} Note that when a path is not continuous, the algorithms returns a continuous portion of the path starting at the beginning of the path.
+This enables function \EXTEND in Algorithm~\ref{algo:M-RRT} to create a new node.}
 
@@
   \begin{itemize}
-    \item each object is either grasped by a robot, or lies in a stable contact pose.
+  \item each object is either grasped by a robot, or lies in a stable contact pose,
+  \item {\color{blue} the volumes occupied by the links of the robots and by
+    the objects are pair-wise disjoint.}
   \end{itemize}
@@
 $$
-\CS = \CS_{r_1}\times\CS_{r_{nr}}\times SE(3)^{no}
+\CS = \CS_{r_1}{\color{blue}\times\cdots}\times\CS_{r_{nr}}\times SE(3)^{no}
 $$
@@
 \end{definition}
+    {\color{blue}Note that $\reals^3\times SO(3)$ and $SE(3)$ have
+      different group operators, exponential maps and
+      logarithms. Constant velocity motions in $SE(3)$ are screw
+      motions while constant velocity motions in $\reals^3\times
+      SO(3)$ consist of linear interpolation of the center of the
+      frame and constant angular velocity.}
 \begin{definition}\textbf{Grasp complement.}
@@
   \end{align*}
+\item {\color{blue}if $flags=(\true,\true,\false,\true,\true,\false)$
+  the grasp is satisfied iff the center of $\handle(\conf)$ is on the
+  $z$ axis of $\gripper(\conf)$ and if the $z$ axes of
+  $\gripper(\conf)$ and $\handle(\conf)$ coincide (free translation
+  and rotation around $z$). This is also useful for cylindrical
+  objects:
+  \begin{align*}
+    h &= (\bar{h}_1,\bar{h}_2,\bar{h}_4,\bar{h}_5),\\
+    h_{comp} &= (,\bar{h}_3,\bar{h}_6),
+  \end{align*}
+  however inequality constraints need to be added manually on $\bar{h}_3$ to
+  limit the translation.
+  }
 \item if $flags=(\true,\true,\true,\false,\false,$ $\false)$ the grasp is satisfied iff the centers of $\gripper(\conf)$ and $\handle(\conf)$ coincide (free rotation).
@@
 
-\paragraph{Number of states:} note that for $i\in\{1,\cdots,\nhandles\}$ the number of occurrences of $i$ in $\state$ is at most 1: a handle cannot be grasped by several grippers.
+\paragraph{Number of states} note that for $i\in\{1,\cdots,\nhandles\}$ the number of occurrences of $i$ in $\state$ is at most 1: a handle cannot be grasped by several grippers.
 Note also that the number of occurrences of $\emptyset$ is not limited: several grippers may hold nothing.
@@
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\linewidth]{figures/example.png}
-    \def\svgwidth{\columnwidth}
-    \graphicspath{{./figures/}}
-    \input{figures/example-graph.pdf_tex}
-  \end{center}
-  \caption{Top: two UR3 robots with one gripper each (X=red, Y=green, Z=blue) manipulating a cylinder with two handles.
-The environment contains one rectangular contact surface (in red).
-The cylinder has two rectangular contact surfaces (in green).
-  Bottom: the corresponding constraint graph.
-Names of states follow Expression~(\ref{eq:state}): for instance, $(\emptyset,1)$ means that gripper of robot 2 grasps handle 1 of the cylinder.
-In this state, there is no placement constraint.}
-  \label{fig:2ur3-cylinder}
-\end{figure}
 To illustrate the notions exposed in the previous sections, let us consider an example of two UR3 robots manipulating a cylinder illustrated in Figure~\ref{fig:2ur3-cylinder}.
@@
     \State{\Call{createTransition}{$Gr$, $Gr^{\prime}$}}
-    \If{$isNewState$}{\Call{recurse}{$\mathcal{G}^{\prime}$,$\mathcal{H}^{\prime}$,$Gr^{\prime}$}}\label{cond:recurse}
+    \If{$isNewState$}{\ \Call{recurse}{$\mathcal{G}^{\prime}$,$\mathcal{H}^{\prime}$,$Gr^{\prime}$}}\label{cond:recurse}
     \EndIf
@@
 Given a set of grippers, handles and objects, the constraint graph can be constructed automatically.
-\href{https://github.com/humanoid-path-planner/hpp-manipulation-corba/blob/5af1b3bad68e8c339d5f42eb72173d7356504532/src/hpp/corbaserver/manipulation/constraint_graph_factory.py#L187}{Here} is an implementation in python.
+\href{https://github.com/humanoid-path-planner/hpp-manipulation-corba/blob/5af1b3bad68e8c339d5f42eb72173d7356504532/src/hpp/corbaserver/manipulation/constraint_graph_factory.py\#L187}{Here} is an implementation in python.
 Algorithm~\ref{algo:constraintGraphFactory} describes this implementation.
@@
 In this section, we show how the constraint graph defined in the previous section is used to plan collision-free manipulation paths.
-Although we are working on an extension of Algorithm RMR*~\cite{schmitt17icra} to several grippers, objects and handles, the only manipulation planning algorithm available up to now in HPP is an extension of RRT algorithm described in the next section.
+Although we are working on an extension of the RMR* algorithm~\cite{schmitt17icra} to several grippers, objects and handles, the only manipulation planning algorithm available up to now in HPP is an extension of the RRT algorithm described in the next section.
 
@@
 
-Manipulation Randomly exploring Random Tree is an extension of RRT algorithm~\cite{LavKuf01b} that grows trees in the free configuration space, exploring the different states of the manipulation problem.
-Algorithm~\ref{algo:M-RRT} describes the algorithm implemented in C++ \href{https://github.com/humanoid-path-planner/hpp-manipulation/blob/16369aa291ab1b17ef6176ae8b8b2512b5e6fff7/src/manipulation-planner.cc#L159}{here}.
+Manipulation Randomly exploring Random Tree is an extension of the RRT algorithm~\cite{LavKuf01b} that grows trees in the free configuration space, exploring the different states of the manipulation problem.
+Algorithm~\ref{algo:M-RRT} describes the algorithm implemented in C++ \href{https://github.com/humanoid-path-planner/hpp-manipulation/blob/16369aa291ab1b17ef6176ae8b8b2512b5e6fff7/src/manipulation-planner.cc\#L159}{here}.
 
@@
 Then it returns the linear interpolation between the configurations, projected onto the sub-manifold defined by the transition constraints.
-If the path is in collision, only the collision-free part at the beginning of the path is returned.
+If the path is in collision {\color{blue}or discontinuous}, only the {\color{blue}continuous} collision-free part at the beginning of the path is returned.
 
@@
     \EndIf
-    \If {$p$ collision free} {\Return $p$}
-    \Else { \Return collision-free portion of $p$ starting at $\conf_{near}$}
+    \If {$p$ collision-free \color{blue}and continuous} {\Return $p$}
+    \Else { \Return collision-free {\color{blue}continuous} portion of $p$ starting at $\conf_{near}$}
     \EndIf
@@
   and solvers,
-\item \texttt{hpp-core} a library that implements most of the concept relative
+\item \texttt{hpp-core} a library that implements most of the concepts relative
   to motion planning. The main features are
@@
   \item roadmaps,
-  \item validation of configurations and paths,
+  \item validation of configurations and paths, {\color{blue} notice that this
+    includes an implementation of continuous collision checking first proposed
+    by Schwarzer \textit{et al}~\cite{SchSahLat2004}.}
   \end{itemize}
@@
 \href{https://github.com/humanoid-path-planner/hpp_benchmark/tree/v4.10.0/2020-07-23}{\texttt{hpp\_benchmark}} package. We present here only a few test cases.
-The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with 32 Megabytes of RAM and 3 Gbytes of cache memory.
+The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with 32 Gigabytes of RAM and 3 Gigabytes of cache memory.
 For each test case, we report the minimum, maximum, mean and standard deviation of the time of computation on the one hand, and of the number of nodes in the roadmap built to solve the problem on the other hand.
@@
 The goal configuration is shown on the right.
-The algorithm is a constrained RRT close to the one described in~\cite{DalNakLamLau2009}}.
+The algorithm is a constrained RRT close to the one described in {\color{blue}Dalibard \textit{et al}}~\cite{DalNakLamLau2009}}.
   \label{fig:benchmark-hrp2}
@@
 One test case concerns constrained motion planning.
-The robot is an HRP-2 humanoid robot in quasi-static equilibrium that can slide on the ground.
-This type of motion can be post-processed into a walking motion using the method described in \cite{dalibard:hal-00654175}.
+The robot is an HRP-2 humanoid robot in quasi-static equilibrium that can slide on the ground ( {\color{blue}Figure~\ref{fig:benchmark-hrp2}} ).
+This type of motion can be post-processed into a walking motion using the method described in {\color{blue}Dalibard \textit{et al}}~\cite{dalibard:hal-00654175}.
 The results are displayed in Table~\ref{tab:benchmark-hrp2}.
@@
   \end{center}
-  \caption{Experimental results for HRP-2 sliding on the ground: time of computation and number of nodes.}
+  \caption{Experimental results for HRP-2 sliding on the ground {\color{blue}(36 degrees of freedom)}: time of computation and number of nodes.}
   \label{tab:benchmark-hrp2}
@@
   \end{center}
-  \caption{Experimental results for Baxter robot manipulating two boxes on a table.}
+  \caption{Experimental results for Baxter robot manipulating two boxes on a table {\color{blue}(31 degrees of freedom)}.}
   \label{tab:baxter-manipulation-boxes}
@@
   \end{center}
-  \caption{Experimental results for PR-2 robot manipulating a box on a table.}
+  \caption{Experimental results for PR-2 robot manipulating a box on a table
+     {\color{blue}(39 degrees of freedom)}.}
   \label{tab:pr2-manipulation-two-hands}
@@
   \end{center}
-  \caption{Experimental results for Romeo robot manipulating a placard.}
+  \caption{Experimental results for Romeo robot manipulating a placard
+    {\color{blue}(67 degrees of freedom)}.}
   \label{tab:romeo-placard}
@@
   \end{center}
-  \caption{Experimental results for construction set assembly.}
+  \caption{Experimental results for construction set assembly
+    {\color{blue}(36 degrees of freedom)}.}
   \label{tab:construction-set}
@@
 \item robot 1 releases sphere 2,
-\item robot 2 put cylinder 1 on the ground.
+\item robot 2 puts cylinder 1 on the ground.
 \end{enumerate}
@@
 foliation). The sub-goals are then linked by running a constrained visibility
-PRM algorithm (\cite{SimLauNis00}) on each leaf. The python code can be found
+PRM algorithm \cite{SimLauNis00} on each leaf. The python code can be found
 on \href{https://github.com/humanoid-path-planner/hpp_benchmark/blob/master/2020-07-23/construction-set/script.py}{github.com}.
@@
   \end{center}
-  \caption{UR-5 manipulating a box with and without waypoint transitions}
+  \caption{UR-5 manipulating a ball with and without waypoint transitions}
   \label{tab:waypoint}

### 2021-09-06 hpp.tex
Start applying corrections.

@@
   Bottom: the corresponding constraint graph.
-Names of states follow Expression~(\ref{eq:state}): for instance, $(\emptyset,1)$ means that gripper of robot 2 grasps handle 1 of the cylinder.
+Names of states follow Expression~(\ref{eq:state}): for example, $(\emptyset,1)$ means that gripper of robot 2 grasps handle 1 of the cylinder.
 In this state, there is no placement constraint.}
@@
 
-Robots in industrial manufacturing are today mostly programmed by
+Today, robots in industrial manufacturing are mostly programmed by
 hand. They repeat the same motion thousands of times with a high
-accuracy. Automating a task with some variability is however very
+accuracy. However, automating a task with some variability is very
 challenging since it requires more programming effort to integrate
@@
 
-This \paper therefore describes a software platform called {Humanoid Path
-  Planner} taylored for manipulation planning in robotics. It can handle many types of robots, from manipulator arms to legged humanoid robots.
+Therefore this \paper describes a software platform called {Humanoid Path
+  Planner} tailored for manipulation planning in robotics. It can handle many types of robots, from manipulator arms to legged humanoid robots.
 The main contributions are:
 \begin{itemize}
-\item an original and general modeling of prehensile manipulation based on
-  non-linear constraints,
-\item an original solver for non-linear constraints that is able to handle implicit and explicit constraints,
+\item an original and general modelling of prehensile manipulation based on
+  nonlinear constraints,
+\item an original solver for nonlinear constraints that is able to handle implicit and explicit constraints,
 \item a manipulation planning algorithm that is able to tackle a great variety of manipulation planning problems,
 \item an open-source software suite that implements all the above following state-of-the-art development tools and methods.
-\item a docker image of the above software with installation instructions provided with this paper. This image makes the experimental results replicable.
+\item a docker image of the aforementioned software with installation instructions provided with this paper. This image makes the experimental results replicable.
 \end{itemize}
@@
 \Partie~\ref{partie:preliminary} introduces some preliminary notions like kinematic chains and Lie groups that are used to model the configuration space of each joint.
-\Partie~\ref{sec:constraints} introduces non-linear constraints and solvers that are at the core of manipulation problem definition.
+\Partie~\ref{sec:constraints} introduces nonlinear constraints and solvers that are at the core of manipulation problem definition.
 \Partie~\ref{sec:manipulation-problem} defines the problem of prehensile manipulation in the general setting.
@@
 
-\subsection{Path planning with non-linear constraints}
-
-Some systems are subject to non-linear constraints. These non-linear
+\subsection{Path planning with nonlinear constraints}
+
+Some systems are subject to nonlinear constraints. These nonlinear
 constraints define sub-manifolds of the configuration space the robot
-must stay on. For instance, legged robots that must keep contact with
+must stay on. For example, legged robots that must keep contact with
 the ground and enforce quasi-static equilibrium, or multi-arm systems
@@
 configuration variables with respect to others~\cite{Cortes2002},
-\cite{MirLam2018} when this is possible. {\color{blue}Jaillet \textit{et al}}~\cite{atlasRRT} propose
-another method based on non-linear projection. They cover the
+\cite{MirLam2018} whenever this is possible. {\color{blue}Jaillet \textit{et al}}~\cite{atlasRRT} propose
+another method based on nonlinear projection. They cover the
 constrained manifold by growing an atlas composed of local
@@
 provide an in-depth review of the various approaches to motion
-planning with non-linear constraints.
+planning with nonlinear constraints.
 
@@
 
-Manipulation planning is a particular instance of path planning were
+Manipulation planning is a particular instance of path planning where
 some objects are moved by some robots.
@@
 The configuration space of the whole system is constrained by
-non-linear constraints due to the fact that objects cannot move by
+nonlinear constraints due to the fact that objects cannot move by
 themselves and should stay in a stable pose when not grasped by a
@@
 simplifying assumption is the existence of a monotone solution, that is
-a solution where each object is grasped at most once and is moved from
+a solution where each object is grasped at the most once and is moved from
 its initial pose to its final pose \cite{stilman2008planning},
@@
 graph is present, although not as clearly expressed as in this paper. The main
-contribution of our work with respect to this latter paper is that the
+contribution of our work with respect to the latter paper is that the
 constraint graph is built automatically at the cost of a more restricted range
@@
 OMPL integrates many randomized path planning algorithms and is widely used for teaching.
-Recently, {\color{blue}Kingston \textit{et al}}~\cite{KinMolKav2019} proposed an extension for systems subject to non-linear constraints.
+Recently, {\color{blue}Kingston \textit{et al}}~\cite{KinMolKav2019} proposed an extension for systems subject to nonlinear constraints.
 
@@
 One of the main differences between our solution and the previously cited ones resides in the way manipulation constraints are compiled into a graph.
-To our knowledge none of the previous solutions can handle a variety of problems as large as those exposed in Section~\ref{sec:benchmarks}.
+To our knowledge none of the previous solutions can handle a variety of problems as large as those described in Section~\ref{sec:benchmarks}.
 
@@
 We refer to~{\color{blue}Murray \textit{et al}}~\cite[Appendix A]{LMS94} for a thorough definition of Lie groups.
-We expose here only properties that are useful for the following developments.
+Here we detail only properties that are useful for the following developments.
 
@@
 
-Therefore elements of $SE(3)$ are represented by a vector in $\reals^7$ where the 3 first components represent the image of the origin (vector $\mathbf{t}$ in Equation~\ref{eq:homogeneous matrix}), the 4 last components $(x,y,z,w)$ represent unit quaternion $w + x {i} + y {j} + z {k}$.
+Therefore elements of $SE(3)$ are represented by a vector in $\reals^7$ where the first 3 components represent the image of the origin (vector $\mathbf{t}$ in Equation~\ref{eq:homogeneous matrix}), the last 4 components $(x,y,z,w)$ represent unit quaternion $w + x {i} + y {j} + z {k}$.
 
@@
 Elements of $SE(2)$ are represented by a vector of dimension 4.
-The 2 first components represent the image of the origin.
-The 2 last components represent the cosine and sine of the rotation angle.
+The first 2 components represent the image of the origin.
+The last 2 components represent the cosine and sine of the rotation angle.
 Therefore the homogeneous matrix associated to $\conf=(q_1,q_2,q_3,q_4)$ is
@@
 
-\section{Non-linear constraints and solvers}\label{sec:constraints}
-
-Some tasks require the robot to enforce some non-linear constraints.
+\section{nonlinear constraints and solvers}\label{sec:constraints}
+
+Some tasks require the robot to enforce some nonlinear constraints.
 Foot contact on the ground for a humanoid robot, center of mass projection on a horizontal plane, gaze constraint are a few examples.
 
-\subsection{Non-linear constraints}
-
-\begin{definition}\textbf{Non-linear constraint.} A non-linear constraint is defined by a {\color{blue} piece-wise} differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and is written
+\subsection{nonlinear constraints}
+
+\begin{definition}\textbf{nonlinear constraint.} A nonlinear constraint is defined by a {\color{blue} piece-wise} differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and is written
 \begin{equation}\label{eq:constraint}
@@
 $$
-It may be useful to use a non zero right hand side for the same function $h$. We define for that parameterized non-linear constraints.
-\begin{definition}\textbf{Parameterized non-linear constraint.}
-  A parameterized non-linear constraint is defined by a {\color{blue} piece-wise} differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and by a vector $\mathbf{h}_0$ of $\reals^m$ and is written
+It may be useful to use a non-zero right hand side for the same function $h$. We define for that parameterized nonlinear constraints.
+\begin{definition}\textbf{Parameterized nonlinear constraint.}
+  A parameterized nonlinear constraint is defined by a {\color{blue} piece-wise} differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and by a vector $\mathbf{h}_0$ of $\reals^m$ and is written
   $$
@@
 
-To optimize constraint resolution, we perform variable substitution when possible in order to reduce both the number of variables and the dimension of the resulting implicit constraint.
-We describe here the method that has been first published in~{\color{blue}Mirabel \textit{et al}}~\cite{MirLam2018}.
+To optimize constraint resolution, we perform variable substitution whenever possible in order to reduce both the number of variables and the dimension of the resulting implicit constraint.
+Here we describe the method that has been first published in~{\color{blue}Mirabel \textit{et al}}~\cite{MirLam2018}.
 Unlike in the former paper, the description we give in Algorithm~\ref{alg:ExplicitConstraintSet} is closer to the real implementation.
@@
 Once several compatible explicit constraints have been inserted in the solver,
-they behave as a single one. For instance, if $\conf=(\conf_1,\conf_2,\conf_3)$,
+they behave as a single one. For example, if $\conf=(\conf_1,\conf_2,\conf_3)$,
 $$
@@
 $$
-The \textit{solver by substitution} applies iteration~(\ref{eq:HierarchicalIterative}) to $\tilde{h}$, instead of $h$. We need therefore to compute the Jacobian of  $\tilde{h}$:
+The \textit{solver by substitution} applies iteration~(\ref{eq:HierarchicalIterative}) to $\tilde{h}$, instead of $h$.  Therefore we need to compute the Jacobian of  $\tilde{h}$:
 $$
@@
 In return, the projection of a continuous path may be discontinuous.
-Before inserting a projected path in a roadmap for instance, it is therefore necessary to detect possible discontinuities.
+Thus, before inserting a projected path in a roadmap for example, it is necessary to detect possible discontinuities.
 {\color{blue}Hauser} \cite{Hauser-RSS-13} proposes a solution to this problem.
@@
 classes \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1pathProjector_1_1Dichotomy.html}{\texttt{pathProjector::Dichotomy}} and \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1pathProjector_1_1Progressive.html}{\texttt{pathProjector::Progressive}}.
-{\color{blue} Note that when a path is not continuous, the algorithms returns a continuous portion of the path starting at the beginning of the path.
+{\color{blue} Note that when a path is not continuous, the algorithms return a continuous portion of the path starting at the beginning of the path.
 This enables function \EXTEND in Algorithm~\ref{algo:M-RRT} to create a new node.}
@@
 \begin{definition}\textbf{Grasp complement.}
-  Given a grasp constraint defined by gripper $\gripper$, handle $\handle$ and some flag vector, the grasp complement is a parameterized non-linear constraint defined by
+  Given a grasp constraint defined by gripper $\gripper$, handle $\handle$ and some flag vector, the grasp complement is a parameterized nonlinear constraint defined by
   $$
@@
 \paragraph{Geometric interpretation and examples}
-The 3 first components of $\bar{h}(\conf)$ in equation~(\ref{eq:barh}) correspond to the position of the center of $\handle(\conf)$ in the frame of $\gripper(\conf)$.
-The 3 last components of $\bar{h}(\conf)$ is a vector representing the relative orientation of $\handle(\conf)$ with respect to $\gripper(\conf)$.
+The first 3 components of $\bar{h}(\conf)$ in equation~(\ref{eq:barh}) correspond to the position of the center of $\handle(\conf)$ in the frame of $\gripper(\conf)$.
+The last 3 components of $\bar{h}(\conf)$ is a vector representing the relative orientation of $\handle(\conf)$ with respect to $\gripper(\conf)$.
 The direction of the vector represents the axis of rotation, the norm of the vector represents the angle of rotation.
@@
 
-We propose here a third method that enables users to define contact surfaces in a more flexible way. In this aim, we denote by:
+Here we propose a third method that enables users to define contact surfaces in a more flexible way. In this aim, we denote by:
 \begin{itemize}
 \item $(o_i)_{i\in I}$ a set of convex polygons attached to an object,
-\item $(f_j)_{j\in J}$ a set of convex polygons attached to the environment or to a mobile part of a robot that can receive objects (mobile robot for instance),
+\item $(f_j)_{j\in J}$ a set of convex polygons attached to the environment or to a mobile part of a robot that can receive objects (mobile robot for example),
 \item respectively $C_{o_i}$, $\mathbf{n}_{o_i}$ the barycenter of $o_i$ and the normal to the plane containing $o_i$,
@@
 contains one object pose for each pair of polygons $(o_i,f_j)$, and there are
-$|I| . |J|$ possible combinations. This constraint is therefore not suitable to
+$|I| . |J|$ possible combinations. Thus this constraint is not suitable to
 enforce object immobility along a path since the object may jump from one pose to another one.
@@
 \end{equation}
-where $h_i\in\{\emptyset,1,\cdots,\nhandles\}$ denotes the id of the handle grasped by gripper $i$; $h_i=\emptyset$ means that gripper $i$ does not grasp any handle.
-
-\paragraph{Number of states} note that for $i\in\{1,\cdots,\nhandles\}$ the number of occurrences of $i$ in $\state$ is at most 1: a handle cannot be grasped by several grippers.
+where $h_i\in\{\emptyset,1,\cdots,\nhandles\}$ denotes the index of the handle grasped by gripper $i$; $h_i=\emptyset$ means that gripper $i$ does not grasp any handle.
+
+\paragraph{Number of states} note that for $i\in\{1,\cdots,\nhandles\}$ the number of occurrences of $i$ in $\state$ is at the most 1: a handle cannot be grasped by several grippers.
 Note also that the number of occurrences of $\emptyset$ is not limited: several grippers may hold nothing.
@@
   \item two edges (back and forth) connect two states if they are neighboring,
-  \item one edge connect each state to itself.
+  \item one edge connects each state to itself.
   \end{itemize}
@@
 
-To illustrate the notions exposed in the previous sections, let us consider an example of two UR3 robots manipulating a cylinder illustrated in Figure~\ref{fig:2ur3-cylinder}.
-The robot is equipped with one gripper attached to the end-effector.
+To illustrate the notions expounded in the previous sections, let us consider an example of two UR3 robots manipulating a cylinder illustrated in Figure~\ref{fig:2ur3-cylinder}.
+The robot is fitted with one gripper attached to the end-effector.
 The cylinder is equipped with two handles and with two square contact surfaces corresponding to the top and bottom faces of the cylinder.
@@
 composed of two robots and a cylinder with two handles. The top picture displays
-$\conf_{rand}$. The middle picture displays $\conf_{near}$ that lies in state
+$\conf_{rand}$. The middle picture displays $\conf_{near}$ that belongs to state
 $(\emptyset,\emptyset)$. The transition that is randomly selected
@@
 \end{itemize}
-According to Section~\ref{subsec:explicit}, the two first constraints can be
+According to Section~\ref{subsec:explicit}, the first two constraints can be
 replaced by an explicit constraint: the position of the object can be deduced
@@
 Figure~\ref{fig:connect} illustrates method \CONNECT applied to two configurations $\conf_1$ (top) and $\conf_2$ (bottom).
-Both configurations lie in state $(1,\emptyset)$\footnote{Note that $\conf_1$ lies at the intersection between states
+Both configurations belong to state $(1,\emptyset)$\footnote{Note that $\conf_1$ is at the intersection between states
   $(\emptyset,\emptyset)$ and $(1,\emptyset)$.}.
@@
 
-Even with these simple tricks however, solution paths to a manipulation problem need to come close to collision, raising the well-known issue of narrow passages.
+However even with these simple tricks, solution paths to a manipulation problem need to come close to collision, raising the well-known issue of narrow passages.
 
@@
   \item $(o_i)_{i\in I}$ a set of convex polygons attached to an object,
-  \item $(f_j)_{j\in J}$ a set of convex polygons attached to the environment or to a mobile part of a robot that can receive objects (mobile robot for instance),
+  \item $(f_j)_{j\in J}$ a set of convex polygons attached to the environment or to a mobile part of a robot that can receive objects (mobile robot for example),
   \item a non-negative real number $\Delta$.
@@
 
-In case 2, we replace the transitions between $\state_1$ and $\state_2$ by three additional waypoint states and eight waypoint transitions as explain in Figure~\ref{fig:pregrasp-placement}.
+In case 2, we replace the transitions between $\state_1$ and $\state_2$ by three additional waypoint states and eight waypoint transitions as explained in Figure~\ref{fig:pregrasp-placement}.
 
@@
     $(\emptyset,\emptyset)$ and $(1,\emptyset)$. Each picture represents a
-    waypoint. The last waypoint lies in state $(1,\emptyset)$. $\conf_{near}$
+    waypoint. The last waypoint is in state $(1,\emptyset)$. $\conf_{near}$
     and $\conf_{rand}$ are the same as in Figure~\ref{fig:MRRT-extension}.}
@@
 transition constraints. The right hand side of the transition constraint is
-priorly initialized with the initial configuration.
+first initialized with the initial configuration.
 
 A waypoint transition builds a path by defining a sequence of configurations
-that lie in the intermediate waypoint states, each waypoint is obtained by
+that belong to the intermediate waypoint states, each waypoint is obtained by
 projecting the previous waypoint onto the corresponding manifold.
@@
 waypoints. The random configuration $\conf_{rand}$ is displayed in
-Figure~\ref{fig:MRRT-extension} top. Table~\ref{table:waypoints} shows the
+Figure~\ref{fig:MRRT-extension} top. Table~\ref{table:waypoints} lists the
 waypoint configurations that are produced when extending $\conf_{near}$ toward
@@
 Note that waypoint states are internal to waypoint edges and are not known by
-the constraint graph when determining in which state a configuration lies
+the constraint graph when determining to which state a configuration belongs
 (Algorithm~\ref{algo:M-RRT} lines~17 and 23) and when visiting the states of the
@@
 examples described in the next sections. An archive is provided with this paper.
-Inflate the archive and follow instructions in the README file.
+Decompress the archive and follow instructions in the README file.
 
@@
 planning problems. The raw data can be found in
-\href{https://github.com/humanoid-path-planner/hpp_benchmark/tree/v4.10.0/2020-07-23}{\texttt{hpp\_benchmark}} package. We present here only a few test cases.
+\href{https://github.com/humanoid-path-planner/hpp_benchmark/tree/v4.10.0/2020-07-23}{\texttt{hpp\_benchmark}} package. Here we present only a few test cases.
 The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with 32 Gigabytes of RAM and 3 Gigabytes of cache memory.
@@
 with two grippers and the placard with two handles. Each handle is associated
-to a unique gripper. The number of states of the constraint graph is thus 3.
+to a single gripper. The number of states of the constraint graph is thus 3.
 
@@
 and handles increases, the number of states in the constraint may increase very
-quickly. Using python bindings, it is possible however to define constraint
+quickly. However using python bindings, it is possible to define constraint
 graphs with only the necessary states. We now present a test case that
@@
 For this reason, M-RRT is not able to solve the construction set
-problem in a reasonable amount of time. It remains to our knowledge however
+problem in a reasonable amount of time. However, it remains to our knowledge
 the only algorithm in the literature able to solve a variety of
-problems as large as those exposed in this section.
+problems as large as those presented in this section.
 

### 2021-09-08 hpp.tex
Finish applying corrections.

@@
 the Amazon picking challenge \cite{brock-2018}. The work described in this
-paper is a small step to the simplification of industrial process
+paper is a small step towards simplifying industrial process
 automation in the presence of some variability, like the variation of
-the initial position of some object or the presence of unknown
-obstacles. The work only covers the motion planning -- and
-more accurately the manipulation planning -- issue. The integration in a whole
+the initial position of some object or unknown
+obstacles. The work only covers motion planning -- and
+more accurately manipulation planning. The integration into a whole
 process is still under development. We think that it
@@
 \begin{itemize}
-\item an original and general modelling of prehensile manipulation based on
+\item an original and general modeling of prehensile manipulation based on
   nonlinear constraints,
-\item an original solver for nonlinear constraints that is able to handle implicit and explicit constraints,
-\item a manipulation planning algorithm that is able to tackle a great variety of manipulation planning problems,
+\item an original solver for nonlinear constraints that can handle implicit and explicit constraints,
+\item a manipulation planning algorithm that tackles a great variety of manipulation planning problems,
 \item an open-source software suite that implements all the above following state-of-the-art development tools and methods.
@@
 \end{itemize}
-{\color{blue} Installation instructions can be found by following this \href{https://humanoid-path-planner.github.io/hpp-doc}{link}.}
+{\color{blue} Installation instructions can be found at \href{https://humanoid-path-planner.github.io/hpp-doc}{https://humanoid-path-planner.github.io/hpp-doc}.}
 This paper extends the work presented in {\color{blue} previous papers }\cite{mirabel-iros2016},\cite{MirLam2018} with the following new material:
@@
 
-Motion planning has given rise to a lot of research work for the past decades.
+Motion planning has given rise to a lot of research work over the past decades.
 The problem consists in finding a collision-free path for a given system in
@@
 Planning motions for high dimensional robots like humanoid robots or multi-arm
-systems has been proven to be of a very high complexity~\cite{SchSha1983}, \cite{Canny1983}. Starting in the 1990's random sampling methods have been proposed
+systems has been shown to be highly complex~\cite{SchSha1983}, \cite{Canny1983}. Starting in the 1990's random sampling methods have been proposed
 to solve the problem, trading the completeness property against efficiency in
 solving problems in high dimensional configuration spaces \cite{KSLO1996}, \cite{HsuLatMot99}, \cite{KufLav00}.
-The latter methods are said probabilistically complete since the probability to find a solution if one exists converges to 1 when the time of computation tends to infinity.
+The latter methods are said to be probabilistically complete since the probability to find a solution if one exists converges to 1 when the time of computation tends to infinity.
 Since then, asymptotically optimal random sampling algorithms have
@@
 the ground and enforce quasi-static equilibrium, or multi-arm systems
-grasping a same object are subject to this type of constraints. As the
+grasping the same object are subject to this type of constraints. As the
 volume of the constrained manifold is usually equal to zero, sampling
@@
 \textit{et al}}~\cite{berenson2011} project random configurations using
-a generalization of Newton-Raphson algorithm. Another way to sample
-configurations on a constrained manifold consists in expressing some
+a generalization of Newton-Raphson algorithm. Another method consists in expressing some
 configuration variables with respect to others~\cite{Cortes2002},
-\cite{MirLam2018} whenever this is possible. {\color{blue}Jaillet \textit{et al}}~\cite{atlasRRT} propose
+\cite{MirLam2018} whenever this can be done. {\color{blue}Jaillet \textit{et al}}~\cite{atlasRRT} propose
 another method based on nonlinear projection. They cover the
@@
 space. {\color{blue}Beobkyoon \textit{et al}}~\cite{tb-rrt2016} propose a variation of the latter paper.  The
-main difference resides in the fact that the nodes built on the
-tangent space are not immediately projected on the manifold.
-{\color{blue}Cefalo \textit{et al}}~\cite{CefOri2019} propose a general framework to plan task constrained
-motions in the presence of moving obstacle. {\color{blue}Kingston \textit{et al}}~\cite{KinMolKav2019}
+main difference lies in the fact that the nodes built on the
+tangent space are not immediately projected onto the manifold.
+{\color{blue}Cefalo \textit{et al}}~\cite{CefOri2019} put forward a general framework to plan task-constrained
+motions in the presence of moving obstacles. {\color{blue}Kingston \textit{et al}}~\cite{KinMolKav2019}
 provide an in-depth review of the various approaches to motion
@@
 Manipulation planning is a particular instance of path planning where
-some objects are moved by some robots.
+some objects are moved by robots.
 Although several instances of the
@@
 prehensile manipulation.
-The configuration space of the whole system is constrained by
+The configuration space of the whole system is subject to
 nonlinear constraints due to the fact that objects cannot move by
@@
 The first attempt to solve manipulation planning problems using random sampling
-was proposed by {\color{blue}Siméon \textit{et al}}~\cite{simeon2004ijrr} where a reduction property reduces the
-complexity of the problem.
+was proposed by {\color{blue}Siméon \textit{et al}}~\cite{simeon2004ijrr} where a reduction property simplifies the problem.
 
@@
 final pose. The final configuration of the robot is not specified. A
-simplifying assumption is the existence of a monotone solution, that is
-a solution where each object is grasped at the most once and is moved from
+simplifying assumption is the existence of a monotone solution, where each object is grasped at the most once and is moved from
 its initial pose to its final pose \cite{stilman2008planning},
@@
 \cite{ota2004rearrangement}. They mostly use two-level methods
-composed of a symbolic or task planner and of a motion
+composed of a symbolic task planner and of a motion
 planner~\cite{cambon:hal-01976081}, \cite{KaeLoz2013}, \cite{LozKae2014}, \cite{18-toussaint-RSS}.
@@
 constraint graph is built automatically at the cost of a more restricted range
-of applications. We address indeed only prehensile manipulation.
+of applications. We only address prehensile manipulation.
 
@@
 
-Open-source software platforms are an important tool to make possible fair
+Open-source software platforms are an important tool to enable fair
 comparison between algorithms.
@@
 planning in the robotics community.
-The most popular one is undoubtedly OMPL \cite{ompl2012}.
-OMPL integrates many randomized path planning algorithms and is widely used for teaching.
+Undoubtedly the most popular one is OMPL \cite{ompl2012} which integrates many randomized path planning algorithms and is widely used for teaching purposes.
 Recently, {\color{blue}Kingston \textit{et al}}~\cite{KinMolKav2019} proposed an extension for systems subject to nonlinear constraints.
@@
 
-One of the main differences between our solution and the previously cited ones resides in the way manipulation constraints are compiled into a graph.
-To our knowledge none of the previous solutions can handle a variety of problems as large as those described in Section~\ref{sec:benchmarks}.
+One of the main differences between our solution and the previously cited ones lies in the way manipulation constraints are compiled into a graph.
+To our knowledge, none of the previous solutions can handle such a variety of problems as large as those described in Section~\ref{sec:benchmarks}.
 
@@
 
-This representation although well suited for fixed base manipulator
-arms is not well adapted for robots with a mobile base like wheeled
-mobile robots, aerial robots or legged robots, since the mobility of the base
+Although well-suited for fixed base manipulator
+arms, this representation is ill-suited for robots with a mobile base like wheeled
+mobile, aerial or legged robots, since the mobility of the base
 cannot be correctly represented by translation or rotation joints.
 Representing a free-flying object by three virtual translations followed by
-three virtual rotations called roll, pitch and yaw is indeed a poor workaround
+three virtual rotations referred to as roll, pitch and yaw is indeed a poor workaround
 due to the presence of singularities. A good illustration of this is the gimball
-lock issue that arose during Apollo~13 flight. To avoid singularities, we
-propose the following definition.
-
-\subsubsection{Kinematic chain}
+lock issue that arose during Apollo~13 flight. To avoid singularities, the following definition is proposed.
+
+\subsection{Kinematic chain}
 
 A \textit{kinematic chain} is a tree of \textit{joints} where each
-joint represents a mobility of a rigid-body link with respect to
-another link or with respect to the world reference frame. To each joint is
-associated a configuration space called the \textit{joint configuration space}.
+joint represents the mobility of a rigid-body link with respect to
+another link or to the world reference frame.
+A configuration space called the \textit{joint configuration space} is
+associated to each joint.
 The most common joints with their respective configuration spaces are
@@
 
-\subsubsection{Lie groups}
+\subsection{Lie groups}
 
@@
 
-\paragraph{Geometric interpretations}
+\subsubsection{Geometric interpretations}
 \begin{itemize}
@@
 Note that $\mathbf{x}$ is also the coordinate vector of $g(\mathbf{x})$ in the moving frame $g$.
-If $\mathbf{v}$, $\omega$ are a linear and an angular velocities at the origin, $(\mathbf{v},\omega)$ is \textit{transported} to $g$
+If $\mathbf{v}$, $\omega$ are a linear and angular velocities at the origin, $(\mathbf{v},\omega)$ is \textit{transported} to $g$
   as the same linear and angular velocities expressed in the moving frame. In other words, if
@@
 
-\paragraph{Vector representations}
+\subsubsection{Vector representations}
 Each Lie group element is represented by a vector. Rotations are represented by
@@
 
-Therefore elements of $SE(3)$ are represented by a vector in $\reals^7$ where the first 3 components represent the image of the origin (vector $\mathbf{t}$ in Equation~\ref{eq:homogeneous matrix}), the last 4 components $(x,y,z,w)$ represent unit quaternion $w + x {i} + y {j} + z {k}$.
+Therefore elements of $SE(3)$ are represented by a vector in $\reals^7$ where the first three components represent the image of the origin (vector $\mathbf{t}$ in Equation~\ref{eq:homogeneous matrix}), the last four components $(x,y,z,w)$ represent unit quaternion $w + x {i} + y {j} + z {k}$.
 
@@
 
-\paragraph{Exponential map}
+\subsubsection{Exponential map}
 
@@
 
-\paragraph{Sum and difference notations}
+\subsubsection{Sum and difference notations}
 
@@
 
-\subsubsection{Robot configuration space}
+\subsection{Robot configuration space}
 
@@
 
-\paragraph{Jacobian}
+\subsubsection{Jacobian}
 
 In this \paper, we will make use of the term Jacobian in a generalized way.
-If $h$ is a {\color{blue} piece-wise} differentiable function from a Lie group $\liegroup_1$ to a Lie group $\liegroup_2$, and $\conf_1$ an element of $\liegroup_1$, we will denote by $\frac{\partial h}{\partial\conf}(\conf_1)$ the operator that maps velocities in $T_{\conf_1}\liegroup_1$ to the velocity in $T_{h(\conf_1)}\liegroup_2$ \textit{transported by} $h$\footnote{If $\dconf\in T_{\conf_1}\liegroup_1$ is a velocity along a time parameterized curve $\gamma$, $\frac{\partial h}{\partial\conf}(\conf_1)\dconf$ is the velocity at
-  the same time along Curve $h(\gamma)$.}.
+If $h$ is a {\color{blue} piece-wise} differentiable function from a Lie group $\liegroup_1$ to a Lie group $\liegroup_2$, and $\conf_1$ an element of $\liegroup_1$, we will denote by $\frac{\partial h}{\partial\conf}(\conf_1)$ the operator that maps velocities in $T_{\conf_1}\liegroup_1$ to the velocity in $T_{h(\conf_1)}\liegroup_2$ \textit{transported by} $h$\footnote{If $\dconf\in T_{\conf_1}\liegroup_1$ is a velocity along a time parameterized curve $\gamma$, $\frac{\partial h}{\partial\conf}(\conf_1)\dconf$ is the velocity along Curve $h(\gamma)$.}.
 
@@
 \end{equation}
-where $.^{+}$ denotes the Moore Penrose\footnote{He has just been awarded the Nobel Prize.} pseudo inverse, and $\alpha_i$ is a positive real number called the step size.
-Taking $\alpha_i=1$ exactly solves the linear approximation, but it may not be the best choice in general.
+where $.^{+}$ denotes the Moore Penrose\footnote{who has just been awarded the Nobel Prize.} pseudo inverse, and $\alpha_i$ is a positive real number called the step size.
+Taking $\alpha_i=1$ solves the linear approximation, but it may not be the best choice in general.
 
@@
 
-In manipulation planning applications, where robots manipulate objects, once an object is grasped, the position of the object can be computed explicitly from the configuration of the robot.
-In this case, some configuration variables of the system depend on other:
+In manipulation planning applications, in which robots manipulate objects, once an object is grasped, the position of the object can be explicitly computed from the configuration of the robot.
+In this case, some configuration variables of the system depend on other configuration variables:
 $$
@@
 If $\conf\in\CS$ is a configuration, we denote by $\conf_{I}\in\reals^{|I|}$ the vector composed of the components of $\conf$ of increasing indices in $I$.
-\paragraph {Example} if $\conf = (q_1,q_2,q_3,q_4,q_5,q_6,q_7)$ and $I=\{1,2,6\}$, then $\conf_I=(q_1,q_2,q_6)$, $\conf_{\bar{I}}=(q_3,q_4,q_5,q_7)$.
+\subsubsection {Example} if $\conf = (q_1,q_2,q_3,q_4,q_5,q_6,q_7)$ and $I=\{1,2,6\}$, then $\conf_I=(q_1,q_2,q_6)$, $\conf_{\bar{I}}=(q_3,q_4,q_5,q_7)$.
 
@@
 
-\paragraph{Example}
+\subsubsection{Example}
 If $m=3$, $n=4$, $M=\{2,3\}$ and $N=\{1,2,4\}$,
@@
 To optimize constraint resolution, we perform variable substitution whenever possible in order to reduce both the number of variables and the dimension of the resulting implicit constraint.
-Here we describe the method that has been first published in~{\color{blue}Mirabel \textit{et al}}~\cite{MirLam2018}.
+Here we describe the method first published in~{\color{blue}Mirabel \textit{et al}}~\cite{MirLam2018}.
 Unlike in the former paper, the description we give in Algorithm~\ref{alg:ExplicitConstraintSet} is closer to the real implementation.
@@
 Once several compatible explicit constraints have been inserted in the solver,
-they behave as a single one. For example, if $\conf=(\conf_1,\conf_2,\conf_3)$,
+they behave as a single constraint. For example, if $\conf=(\conf_1,\conf_2,\conf_3)$,
 $$
@@
 
-\paragraph{Substitution} When an explicit constraint is not successfully added following Algorithm~\ref{alg:ExplicitConstraintSet}, it is handled as an implicit constraint.
+\subsubsection{Substitution} When an explicit constraint is not successfully added following Algorithm~\ref{alg:ExplicitConstraintSet}, it is handled as an implicit constraint.
 Therefore, after inserting implicit and explicit constraints, the solver stores a system of equations equivalent to one explicit and one implicit constraints that we denote by:
@@
 Other coefficients of $J_{i}J$ are equal to the corresponding coefficients of $J$.
-An implementation of the above Jacobian product can be found \href{https://github.com/humanoid-path-planner/hpp-constraints/blob/e21490c8c713949bd3038dccb6fe02cf254a615f/src/explicit-constraint-set.cc\#L267}{here}.
+An implementation of the aforementioned Jacobian product can be found \href{https://github.com/humanoid-path-planner/hpp-constraints/blob/e21490c8c713949bd3038dccb6fe02cf254a615f/src/explicit-constraint-set.cc\#L267}{here}.
 
@@
 
-\paragraph{Important remark} As mentioned in Table~\ref{tab:vector representation}, the configuration and velocity vectors may have different sizes.
+\subsubsection{Important remark} As mentioned in Table~\ref{tab:vector representation}, the configuration and velocity vectors may have different sizes.
 As a consequence, index sets $in$ and $out$ in Definition~\ref{def:Explicit} correspond to configuration vector indices, while in Expression~(\ref{eq:jacobian-explicit}), they correspond to velocity vector indices.
@@
 
-Now that we are able to project configurations onto sub-manifolds defined by numerical constraints -- up to some numerical threshold, we need now to define paths on such sub-manifolds.
-The usual way of doing so is to discretize the path and to project each sample configuration.
-The shortcoming of this method is that it requires to choose a discretization step at path construction and to lose the continuous information of the path.
+Now that we are able to project configurations onto sub-manifolds defined by numerical constraints -- up to some numerical threshold, we need to define paths on such sub-manifolds.
+The usual way of doing so is by discretizing the path and projecting each sample configuration.
+The shortcoming is that it requires choosing a discretization step at path construction thus losing the continuous information of the path.
 
@@
 
-\paragraph{Configuration space} The configuration space of a manipulation problem is the Cartesian product of the configuration spaces of the robots and of the objects.
+\subsubsection{Configuration space} The configuration space of a manipulation problem is the Cartesian product of the configuration spaces of the robots and of the objects.
 $$
@@
   \end{itemize}
-  Let $\bar{h}$ be the smooth mapping from $\CS$ to $\reals^6$ that maps to any configuration $\conf$ of the system, the expression
+  Let $\bar{h}$ be the smooth mapping from $\CS$ to $\reals^6$ defined by
   \begin{equation}\label{eq:barh}
@@
 
-\paragraph{Geometric interpretation and examples}
-The first 3 components of $\bar{h}(\conf)$ in equation~(\ref{eq:barh}) correspond to the position of the center of $\handle(\conf)$ in the frame of $\gripper(\conf)$.
-The last 3 components of $\bar{h}(\conf)$ is a vector representing the relative orientation of $\handle(\conf)$ with respect to $\gripper(\conf)$.
+\subsubsection{Geometric interpretation and examples}
+The first three components of $\bar{h}(\conf)$ in equation~(\ref{eq:barh}) correspond to the position of the center of $\handle(\conf)$ in the frame of $\gripper(\conf)$.
+The last three components of $\bar{h}(\conf)$ are a vector representing the relative orientation of $\handle(\conf)$ with respect to $\gripper(\conf)$.
 The direction of the vector represents the axis of rotation, the norm of the vector represents the angle of rotation.
@@
 $$
-contains all the configurations that are reachable from $\conf_0$ while keeping the grasp.
+contains all the configurations that are reachable from $\conf_0$ while maintaining the grasp.
 Note that this representation of relative pose constraints has been used in the Stack of Task software, although it is not described in the corresponding paper~\cite{SoT}.
@@
 \item defining virtual grippers in the environment and virtual handles on the object, implicitly defines a discrete set of poses,
-\item defining a virtual gripper on an horizontal plane and a virtual handle on the object, and using a grasp with flags $(\false,\false,\true,\true,\true,\false)$ constrains the object to move on a infinite horizontal plane.
+\item defining a virtual gripper on a horizontal plane and a virtual handle on the object, and using a grasp with flags $(\false,\false,\true,\true,\true,\false)$ constrains the object to move on an infinite horizontal plane.
 \end{enumerate}
@@
 We denote by $o_i(\conf)$ and $f_j(\conf)$ the poses in configuration $\conf$ of frames with respective centers $C_{o_i}$ and $C_{f_j}$ and with $x$-axis normal to each polygon.
-In an analogous way as in Definition~\ref{def:grasp}, we define
+Similarly as in Definition~\ref{def:grasp}, we define
 \begin{equation}\label{eq:convex-shape-contact-hbar}
@@
 \end{equation}
-It is straightforward that the above function vanishes if and only if two convex polygons $o_i$ and $f_j$ are in contact and $C_{o_i}$ is inside $f_j$.
+It is straightforward that this function vanishes if and only if two convex polygons $o_i$ and $f_j$ are in contact and if $C_{o_i}$ is inside $f_j$.
 
@@
 $|I| . |J|$ possible combinations. Thus this constraint is not suitable to
-enforce object immobility along a path since the object may jump from one pose to another one.
+enforce object immobility along a path since the object may jump from one pose to another.
 
@@
 
-\paragraph{Number of states} note that for $i\in\{1,\cdots,\nhandles\}$ the number of occurrences of $i$ in $\state$ is at the most 1: a handle cannot be grasped by several grippers.
+\subsubsection{Number of states} note that for $i\in\{1,\cdots,\nhandles\}$ the number of occurrences of $i$ in $\state$ is at the most 1: a handle cannot be grasped by several grippers.
 Note also that the number of occurrences of $\emptyset$ is not limited: several grippers may hold nothing.
-Let $m$ be a non-negative integer not greater than $\ngrippers$ nor $\nhandles$ and let us count the number of states with $m$ handles grasped.
+Let $m$ be a non-negative integer not greater than $\min(\ngrippers,\nhandles)$ and let us count the number of states with $m$ handles grasped.
 The number of subset of $m$ handles among $\nhandles$ is equal to $\frac{\nhandles!}{(\nhandles-m)!m!}$.
@@
 $$
-\begin{definition}\label{def:neighboring-states}\textbf{Neighboring states}\\
-  Two states $\state_1=(h_{1 1},\cdots,h_{\ngrippers 1})$ and $\state_2=(h_{1 2},\cdots,h_{\ngrippers 2})$ are neighboring if they differ by only one grasp and the grasp is empty in one of the states:
+\begin{definition}\label{def:neighboring-states}\textbf{adjacent states}\\
+  Two states $\state_1=(h_{1 1},\cdots,h_{\ngrippers 1})$ and $\state_2=(h_{1 2},\cdots,h_{\ngrippers 2})$ are adjacent to each other if they differ by only one grasp and the grasp is empty in one of the states:
   \begin{align*}
@@
   \item the nodes of which are states defined by subsets of grasps~(\ref{eq:state}),
-  \item two edges (back and forth) connect two states if they are neighboring,
+  \item two edges (back and forth) connect two states if they are adjacent to each other,
   \item one edge connects each state to itself.
@@
 
-\subsection{Example}\label{subsec:ur3-cylinder}
+\subsubsection{Example}\label{subsec:ur3-cylinder}
 
@@
 The robot is fitted with one gripper attached to the end-effector.
-The cylinder is equipped with two handles and with two square contact surfaces corresponding to the top and bottom faces of the cylinder.
+The cylinder is equipped with two handles and with two square contact surfaces corresponding to the top and bottom sides of the cylinder.
 $\ngrippers=2,\nhandles=2,\nobjects=1$.
@@
 \end{table}
-Table~\ref{tab:state-constraints} indicates which constraints are active for each state. Table~\ref{tab:transition-constraints} indicates which constraints are active for each transition.
+Table~\ref{tab:state-constraints} indicates which constraints are active for each state and Table~\ref{tab:transition-constraints} for each transition.
 
@@
 
-\subsection{Automatic construction}
+\subsubsection{Automatic construction}
 
@@
 In this section, we show how the constraint graph defined in the previous section is used to plan collision-free manipulation paths.
-Although we are working on an extension of the RMR* algorithm~\cite{schmitt17icra} to several grippers, objects and handles, the only manipulation planning algorithm available up to now in HPP is an extension of the RRT algorithm described in the next section.
+Although we are working on an extension of the RMR* algorithm~\cite{schmitt17icra} to several grippers, objects and handles, the only manipulation planning algorithm available so far in HPP is an extension of the RRT algorithm described in the next section.
 
@@
 After initializing the roadmap with the initial and goal configurations, the algorithm iteratively calls method {\scriptsize ONE}{\small S}{\scriptsize TEP} until a solution path is found or the maximum number of iterations is reached.
-This latter method shoots a random configuration (line 6) and for each connected component of the roadmap and each state of the constraint graph, extends the nearest node in the direction of the random configuration (lines 7--10).
-For each successful extension, the end of the extension path is stored for later connection (line 11).
+This method picks a random configuration (line 6) and for each connected component of the roadmap and each state of the constraint graph, extends the nearest node in the direction of the random configuration (lines 7--10).
+For each successful extension, the end of the extension path is stored for subsequent connections (line 11).
 After the extension step, the algorithm tries to connect new nodes to other connected components using two strategies:
@@
 
-In this section, we illustrate the algorithm described in the previous section with two examples.
-Figure~\ref{fig:MRRT-extension} illustrates function \EXTEND defined in the previous section on the example of Figure~\ref{fig:2ur3-cylinder}.
+In this section, the algorithm described in the previous section is depicted with two examples.
+Figure~\ref{fig:MRRT-extension} shows function \EXTEND defined in the previous section on the example of Figure~\ref{fig:2ur3-cylinder}.
 The system considered is
-composed of two robots and a cylinder with two handles. The top picture displays
-$\conf_{rand}$. The middle picture displays $\conf_{near}$ that belongs to state
+composed of two robots and a cylinder with two handles. The picture at the top displays
+$\conf_{rand}$. The picture in the middle displays $\conf_{near}$ that belongs to state
 $(\emptyset,\emptyset)$. The transition that is randomly selected
@@
 try to grasp handle 1. According to tables~\ref{tab:state-constraints} and
-\ref{tab:transition-constraints}, the constraints of the transition are
+\ref{tab:transition-constraints}, the transition constraints are
 $(place_1, place_1/comp)$.
-The first one is of type~(\ref{eq:convex-shape-contact}), the second one is
+The first one is of type~(\ref{eq:convex-shape-contact}), the second
 of type~(\ref{eq:convex-shape-contact-complement}) and is parameterized: the
@@
 According to Section~\ref{subsec:explicit}, the first two constraints can be
-replaced by an explicit constraint: the position of the object can be deduced
+replaced by an explicit constraint: the position of the object can be derived
 from the right hand side of $place_1/comp$ that is initialized with configuration
@@
 on the configuration variables of robot 1 (6 variables). The solution found by the solver, $\conf_{target}$ (line 39) is displayed in
-Figure~\ref{fig:MRRT-extension} bottom. Notice that as expected, the position of the object is the same in $\conf_{target}$ as in $\conf_{near}$.
+Figure~\ref{fig:MRRT-extension} at the bottom. Notice that as expected, the position of the object is the same in $\conf_{target}$ as in $\conf_{near}$.
 
@@
   \end{center}
-  \caption{Along a transition where an object already grasped is grasped by
-    an additional gripper, an intermediate waypoint state called
+  \caption{Along a transition where an object already grasped is grasped a second time, an intermediate waypoint state called
     \textit{pregrasp} (pg) is added. This intermediate state is represented by
@@
     gripper 2 is going to grasp handle 2 from the state where gripper 1 grasps
-    handle 1. The constraints associated to this waypoint state are the
-    constraints of the state with the least active grasps (here $(1,\emptyset)$)
-    and the pregrasp constraint corresponding to the grasp that is added in
-    the node with the most active grasps (here $pregrasp_{22}$). The transition
-    constraints are the same for all transitions (in red) and the same as
+    handle 1. The constraints associated to this waypoint state are those
+    of the state with the least active grasps (here $(1,\emptyset)$)
+    and the pregrasp constraint corresponding to the new grasp (here $pregrasp_{22}$). The transition
+    constraints are the same for all transitions (in red) and identical to
     the loop transition of the state with the least active grasps (in blue: here
@@
 
-A prehensile manipulation motion contains by definition configurations that are in contact:
+By definition, a prehensile manipulation motion contains configurations that are in contact:
 \begin{itemize}
@@
 \end{itemize}
-Contacts are difficult to handle by classical collision detection libraries and are often considered as collisions.
+Contacts are difficult to handle using classical collision detection libraries and are often considered as collisions.
 To overcome this issue, we keep the gripper open during grasp, and objects slightly above contact surfaces in stable poses.
@@
 
-To cope with this issue, we define intermediate states in the
+To cope with this, we define intermediate states in the
 constraint graph called waypoint states. These states are inserted
@@
   \end{itemize}
-  Let $\bar{h}$ be the smooth mapping from $\CS$ to $\reals^6$ that maps to any configuration $\conf$ of the system, the expression
+  Let $\bar{h}$ be the smooth mapping from $\CS$ to $\reals^6$ defined by
   \begin{equation}\label{eq:barh-pregrasp}
@@
 Note that when this constraint is satisfied, the handle is translated
-along $x$ axis by distance $\Delta$ compared to a configuration
+along $x$ axis over a distance $\Delta$ compared to a configuration
 satisfying the grasp constraint. The value of $\Delta$ depends on the geometry
@@
 \end{definition}
-Note that when this constraint is satisfied, the object is translated of a
+Note that when this constraint is satisfied, the object is translated over a
 distance $\Delta$ along the normal to the contact surface.
@@
 Section~\ref{subsec:constraint graph} by a sequence of intermediate
-states and transitions as follows.
-
-Following definition~\ref{def:neighboring-states}, if two states $\state_1$
-and $\state_2$ are neighbors, one of them contains an additional grasp with
-respect to the other one. Without loss of generality, let us consider that
+states and transitions as follows: given
+Definition~\ref{def:neighboring-states}, if two states $\state_1$
+and $\state_2$ are adjacent to each other, one of them contains an additional grasp with
+respect to the other. Without loss of generality, consider that
 $\state_2$ contains the additional grasp $\grasp{\gripper_i}{\handle_j}$,
-$i\in\{1,\cdots\ngrippers\}$, $j\in\{1,\cdots\nhandles\}$. Then there are two
-cases:
+$i\in\{1,\cdots\ngrippers\}$, $j\in\{1,\cdots\nhandles\}$. Let us denote by
+$\object{}$ the object handle $\handle_j$ belongs to. Then either
 \begin{enumerate}
-\item the object handle $\handle_j$ belongs to is already grasped in state
-  $\state_1$, or
-\item the object handle $\handle_j$ belongs to is in placement is
-  state $\state_1$.
+\item $\object{}$ is already grasped in state $\state_1$, or
+\item $\object{}$ is in placement in  state $\state_1$.
 \end{enumerate}
@@
 
-\paragraph{Construction of a path along a waypoint transition} Function
+\subsubsection{Construction of a path along a waypoint transition} Function
 {\scriptsize EXTEND} in Algorithm~\ref{algo:M-RRT} builds a path along a
@@
 A waypoint transition builds a path by defining a sequence of configurations
-that belong to the intermediate waypoint states, each waypoint is obtained by
-projecting the previous waypoint onto the corresponding manifold.
+that belong to the intermediate waypoint states, each configuration being obtained by
+projecting the previous configuration onto the corresponding manifold.
 Figure~\ref{fig:MRRT-extension-wp} proposes an example of extension along
@@
 waypoints. The random configuration $\conf_{rand}$ is displayed in
-Figure~\ref{fig:MRRT-extension} top. Table~\ref{table:waypoints} lists the
+Figure~\ref{fig:MRRT-extension}, top. Table~\ref{table:waypoints} lists the
 waypoint configurations that are produced when extending $\conf_{near}$ toward
@@
 \end{table}
-\paragraph{Implementation} From an implementation point of view, Class
+\subsubsection{Implementation} From an implementation point of view, Class
 \href{https://gepettoweb.laas.fr/hpp/hpp-manipulation/doxygen-html/classhpp_1_1manipulation_1_1graph_1_1WaypointEdge.html}{\texttt{WaypointEdge}} derives from class \href{https://gepettoweb.laas.fr/hpp/hpp-manipulation/doxygen-html/classhpp_1_1manipulation_1_1graph_1_1Edge.html}{\texttt{Edge}}.
@@
 
-Note that waypoint states are internal to waypoint edges and are not known by
+Note that waypoint states are internal to waypoint edges and thus not known by
 the constraint graph when determining to which state a configuration belongs
@@
 
-In this section, we describe in more details the software platform
+In this section, we describe in greater details the software platform
 \href{https://humanoid-path-planner.github.io/hpp-doc}{Humanoid Path
@@
 \begin{itemize}
-\item \texttt{hpp-fcl} a modified version of \href{https://github.com/flexible-collision-library/fcl}{\texttt{fcl}}. The main additional features are the following:
+\item \texttt{hpp-fcl} a modified version of \href{https://github.com/flexible-collision-library/fcl}{\texttt{fcl}}. The main additional features are:
   \begin{itemize}
@@
 that implements CORBA services. These services can be extended via a plugin
-system. The application can then be controlled via python script or C++ code.
+system. The application can then be controlled with python scripts or C++ code.
 CORBA clients are provided in python and C++. The packages implementing CORBA
@@
 \item \texttt{hpp-manipulation-corba} for manipulation problems. This
-latter package also provides an implementation of the automatic constraint graph
+package also provides an implementation of the automatic constraint graph
 construction in python.
@@
 planning problems. The raw data can be found in
-\href{https://github.com/humanoid-path-planner/hpp_benchmark/tree/v4.10.0/2020-07-23}{\texttt{hpp\_benchmark}} package. Here we present only a few test cases.
+\href{https://github.com/humanoid-path-planner/hpp_benchmark/tree/v4.10.0/2020-07-23}{\texttt{hpp\_benchmark}} package. Here we only present a few test cases.
 The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with 32 Gigabytes of RAM and 3 Gigabytes of cache memory.
-For each test case, we report the minimum, maximum, mean and standard deviation of the time of computation on the one hand, and of the number of nodes in the roadmap built to solve the problem on the other hand.
+For each test case, we report the minimum, maximum, mean and standard deviation of the time of computation on the one hand, and of the number of nodes in the roadmap built to solve the problem on the other.
 
@@
 The robot has two grippers and each box is equipped with a handle.
-The constraint graph thus contains 7 nodes.
+Thus the constraint graph contains seven nodes.
 The experimental results are displayed in Table~\ref{tab:baxter-manipulation-boxes}.
@@
 The robot is requested to flip the box upside down from an initial pose to a goal pose as represented in Figure~\ref{fig:pr2-manipulation-two-hands}.
-The robot is equipped with two grippers, the box is equipped with two handles.
-The constraint graph contains 7 nodes.
+The robot is equipped with two grippers and the box with two handles.
+The constraint graph contains seven nodes.
 Table~\ref{tab:pr2-manipulation-two-hands} shows the experimental results.
@@
 The third test case features Humanoid Robot Romeo manipulating a placard. The
-robot is requested to rotate the placard by 180 degrees. The robot is equipped
+robot is requested to rotate the placard by 180 degrees. It is equipped
 with two grippers and the placard with two handles. Each handle is associated
-to a single gripper. The number of states of the constraint graph is thus 3.
+to a single gripper. The number of states of the constraint graph is thus three.
 
@@
 In the three previous test cases, the constraint graph was automatically built
-following Algorithm~\ref{algo:constraintGraphFactory}. If the number of grippers
-and handles increases, the number of states in the constraint may increase very
+by Algorithm~\ref{algo:constraintGraphFactory}. If the number of grippers
+and handles increases, the number of states in the constraint graph may increase very
 quickly. However using python bindings, it is possible to define constraint
 graphs with only the necessary states. We now present a test case that
-illustrates this possibility. The system is represented in Figure~\ref{fig:construction-set}.
+illustrates this possibility. The system is depicted in Figure~\ref{fig:construction-set}.
 
 In this example, an operator provides the sequence of actions (transitions)
-the system needs to go through:
+the system needs to perform:
 \begin{enumerate}
@@
 \item robot 2 grasps cylinder 1,
-\item robot 1 sticks sphere 1 on cylinder 1,
+\item robot 1 sticks sphere 1 to cylinder 1,
 \item robot 1 releases sphere 1,
 \item robot 1 grasps sphere 2,
-\item robot 1 sticks sphere 2 on cylinder 1,
+\item robot 1 sticks sphere 2 to cylinder 1,
 \item robot 1 releases sphere 2,
@@
 From this sequence of actions, the sequence of states visited is
-computed and only those states (9 in total) are built in the
+computed and only those states (nine in total) are built in the
 constraint graph. Then, a sequence of sub-goals in the successive
@@
 PRM algorithm \cite{SimLauNis00} on each leaf. The python code can be found
-on \href{https://github.com/humanoid-path-planner/hpp_benchmark/blob/master/2020-07-23/construction-set/script.py}{github.com}.
+at \href{https://github.com/humanoid-path-planner/hpp_benchmark/blob/master/2020-07-23/construction-set/script.py}{github.com}.
 
@@
 now empirically show the positive effect of waypoints on the
-efficiency of manipulation planning. For that, we run 20 times
+efficiency of manipulation planning. To do that, we run 20 times
 Algorithm M-RRT on the same problem with and without waypoint
@@
 ball as shown in Figure~\ref{fig:ur5-pokeball}. The results are
-reported in Table~\ref{tab:waypoint}. We can notice that for this
-example, the waypoint transitions make the time of computation and the
-number of nodes decrease by 2 orders of magnitude. This can be
-explained by the fact that in grasp configurations, the gripper is
+reported in Table~\ref{tab:waypoint}. We can notice in this
+example, that waypoint transitions decrease the computation time and the
+number of nodes by two orders of magnitude. This is because in grasp
+configurations, the gripper is
 very close to the object and only a small part of the approaching
 directions of the gripper toward the object leads to collision-free
-paths. Waypoint states on the contrary are away from obstacles and
+paths. On the contrary waypoint states are away from obstacles and
 easier to reach. The transition between the pregrasp waypoint and the
-grasp $\cap$ placement waypoint is always collision-free.
+grasp $\cap$ placement waypoint is {\color{blue}almost} always collision-free.
 
@@
 The experimental results show that M-RRT is able to solve a variety of
-manipulation problems including with a legged robot in quasi-static
+manipulation problems including that of a legged robot in quasi-static
 equilibrium. No parameter tuning is required between the different problems. All
@@
 
-As for any random motion planning method, we observe a large standard
-deviation within the 20 runs of each test case, both for the number of nodes
-and for the time of computation.
+As in any random motion planning method, we observe a large standard
+deviation between the 20 runs of each test case, for the number of nodes
+as well as for the time of computation.
 
@@
 \end{enumerate}
-For this reason, M-RRT is not able to solve the construction set
-problem in a reasonable amount of time. However, it remains to our knowledge
-the only algorithm in the literature able to solve a variety of
+Thus, M-RRT is not able to solve the construction set
+problem within a reasonable amount of time. However, to our knowledge it is
+the only algorithm in the literature capable of solving a variety of
 problems as large as those presented in this section.
@@
 This paper presents a software platform aimed at prototyping and
-solving a great variety of prehensile manipulation planning
+solving a large number of prehensile manipulation planning
 problems. The platform provides an original algorithm M-RRT that is an
@@
 
-We think that this platform is a perfect basis for researchers who want to
+It is the authors' opinion that this platform is perfect for researchers who want to
 develop and benchmark new manipulation planning algorithms. Note that some
-on-going work in humanoid locomotion  \cite{tonneau-2018} is based on HPP.
+of the on-going work in humanoid locomotion  \cite{tonneau-2018} is based on HPP.
 
@@
 that can handle use cases as diverse as those proposed in the benchmark section.
-A good candidate is a generalization of RMR*~\cite{schmitt17icra}. We also aim at working on manipulation path optimization since paths computed by
-random algorithms are too long to be applied on real robots as such.
+A good candidate is a generalization of RMR*~\cite{schmitt17icra}. Also we intend to focus on manipulation path optimization since paths computed by
+random algorithms are too long to be applied to real robots as such.

### 2021-09-08 review/reply-to-reviewers.tex
Add answer to reviewer comments.

\documentclass{article}
\title{Paper IEEE-TRO 21-0184: reply to reviewers}
\author{Florent Lamiraux and Joseph Mirabel}
\date{}
\begin{document}
\maketitle
\subsection*{Reviewer 7}
\begin{enumerate}
  \item \textbf{Comment.} The beginning of the main content of the
    paper seems hard to follow for readers not familiar with
    manipulation planning.  I think this could be easily improved by
    adding one or two explanatory images earlier on in the
    paper.\\
    \textbf{Answer.} We agree that the first picture comes very lately in the
    initial version of the paper. To improve the paper, we have moved Figure 2
    at the beginning of the paper to provide quickly an example of manipulation
    planning problem.
\end{enumerate}

\subsection*{Reviewer 9}
\begin{enumerate}
\item \textbf{Comment.} In the modeling of SE(2) and SO(2), why using 2
  coordinates for the angular position (sin and cos)? In 2D, using just one
  coordinate (the unique rotation angle) is not enough?\\
  \textbf{Answer.} Using $(\sin\theta,\cos\theta)$ makes the parameterization
  continuous. We added a comment in the paper.

\item \textbf{Comment.} A bit after equation (3), the unnumbered
  equation $q(t)=\cdots$ is the minus sign correct? Should it be the
  operator defined in equation (3)?
  \textbf{Answer.} You are perfectly right. We fixed the error.

\item \textbf{Comment.} I think that some words are supposed to be
  hyperlinks, like ``here''. They do not work.\\
  \textbf{Answer.} You
  are right, there are many hyperlinks in the paper. We checked them
  one by one and found that they all work.

\item \textbf{Comment.}  Algorithm 1: Line 6 checks whether any output
  variable of E is already computed by a previous explicit
  constraint. What if two explicit constraints are redundant? For
  instance, if we have two constraints that (in addition to other
  configuration variables) set the same identical constraint to the
  same config variable.  \\
  \textbf{Answer.} As stated at the beginning of Paragraph IV.D.a, "when an
  explicit constraint is not successfully added following Algorithm 1, it is
  handled as an implicit constraint."

\item \textbf{Comment.} what happens if you detect that a projected
  path is not continuous?\\
  \textbf{Answer.} In this case, only a
  continuous part starting at the beginning of the path is used by the
  M-RRT algorithm. We have added a remark in paragraph and in IV.E.1
  Algorithm 4.

\item \textbf{Comment.} Has the final configuration to be fully
  described? E.g., can I provide just a final desired object pose in
  the cartesian space but not specify the corresponding robot inverse
  kinematic solution (that could be not unique)?\\
  \textbf{Answer.}
  The M-RRT algorithm as of today is not able to define the goal state
  by some constraints, although it would not be difficult to modify
  it for that. However, the python bindings enable users to quickly develop
  new algorithms where the goal can be defined by some constraints. In the
  construction set example described in Section VII.B.2) and in Figure 13, the
  goal is a state of the constraint graph.

\item \textbf{Comment.} Section V.A Grasp: the equation at the
  beginning. Should it be $C = C_{r_1} \times \cdots \times C_{r_{nr}} \times
  SE(3)^{no}$? (I added the dots).\\
  \textbf{Answer.} You are perfectly right. We fixed the error.

\item \textbf{Comment.} In equation (13) (and also in other
  equations), is $R^3 \times SO(3)$ the same as $SE(3)$?\\
  \textbf{Answer.} They both represent rigid-body motion but they have different
  Lie group operators. We have added a paragraph after Equation (13) to clarify
  this point.

\item \textbf{Comment.} The constraint that arises from the stable
  contact pose is piece-wise differentiable. While in the previous
  section the constraints are required to be differentiable. Is there
  any issue?\\
  \textbf{Answer.} Piece-wise differentiable is enough. We have replaced all
  instances of "differentiable" by "piece-wise differentiable".

\item \textbf{Comment.} Section D, after a) Number of states, there is an extra :.\\
  \textbf{Answer.} This has been fixed.

\item \textbf{Comment.}   The used PC has 23 MB of RAM and 3GB of cache. Please
  check these numbers.\\
  \textbf{Answer.} This is 32 GB of RAM and 3 GB of cache. This has been fixed.

\item \textbf{Comment.} I think that Figure 9 is not cited in the text.\\
  \textbf{Answer.} This has been fixed.

\item \textbf{Comment.}  Since the paths are continuous, how is the
  collision checking carried out? Moreover, If I well understand,
  during the waypoint transition the collision checker is not
  called. Is it implicitly assumed here that in such transitions there
  is no collision?\\
  \textbf{Answer.} The software platform proposes discretized collision checking
  as well as continuous collision checking. We added a sentence and a reference
  in the description of the framework. Collisions are checked also along
  waypoint transitions. We added the word "almost" at the end of Section VII-B3). Since the sentence was confusing.
\end{enumerate}

\subsection*{Reviewer 16}
\begin{enumerate}
\item \textbf{Comment.} In the literature survey, it may be good to refer to
  the paper by Lozano-Perez and Kaelbling
  (https://lis.csail.mit.edu/pubs/tlpk-iros14.pdf) because it very much is
  related to representing manipulation planning problems with constraint graphs.\\
  \textbf{Answer.} Although we had already cited 2 papers from this team, we
  have added the requested reference.

\item \textbf{Comment.} Towards the end of Section III, the paper says
  "We denote by $iq_i$ and $iv_i$ the indices of joint $i$ ...". But there
  are multiple indices of joint $i$ (namely $nq_i$ and $nv_i$ many). So
  maybe the sentence should say "the starting (or first?) indices of
  joint i">\\
  \textbf{Answer.} You are right. This has been fixed.

\item \textbf{Comment.} At the end of Section IV-B, the paper refers
  to different linesearch methods implemented. It may make sense to
  briefly describe what these different methods do, or give a
  reference next to each, so that the reader can understand what they
  refer to if interested.\\
  \textbf{Answer.} We have added a reference for one of them and explanations
  for the other ones.

\item \textbf{Comment.} In Section IV-E, the paper says the
  constraints are applied only during "path evaluation". It is not
  perfectly clear what is meant by "path evaluation". What does
  "evaluating" a path mean?\\
  \textbf{Answer.} "Path evaluation" means "computing the configuration at a
  given parameter". We have added this definition in the paper.

\item \textbf{Comment.} In Section V, the paper defines the
  manipulation problem, and as admissible configuration properties, I
  wonder if it should also mention being "collision-free" as a
  requirement. It is kind of obvious, but may be good to be
  complete.\\
  \textbf{Answer.} You are right. We added the absence of collision in the
  definition of admissible configurations.

\item \textbf{Comment.} At the beginning of Section V-A, the combined
  configuration space is defined, but it seems to include only the
  first and the last robot ($C_{r1}$ and $C_{rnr}$). Where are the other
  robots in betwwen 1 and nr?\\ \textbf{Answer.} You are right. This has
  been fixed.

\item \textbf{Comment.} In Section V-A, the freedom of a grasp seems
  to be defined only using binary true/false flags. Using this method,
  is it possible to define limits? For example, for a cylinder, the
  gripper may be able to slide up and down along the height of the
  cylinder. Is it possible to represent using this approach, or would
  one need to discretize along the height of the cylinder? It would be
  good to state this explicitly (whether possible or not). (This
  capability [e.g. limits on translation on a table] is also what is
  needed for the "Stable contact pose" formalization in Section
  V-B. This raises the question: is there really a need to have two
  separate formalizations for hand-object and object-surface
  attachments, or can there be one unified
  formalization?)\\
  \textbf{Answer.} Allowing a free translation along the axis of a cylindrical
  object is possible using flags. However, the user needs to manually add
  inequality constraints to limit this translation. We have added this case in
  the paper with an explanation. Limits of translation are automatically
  defined for contact surfaces by providing plane convex polygons. It seems
  to us difficult to merge these two types of constraints.

\item \textbf{Comment.} One nice trick in Simeon et al. [30] is the
  "virtual closed chain" motions when an object is both on support
  surface and is grasped. Is it possible to have similar solutions
  using the formalization here?\\
  \textbf{Answer.} The constraint graph formalism actually enables us to
  generalize the reduction property in [30]. For instance, the benchmark
  "Romeo placard" could be solved this way. To do so, we need to modify the
  transition constraints when building the constraint graph.

\item \textbf{Comment.} In the experimental results section (VII-B) it
  would be good to state the dimensionality of the problems
  (i.e. robot number of DoF and maybe even objects' number of DoF) as
  that would put the planning times a bit more in
  perspective.\\
  \textbf{Answer.} We have added the number of degrees of freedom to each
  benchmark.
\end{enumerate}

\subsection*{Associate editor}
\begin{enumerate}
\item \textbf{Comment.} In addition, given the importance of the
  software framework, the authors are advised to make sure to provide
  a link to the software repositories on github and to give a proper
  visibility in the manuscript.\\ \textbf{Answer.} There was already a link
  to the installation page, but we made it more visible in the introduction.
\end{enumerate}

\item \textbf{Comment.} Don't use citations as nouns like in "evoked
  in [2]". Rather use "evoked by Mirabel~\emph{et al.}~\cite{...}.
  Don't start a sentence with a citation like "[47] propose". This
  would be fixed by the above change.\\
  \textbf{answer.} We changed citations as requested.

\item \textbf{Comment.} 
Overall, this paper is substandard with respect to writing. The authors
are strongly encouraged to carefully revise the writing and spelling of
the paper. If this should not be corrected in the revised version, this
paper will not be recommended for publication despite the contributions
and this publication recommendation. The authors are encouraged to ask
a native speaker for advise.\\ \textbf{Answer.}
\end{document}

### 2021-09-09 review/reply-to-reviewers.tex
Some more corrections.

@@
 \maketitle
+
+First we would like to warmly thank the reviewers and associate editor for their
+time and valuable remarks. Below are answers to each of their comments.
 \subsection*{Reviewer 7}
@@
     initial version of the paper. To improve the paper, we have moved Figure 2
-    at the beginning of the paper to provide quickly an example of manipulation
+    at the beginning of the paper to quickly provide an example of manipulation
     planning problem.
@@
   \textbf{Answer.} Using $(\sin\theta,\cos\theta)$ makes the parameterization
-  continuous. We added a comment in the paper.
+  continuous. We have added a comment in the caption of Table I.
 
@@
   operator defined in equation (3)?
-  \textbf{Answer.} You are perfectly right. We fixed the error.
+  \textbf{Answer.} You are perfectly right. We have fixed the error.
 
@@
   \textbf{Answer.} You
-  are right, there are many hyperlinks in the paper. We checked them
-  one by one and found that they all work.
+  are right, there are many hyperlinks in the paper. We have checked them
+  one by one.
 
@@
   same config variable.  \\
-  \textbf{Answer.} As stated at the beginning of Paragraph IV.D.a, "when an
-  explicit constraint is not successfully added following Algorithm 1, it is
-  handled as an implicit constraint."
+
+  \textbf{Answer.} As stated at the beginning of Section IV.D.1:
+  "when an explicit constraint is not successfully added following
+  Algorithm 1, it is handled as an implicit constraint." If redundant
+  explicit constraints are inserted, the second one will be handled as
+  an implicit constraint. By substitution, in the system of implicit
+  constraints, some lines will be of the form $0=0$. Those lines will
+  be ignored when computing the pseudo-inverse of the Jacobian of the
+  implicit system. As a result, the solver will solve both constraints.
+  This will only affect the time of computation since the
+  size of the Jacobian is bigger than if only the first explicit
+  constraint had been inserted. The same is true with redundant
+  implicit constraints.
 
@@
   continuous part starting at the beginning of the path is used by the
-  M-RRT algorithm. We have added a remark in paragraph and in IV.E.1
+  M-RRT algorithm. We have added a remark in Section IV.E.1 and in
   Algorithm 4.
@@
   kinematic solution (that could be not unique)?\\
-  \textbf{Answer.}
-  The M-RRT algorithm as of today is not able to define the goal state
-  by some constraints, although it would not be difficult to modify
-  it for that. However, the python bindings enable users to quickly develop
-  new algorithms where the goal can be defined by some constraints. In the
-  construction set example described in Section VII.B.2) and in Figure 13, the
-  goal is a state of the constraint graph.
+  \textbf{Answer.}  The M-RRT algorithm as of today is not able to
+  define the goal state by some constraints, although it would not be
+  difficult to modify it for that. However, the python bindings enable
+  users to quickly develop new algorithms where the goal can be
+  defined by some constraints. In the construction set example
+  described in Section VII.B.2, page 17 and in Figure 13, the goal is
+  a state of the constraint graph.
 
@@
   equations), is $R^3 \times SO(3)$ the same as $SE(3)$?\\
-  \textbf{Answer.} They both represent rigid-body motion but they have different
-  Lie group operators. We have added a paragraph after Equation (13) to clarify
-  this point.
+  \textbf{Answer.} They both represent rigid-body transformation but
+  they have different Lie group operators. We have added a paragraph
+  after Equation (13) to clarify this point.
 
@@
 \item \textbf{Comment.} I think that Figure 9 is not cited in the text.\\
-  \textbf{Answer.} This has been fixed.
+  \textbf{Answer.} Figure 9 is now cited in Section VII.B.1, page 16.
 
@@
   is no collision?\\
-  \textbf{Answer.} The software platform proposes discretized collision checking
-  as well as continuous collision checking. We added a sentence and a reference
-  in the description of the framework. Collisions are checked also along
-  waypoint transitions. We added the word "almost" at the end of Section VII-B3). Since the sentence was confusing.
+  \textbf{Answer.} The software platform proposes discretized
+  collision checking as well as continuous collision checking. We have
+  added a sentence and a reference in the description of the
+  framework (Section VII, page 15). Collisions are checked also along waypoint
+  transitions. We have added the word "almost" at the end of Section
+  VII-B3, page 18 since the sentence was confusing.
 \end{enumerate}
@@
   related to representing manipulation planning problems with constraint graphs.\\
-  \textbf{Answer.} Although we had already cited 2 papers from this team, we
-  have added the requested reference.
+  \textbf{Answer.} We have added and cited the reference.
 
@@
   maybe the sentence should say "the starting (or first?) indices of
-  joint i">\\
-  \textbf{Answer.} You are right. This has been fixed.
+  joint i".\\
+  \textbf{Answer.} You are right. This has been fixed (page 4, right column).
 
@@
   \textbf{Answer.} "Path evaluation" means "computing the configuration at a
-  given parameter". We have added this definition in the paper.
+  given parameter". We have added the definition in the paper.
 
@@
   complete.\\
-  \textbf{Answer.} You are right. We added the absence of collision in the
+  \textbf{Answer.} You are right. We have added the absence of collision in the
   definition of admissible configurations.
@@
   configuration space is defined, but it seems to include only the
-  first and the last robot ($C_{r1}$ and $C_{rnr}$). Where are the other
-  robots in betwwen 1 and nr?\\ \textbf{Answer.} You are right. This has
+  first and the last robot ($C_{r_1}$ and $C_{r_{nr}}$). Where are the other
+  robots in between 1 and nr?\\ \textbf{Answer.} You are right. This has
   been fixed.
@@
   formalization?)\\
-  \textbf{Answer.} Allowing a free translation along the axis of a cylindrical
-  object is possible using flags. However, the user needs to manually add
-  inequality constraints to limit this translation. We have added this case in
-  the paper with an explanation. Limits of translation are automatically
-  defined for contact surfaces by providing plane convex polygons. It seems
-  to us difficult to merge these two types of constraints.
+  \textbf{Answer.} Allowing a free translation along the axis of a
+  cylindrical object is possible using flags. However, the user needs
+  to manually add inequality constraints to limit this translation. We
+  have added this case in the paper with an explanation (page
+  8). Limits of translation are automatically defined for contact
+  surfaces by providing plane convex polygons. It seems to us
+  difficult to merge these two types of constraints.
 
@@
   using the formalization here?\\
-  \textbf{Answer.} The constraint graph formalism actually enables us to
-  generalize the reduction property in [30]. For instance, the benchmark
-  "Romeo placard" could be solved this way. To do so, we need to modify the
-  transition constraints when building the constraint graph.
+  \textbf{Answer.} The constraint graph representation actually
+  enables us to generalize the reduction property in [30]. For
+  instance, the benchmark "Romeo placard" could be solved this way. To
+  do so, we would need to modify the transition constraints when
+  building the constraint graph. We have added this task as a future
+  work in the conclusion.
 
@@
   perspective.\\
-  \textbf{Answer.} We have added the number of degrees of freedom to each
-  benchmark.
+  \textbf{Answer.} We have added the number of degrees of freedom in the tables
+  compiling the experimental results for each benchmark.
 \end{enumerate}
@@
   a link to the software repositories on github and to give a proper
-  visibility in the manuscript.\\ \textbf{Answer.} There was already a link
-  to the installation page, but we made it more visible in the introduction.
-\end{enumerate}
+  visibility in the manuscript.\\ \textbf{Answer.} There was already a
+  link to the installation page, but we have made it more visible in
+  the introduction (page 1, left column).
 
 \item \textbf{Comment.} Don't use citations as nouns like in "evoked
-  in [2]". Rather use "evoked by Mirabel~\emph{et al.}~\cite{...}.
+  in [2]". Rather use "evoked by Mirabel~\textbackslash emph\{et al.\}~\textbackslash cite\{...\}".
   Don't start a sentence with a citation like "[47] propose". This
   would be fixed by the above change.\\
-  \textbf{answer.} We changed citations as requested.
+  \textbf{answer.} We have changed citations as requested.
 
-\item \textbf{Comment.} 
-Overall, this paper is substandard with respect to writing. The authors
-are strongly encouraged to carefully revise the writing and spelling of
-the paper. If this should not be corrected in the revised version, this
-paper will not be recommended for publication despite the contributions
-and this publication recommendation. The authors are encouraged to ask
-a native speaker for advise.\\ \textbf{Answer.}
+\item \textbf{Comment.}  Overall, this paper is substandard with
+  respect to writing. The authors are strongly encouraged to carefully
+  revise the writing and spelling of the paper. If this should not be
+  corrected in the revised version, this paper will not be recommended
+  for publication despite the contributions and this publication
+  recommendation. The authors are encouraged to ask a native speaker
+  for advise.\\ \textbf{Answer.} %You probably mean "for advice" ?
+  We have asked a professional interpreter to read and correct our
+  paper. The new version should be perfect.
+\end{enumerate}
 \end{document}

### 2021-09-09 hpp.tex
Some more corrections.

@@
     h &= (\bar{h}_1,\bar{h}_2,\bar{h}_4,\bar{h}_5),\\
-    h_{comp} &= (,\bar{h}_3,\bar{h}_6),
+    h_{comp} &= (\bar{h}_3,\bar{h}_6),
   \end{align*}
@@
 \href{https://github.com/humanoid-path-planner/hpp_benchmark/tree/v4.10.0/2020-07-23}{\texttt{hpp\_benchmark}} package. Here we only present a few test cases.
-The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with 32 Gigabytes of RAM and 3 Gigabytes of cache memory.
+The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with {\color{blue}32 Gigabytes of RAM and 3 Gigabytes of cache memory}.
 For each test case, we report the minimum, maximum, mean and standard deviation of the time of computation on the one hand, and of the number of nodes in the roadmap built to solve the problem on the other.
@@
 A good candidate is a generalization of RMR*~\cite{schmitt17icra}. Also we intend to focus on manipulation path optimization since paths computed by
-random algorithms are too long to be applied to real robots as such.
+random algorithms are too long to be applied to real robots as such. {\color{blue}Finally, we would like to generalize the reduction property proposed by Siméon
+  \textit{et al}~\cite{simeon2004ijrr} . The constraint graph representation is
+  a perfect tool for that.}

### 2021-10-15 hpp.tex
Last corrections before submission.

@@
 Today, robots in industrial manufacturing are mostly programmed by
-hand. They repeat the same motion thousands of times with a high
+hand. They repeat the same motion thousands of times with great
 accuracy. However, automating a task with some variability is very
@@
 \item a manipulation planning algorithm that tackles a great variety of manipulation planning problems,
-\item an open-source software suite that implements all the above following state-of-the-art development tools and methods.
+\item an open-source software suite that implements all the above, following state-of-the-art development tools and methods.
 \item a docker image of the aforementioned software with installation instructions provided with this paper. This image makes the experimental results replicable.
@@
 \item description of the configuration space as a Cartesian product of Lie groups (Section~\ref{partie:preliminary}),
-\item unified and detailed definition of the grasp and placement constraints that are only evoked in {\color{blue}Mirabel \textit{et al}}~\cite{mirabel-iros2016} (Section~\ref{sec:manipulation-problem}),
+\item unified and detailed definition of the grasp and placement constraints that are only mentioned in {\color{blue}Mirabel \textit{et al}}~\cite{mirabel-iros2016} (Section~\ref{sec:manipulation-problem}),
 \item automatic construction of the constraint graph (Section~\ref{sec:manipulation-problem}),
@@
 \item a description of the software platform (Section~\ref{sec:humanoid-path-planner}),
-\item experimental results on several different problems.
+\item experimental results for several different problems.
 \end{itemize}
@@
 \Partie~\ref{sec:manipulation-planning} provides a general algorithm that solves manipulation planning problems.
-Finally, \partie~\ref{sec:humanoid-path-planner} is devoted to the software platform implementing the notions introduced in the previous sections.
+Finally, \Partie~\ref{sec:humanoid-path-planner} is devoted to the software platform implementing the notions introduced in the previous sections.
 Experimental results are provided for a large variety of problems.
 
-Each \partie is implemented by one or several software packages.
-For some values that need to be computed, instead of providing formulas, we sometimes provide a link to the C++ or python implementation.
+Each \partie is implemented by one or several software packages. For some values that need to be computed, rather than providing formulas, we sometimes give a link to the C++ or python implementation.
 
@@
 
-Some systems are subject to nonlinear constraints. These nonlinear
-constraints define sub-manifolds of the configuration space the robot
+Some systems are subject to nonlinear constraints. These
+constraints define submanifolds of the configuration space the robot
 must stay on. For example, legged robots that must keep contact with
@@
 robot.  The accessible configuration space is thus a union of
-sub-manifolds as defined in the previous section. Each of these
+submanifolds as defined in the previous section. Each of these
 manifolds may moreover be a foliation where each leaf corresponds to a
@@
 \cite{stilman2007manipulation}, \cite{srivastavaetal2014icra}, \cite{nieuwenhuisen2008effective},
-\cite{ota2004rearrangement}. They mostly use two-level methods
+\cite{ota2004rearrangement}. They mainly rely on two-level methods
 composed of a symbolic task planner and of a motion
@@
 \end{itemize}
-$SO(n)$ and $SE(n)$ respectively stand for \textit{special orthogonal group} and \textit{special Euclidean group}.
+$SO(n)$ and $SE(n)$ stand for \textit{special orthogonal group} and \textit{special Euclidean group} respectively.
 They represent the group of rotations and the group of rigid-body transformations in $\reals^n$.
@@
 
-The joint configuration spaces enumerated in the previous paragraph: $\mathbb{R}^n$, $SO(n)$, and $SE(n)$ are all Lie groups.
+The joint configuration spaces listed in the previous paragraph: $\mathbb{R}^n$, $SO(n)$, and $SE(n)$ are all Lie groups.
 The group operation is $+$ for $\mathbb{R}^n$, and composition denoted as $"."$ for $SE(n)$.
 We refer to~{\color{blue}Murray \textit{et al}}~\cite[Appendix A]{LMS94} for a thorough definition of Lie groups.
-Here we detail only properties that are useful for the following developments.
+Here we detail only those properties that are useful for the following developments.
 
@@
 Note that $\mathbf{x}$ is also the coordinate vector of $g(\mathbf{x})$ in the moving frame $g$.
-If $\mathbf{v}$, $\omega$ are a linear and angular velocities at the origin, $(\mathbf{v},\omega)$ is \textit{transported} to $g$
+If $\mathbf{v}$, $\omega$ are linear and angular velocities at the origin, $(\mathbf{v},\omega)$ is \textit{transported} to $g$
   as the same linear and angular velocities expressed in the moving frame. In other words, if
@@
 \end{itemize}
-$SE(2)$, $SO(3)$, and $SO(2)$ are (or are isomorphic to) subgroups of $SE(3)$ and follow the same geometrical interpretation.
+$SE(2)$, $SO(3)$, and $SO(2)$ are subgroups of $SE(3)$ and follow the same geometrical interpretation.
 
@@
 
-\section{nonlinear constraints and solvers}\label{sec:constraints}
+\section{Nonlinear constraints and solvers}\label{sec:constraints}
 
@@
 
-\subsection{nonlinear constraints}
+\subsection{Nonlinear constraints}
 
@@
 $$
-It may be useful to use a non-zero right hand side for the same function $h$. We define for that parameterized nonlinear constraints.
+It may be useful to use a non-zero right hand side for the same function $h$. For that we define parameterized nonlinear constraints.
 \begin{definition}\textbf{Parameterized nonlinear constraint.}
@@
 
-This operator is represented by a matrix with $nv_2$ lines and $nv_1$ columns, where $nv_1$ and $nv_2$ are the dimensions of the tangent spaces of respectively $\liegroup_1$ and $\liegroup_2$.
-
-\subsection{Newton based solver}
+This operator is represented by a matrix with $nv_2$ lines and $nv_1$ columns, where $nv_1$ and $nv_2$ are the dimensions of the tangent spaces of $\liegroup_1$ and $\liegroup_2$ respectively.
+
+\subsection{Newton-based solver}
 
 It is sometimes useful to produce a configuration $\conf$ that satisfies a constraint (or a set of constraints) of type~(\ref{eq:constraint}) from a configuration $\conf_0$ that does not.
-This action is called the \textit{projection of $\conf_0$ onto the sub-manifold defined by the constraint} and is performed by a Gauss-Newton solver~\cite[Chapter~10]{NoceWrig06} that iteratively linearizes the constraint as follows:
+This action is called the \textit{projection of $\conf_0$ onto the submanifold defined by the constraint} and is performed by a Gauss-Newton solver~\cite[Chapter~10]{NoceWrig06} that iteratively linearizes the constraint as follows:
 $$
@@
 
-In manipulation planning applications, in which robots manipulate objects, once an object is grasped, the position of the object can be explicitly computed from the configuration of the robot.
+In manipulation planning applications in which robots manipulate objects, once an object is grasped, the position of the object can be explicitly computed from the configuration of the robot.
 In this case, some configuration variables of the system depend on other configuration variables:
@@
 \item $m$ and $n$ are two integers,
-\item $M$ and $N$ are two subsets of respectively $I_{m}$ and $I_{n}$,
+\item $M$ and $N$ are two subsets of $I_{m}$ and $I_{n}$ respectively,
 \item $J$ is a matrix with $m$ rows and $n$ columns,
@@
 
-To optimize constraint resolution, we perform variable substitution whenever possible in order to reduce both the number of variables and the dimension of the resulting implicit constraint.
+To optimize constraint resolution, we perform variable substitution whenever possible in order to reduce the number of variables as well as the dimension of the resulting implicit constraint.
 Here we describe the method first published in~{\color{blue}Mirabel \textit{et al}}~\cite{MirLam2018}.
@@
 $nc$ is the size of the latter.
-$args$ is an array that stores for each configuration variable, the index in $explicit$ of the constraint that computes this configuration variable, -1 if no constraint computes the index.
+$args$ is an array that, for each configuration variable stores the index in $explicit$ of the constraint that computes this configuration variable, -1 if no constraint computes the index.
 Procedure \href{https://github.com/humanoid-path-planner/hpp-constraints/blob/2555dfec945575f824bd973c30c5c13ec3b67645/src/explicit-constraint-set.cc\#L144}{ADD} tests whether explicit constraint $E$ is compatible with the previously inserted constraints.
@@
 The Loop at line 9 recursively checks that any element of $out$ is not an input variable of a previously inserted constraint.
-If the loop ends without returning failure,  line 18 stores that elements of $out$ are computed by $E$ and $E$ is inserted in the vector of constraints.
+If the loop ends without returning failure,  line 18 stores the information that elements of $out$ are computed by $E$ and $E$ is inserted in the vector of constraints.
 Function \href{https://github.com/humanoid-path-planner/hpp-constraints/blob/2555dfec945575f824bd973c30c5c13ec3b67645/src/explicit-constraint-set.cc\#L301}{computeOrder} at line 20 recursively computes the order in which the explicit constraints are evaluated, following the rule that the input of a constraint should be evaluated before the output.}
@@
 
-Let us now detail the iterative computation of~(\ref{eq:jacobian-product}). Let $J$ be the product of of $J_j$ for $j$ from $nc$ to $i+1$.
+Let us now detail the iterative computation of~(\ref{eq:jacobian-product}). Let $J$ be the product of $J_j$ for $j$ from $nc$ to $i+1$.
 Note that if $J_i$ and $J$ are square matrices of size $nv$, of the form~(\ref{eq:jacobian-explicit}), $J_i.J$ can be computed by block as follows:
@@
 
-Now that we are able to project configurations onto sub-manifolds defined by numerical constraints -- up to some numerical threshold, we need to define paths on such sub-manifolds.
+Now that we are able to project configurations onto submanifolds defined by numerical constraints -- up to some numerical threshold, we need to define paths on such submanifolds.
 The usual way of doing so is by discretizing the path and projecting each sample configuration.
@@
 Instead, we propose an alternative architecture where paths store the constraints they are subject to and apply the constraints at path evaluation {\color{blue}(\textit{i.e.} when computing the configuration at a given parameter)}.
-Let $P\in C^1([0,T],\CS)$ be a path without constraint defined on an interval $[0,T]$, and $\mathbf{proj}$ a projector onto a sub-manifold defined by numerical constraints (\textit{i.e.} an instance of\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}).
+Let $P\in C^1([0,T],\CS)$ be a path without constraint defined on an interval $[0,T]$, and $\mathbf{proj}$ a projector onto a submanifold defined by numerical constraints (\textit{i.e.} an instance of\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}).
 
@@
 
-The previous sections have presented how we model kinematic chains, configurations and velocities for a given robotic system, how configurations and paths can be projected onto a sub-manifold of the configuration space defined by numerical constraints.
+The previous sections have presented how we model kinematic chains, configurations and velocities for a given robotic system and how configurations and paths can be projected onto a submanifold of the configuration space defined by numerical constraints.
 
@@
 $$
-where $nr$ is the number of robots, $no$ is the number of objects, $\CS_{r_i},\ i\in\{1,...,nr\}$ is the configuration space of robot $r_i$.
+where $nr$ is the number of robots, $no$ the number of objects, $\CS_{r_i},\ i\in\{1,...,nr\}$ is the configuration space of robot $r_i$.
 \begin{definition}\textbf{Gripper.}\label{def:gripper}
@@
 \end{itemize}
-If $\conf_0$ is a configuration satisfying the grasp constraint: $h(\conf_0)$=0, then the sub-manifold defined by
+If $\conf_0$ is a configuration satisfying the grasp constraint: $h(\conf_0)$=0, then the submanifold defined by
 $$
@@
 
-\subsection{Stable Contact pose}\label{subsec:placement}
+\subsection{Stable contact pose}\label{subsec:placement}
 
@@
 \item defining virtual grippers in the environment and virtual handles on the object, implicitly defines a discrete set of poses,
-\item defining a virtual gripper on a horizontal plane and a virtual handle on the object, and using a grasp with flags $(\false,\false,\true,\true,\true,\false)$ constrains the object to move on an infinite horizontal plane.
+\item defining a virtual gripper on a horizontal plane and a virtual handle on the object, and using a grasp with flags $(\false,\false,\true,\true,\true,\false)$ constrains the object to move along an infinite horizontal plane.
 \end{enumerate}
 
-Here we propose a third method that enables users to define contact surfaces in a more flexible way. In this aim, we denote by:
+Here we propose a third method that enables users to define contact surfaces in a more flexible way. To this end, we denote by:
 \begin{itemize}
@@
 
-As for grasps, we need to define a parameterized complement constraint for the contact constraint in order to specify the sub-manifold of configurations that are reachable from one configuration while keeping the object in a constant stable pose.
+As for grasps, we need to define a parameterized complement constraint for the contact constraint in order to specify the submanifold of configurations reachable from one configuration while keeping the object in a constant stable pose.
 The naive way consists in defining
@@
 Let $\conf_0$ be a configuration such that $h(\conf_0)=0$.
-The sub-manifold defined by
+The submanifold defined by
 \begin{equation}\label{eq:stable-pose}
@@
 \end{itemize}
-With this definition, the sub-manifold defined by~(\ref{eq:stable-pose}), (\ref{eq:convex-shape-contact}), (\ref{eq:convex-shape-contact-complement}) contains configurations where the object is in a unique stable pose.
+With this definition, the submanifold defined by~(\ref{eq:stable-pose}), (\ref{eq:convex-shape-contact}), (\ref{eq:convex-shape-contact-complement}) contains configurations where the object is in a unique stable pose.
 The polygon indices $i$ and $j$, as well as their relative position can indeed be recovered from~(\ref{eq:convex-shape-contact-complement}):
@@
 
-\subsection{Constraint Graph}\label{subsec:constraint graph}
-
-According to Definition~\ref{def:manipulation-problem}, the set of admissible configurations of a manipulation problem is the union of sub-manifolds of the configuration space of the system.
-Each sub-manifold is defined by grasp and/or stable contact constraints.
-We call each sub-manifold a \textit{state} of the problem.
+\subsection{Constraint graph}\label{subsec:constraint graph}
+
+According to Definition~\ref{def:manipulation-problem}, the set of admissible configurations of a manipulation problem is the union of submanifolds of the configuration space of the system.
+Each submanifold is defined by grasp and/or stable contact constraints.
+We call each submanifold a \textit{state} of the problem.
 
@@
   \end{algorithmic}
-  \caption{Method \CREATESTATE builds the constraints relative to a state: one grasp constraint for each grasp, and one placement constraint for each object non grasped.
+  \caption{Method \CREATESTATE builds the constraints relative to a state: one grasp constraint for each grasp, and one placement constraint for each object not grasped.
 \CREATETRANSITION builds the constraints relative to a transition: the constraints of the initial state (with the fewest grasps) and their complements.}
@@
 First, it checks whether there exists a transition between the states.
-If so, it checks that the right hand side of the transition parameterized constraints is the same for both configurations (up to the error threshold).
-Then it returns the linear interpolation between the configurations, projected onto the sub-manifold defined by the transition constraints.
+If so, it checks that the right hand side of the parameterized constraints of the transition is the same for both configurations (up to the error threshold).
+Then it returns the linear interpolation between the configurations, projected onto the submanifold defined by the transition constraints.
 If the path is in collision {\color{blue}or discontinuous}, only the {\color{blue}continuous} collision-free part at the beginning of the path is returned.
@@
 Function \EXTEND attempts to generate a path from a configuration in a state to another state following a random transition.
-Similarly as for function \CONNECT, the path is projected onto the sub-manifold defined by the transition constraints.
+Similarly as for function \CONNECT, the path is projected onto the submanifold defined by the transition constraints.
 The end configuration is obtained by applying to the random configuration the constraints of the transition and of the goal state.
@@
 In this section, the algorithm described in the previous section is depicted with two examples.
-Figure~\ref{fig:MRRT-extension} shows function \EXTEND defined in the previous section on the example of Figure~\ref{fig:2ur3-cylinder}.
+Figure~\ref{fig:MRRT-extension} shows function \EXTEND defined in the previous section applied to the example of Figure~\ref{fig:2ur3-cylinder}.
 The system considered is
@@
 Section~\ref{subsec:constraint graph} by a sequence of intermediate
-states and transitions as follows: given
+states and transitions: given
 Definition~\ref{def:neighboring-states}, if two states $\state_1$
@@
 $i\in\{1,\cdots\ngrippers\}$, $j\in\{1,\cdots\nhandles\}$. Let us denote by
-$\object{}$ the object handle $\handle_j$ belongs to. Then either
+$\object{}$ the object to which handle $\handle_j$ belongs. Then either
 \begin{enumerate}
@@
 transition from an initial configuration by projecting the configuration
-onto the sub-manifold defined by the goal state constraints and by the
+onto the submanifold defined by the goal state constraints and by the
 transition constraints. The right hand side of the transition constraint is
@@
 \item \texttt{hpp-core} a library that implements most of the concepts relative
-  to motion planning. The main features are
+  to motion planning. The main features are:
   \begin{itemize}
@@
 The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with {\color{blue}32 Gigabytes of RAM and 3 Gigabytes of cache memory}.
-For each test case, we report the minimum, maximum, mean and standard deviation of the time of computation on the one hand, and of the number of nodes in the roadmap built to solve the problem on the other.
+For each test case, we report the minimum, maximum, mean and standard deviation of the time of computation on the one hand, and of the number of nodes in the roadmap built to solve the problem, on the other.
 
@@
 computed and only those states (nine in total) are built in the
-constraint graph. Then, a sequence of sub-goals in the successive
-states is computed, in such a way that each sub-goal is accessible by
+constraint graph. Then, a sequence of subgoals in the successive
+states is computed, in such a way that each subgoal is accessible by
 the previous one (on the same leaf of the corresponding transition
-foliation). The sub-goals are then linked by running a constrained visibility
+foliation). The subgoals are then linked by running a constrained visibility
 PRM algorithm \cite{SimLauNis00} on each leaf. The python code can be found

### 2021-11-18 hpp.tex
Final version before submission
  Remove color blue for text that had been added,
  Add bios.

@@
   \end{center}
-  \caption{{\color{blue} Example of manipulation planning problem.} Top: two UR3 robots with one gripper each (X=red, Y=green, Z=blue) manipulating a cylinder with two handles.
+  \caption{{ Example of manipulation planning problem.} Top: two UR3 robots with one gripper each (X=red, Y=green, Z=blue) manipulating a cylinder with two handles.
 The environment contains one rectangular contact surface (in red).
@@
 Therefore this \paper describes a software platform called {Humanoid Path
-  Planner} tailored for manipulation planning in robotics. It can handle many types of robots, from manipulator arms to legged humanoid robots.
+  Planner} tailored for manipulation planning in robotics.
+It can handle many types of robots, from manipulator arms to legged humanoid robots.
+Figure~\ref{fig:2ur3-cylinder} displays an example of manipulation problem.
 The main contributions are:
@@
 \end{itemize}
-{\color{blue} Installation instructions can be found at \href{https://humanoid-path-planner.github.io/hpp-doc}{https://humanoid-path-planner.github.io/hpp-doc}.}
-This paper extends the work presented in {\color{blue} previous papers }\cite{mirabel-iros2016},\cite{MirLam2018} with the following new material:
+{ Installation instructions can be found at \href{https://humanoid-path-planner.github.io/hpp-doc}{https://humanoid-path-planner.github.io/hpp-doc}.}
+This paper extends the work presented in { previous papers }\cite{mirabel-iros2016},\cite{MirLam2018} with the following new material:
 \begin{itemize}
 \item description of the configuration space as a Cartesian product of Lie groups (Section~\ref{partie:preliminary}),
-\item unified and detailed definition of the grasp and placement constraints that are only mentioned in {\color{blue}Mirabel \textit{et al}}~\cite{mirabel-iros2016} (Section~\ref{sec:manipulation-problem}),
+\item unified and detailed definition of the grasp and placement constraints that are only mentioned in {Mirabel \textit{et al}}~\cite{mirabel-iros2016} (Section~\ref{sec:manipulation-problem}),
 \item automatic construction of the constraint graph (Section~\ref{sec:manipulation-problem}),
@@
 \Partie~\ref{partie:preliminary} introduces some preliminary notions like kinematic chains and Lie groups that are used to model the configuration space of each joint.
-\Partie~\ref{sec:constraints} introduces nonlinear constraints and solvers that are at the core of manipulation problem definition.
+\Partie~\ref{sec:constraints} introduces nonlinear constraints and solvers that are at the core of the manipulation problem definition.
 \Partie~\ref{sec:manipulation-problem} defines the problem of prehensile manipulation in the general setting.
@@
 probability. To sample configurations on the constrained manifold,
-{\color{blue}Dalibard \textit{et al}}~\cite{DalNakLamLau2009} and {\color{blue}Benrenson
+{Dalibard \textit{et al}}~\cite{DalNakLamLau2009} and {Benrenson
 \textit{et al}}~\cite{berenson2011} project random configurations using
@@
 configuration variables with respect to others~\cite{Cortes2002},
-\cite{MirLam2018} whenever this can be done. {\color{blue}Jaillet \textit{et al}}~\cite{atlasRRT} propose
+\cite{MirLam2018} whenever this can be done. {Jaillet \textit{et al}}~\cite{atlasRRT} propose
 another method based on nonlinear projection. They cover the
@@
 projection of a uniform distribution over the configuration
-space. {\color{blue}Beobkyoon \textit{et al}}~\cite{tb-rrt2016} propose a variation of the latter paper.  The
+space. {Beobkyoon \textit{et al}}~\cite{tb-rrt2016} propose a variation of the latter paper.  The
 main difference lies in the fact that the nodes built on the
 tangent space are not immediately projected onto the manifold.
-{\color{blue}Cefalo \textit{et al}}~\cite{CefOri2019} put forward a general framework to plan task-constrained
-motions in the presence of moving obstacles. {\color{blue}Kingston \textit{et al}}~\cite{KinMolKav2019}
+{Cefalo \textit{et al}}~\cite{CefOri2019} put forward a general framework to plan task-constrained
+motions in the presence of moving obstacles. {Kingston \textit{et al}}~\cite{KinMolKav2019}
 provide an in-depth review of the various approaches to motion
@@
 The first attempt to solve manipulation planning problems using random sampling
-was proposed by {\color{blue}Siméon \textit{et al}}~\cite{simeon2004ijrr} where a reduction property simplifies the problem.
+was proposed by {Siméon \textit{et al}}~\cite{simeon2004ijrr} where a reduction property simplifies the problem.
 
@@
 
-{\color{blue}Schmitt \textit{et al}}~\cite{schmitt19icra} propose an approach where two robots manipulate an object in a dynamic environment.
+{Schmitt \textit{et al}}~\cite{schmitt19icra} propose an approach where two robots manipulate an object in a dynamic environment.
 The output of the algorithm is a sequence of controllers rather than a sequence of paths.
 
-Our work shares many ideas with {\color{blue}Hauser and Ng-Thow-Hing} \cite{HauNgt2011} where the notion of constraint
+Our work shares many ideas with {Hauser and Ng-Thow-Hing} \cite{HauNgt2011} where the notion of constraint
 graph is present, although not as clearly expressed as in this paper. The main
@@
 Undoubtedly the most popular one is OMPL \cite{ompl2012} which integrates many randomized path planning algorithms and is widely used for teaching purposes.
-Recently, {\color{blue}Kingston \textit{et al}}~\cite{KinMolKav2019} proposed an extension for systems subject to nonlinear constraints.
+Recently, {Kingston \textit{et al}}~\cite{KinMolKav2019} proposed an extension for systems subject to nonlinear constraints.
 
@@
 The group operation is $+$ for $\mathbb{R}^n$, and composition denoted as $"."$ for $SE(n)$.
-We refer to~{\color{blue}Murray \textit{et al}}~\cite[Appendix A]{LMS94} for a thorough definition of Lie groups.
+We refer to~{Murray \textit{et al}}~\cite[Appendix A]{LMS94} for a thorough definition of Lie groups.
 Here we detail only those properties that are useful for the following developments.
@@
 Notice that the dimensions of the configuration representation and of the velocity representation may differ.
-{\color{blue} Using $(\cos\theta,\sin\theta)$ instead of $\theta$ for $SO(2)$ and $SE(2)$ makes the parameterization continuous when $\theta$ discontinuously switches from $-\pi$ to $\pi$}}
+{ Using $(\cos\theta,\sin\theta)$ instead of $\theta$ for $SO(2)$ and $SE(2)$ makes the parameterization continuous when $\theta$ discontinuously switches from $-\pi$ to $\pi$}}
   \label{tab:vector representation}
@@
 $$
-We denote by $iq_i$, and $iv_i$ the {\color{blue}starting} indices of joint $i$ in the robot configuration and velocity vectors.
+We denote by $iq_i$, and $iv_i$ the {starting} indices of joint $i$ in the robot configuration and velocity vectors.
 $$
@@
 $$
-\conf(t) = \conf_0 \oplus t (\conf_1 {\color{blue}\ominus} \conf_0)
+\conf(t) = \conf_0 \oplus t (\conf_1 {\ominus} \conf_0)
 $$
@@
 
-\begin{definition}\textbf{nonlinear constraint.} A nonlinear constraint is defined by a {\color{blue} piece-wise} differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and is written
+\begin{definition}\textbf{nonlinear constraint.} A nonlinear constraint is defined by a { piece-wise} differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and is written
 \begin{equation}\label{eq:constraint}
@@
 \begin{definition}\textbf{Parameterized nonlinear constraint.}
-  A parameterized nonlinear constraint is defined by a {\color{blue} piece-wise} differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and by a vector $\mathbf{h}_0$ of $\reals^m$ and is written
+  A parameterized nonlinear constraint is defined by a { piece-wise} differentiable mapping $h$ from $\CS$ to a vector space $\reals^m$ and by a vector $\mathbf{h}_0$ of $\reals^m$ and is written
   $$
@@
 %% begin impl
-{\color{blue} Piece-wise} differentiable mappings are represented by abstract Class\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1DifferentiableFunction.html}{\texttt{DifferentiableFunction}}.
+{ Piece-wise} differentiable mappings are represented by abstract Class\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1DifferentiableFunction.html}{\texttt{DifferentiableFunction}}.
 %% end impl
@@
 In this \paper, we will make use of the term Jacobian in a generalized way.
-If $h$ is a {\color{blue} piece-wise} differentiable function from a Lie group $\liegroup_1$ to a Lie group $\liegroup_2$, and $\conf_1$ an element of $\liegroup_1$, we will denote by $\frac{\partial h}{\partial\conf}(\conf_1)$ the operator that maps velocities in $T_{\conf_1}\liegroup_1$ to the velocity in $T_{h(\conf_1)}\liegroup_2$ \textit{transported by} $h$\footnote{If $\dconf\in T_{\conf_1}\liegroup_1$ is a velocity along a time parameterized curve $\gamma$, $\frac{\partial h}{\partial\conf}(\conf_1)\dconf$ is the velocity along Curve $h(\gamma)$.}.
+If $h$ is a { piece-wise} differentiable function from a Lie group $\liegroup_1$ to a Lie group $\liegroup_2$, and $\conf_1$ an element of $\liegroup_1$, we will denote by $\frac{\partial h}{\partial\conf}(\conf_1)$ the operator that maps velocities in $T_{\conf_1}\liegroup_1$ to the velocity in $T_{h(\conf_1)}\liegroup_2$ \textit{transported by} $h$\footnote{If $\dconf\in T_{\conf_1}\liegroup_1$ is a velocity along a time parameterized curve $\gamma$, $\frac{\partial h}{\partial\conf}(\conf_1)\dconf$ is the velocity along Curve $h(\gamma)$.}.
 
@@
 $$
-h(\conf_{i+1}) \approx h(\conf_i) + \frac{\partial h}{\partial \conf}(\conf_{i})(\conf_{i+1}{\color{blue}\ominus}\conf_i) = 0
+h(\conf_{i+1}) \approx h(\conf_i) + \frac{\partial h}{\partial \conf}(\conf_{i})(\conf_{i+1}{\ominus}\conf_i) = 0
 $$
@@
 \begin{equation}\label{eq:HierarchicalIterative}
-\conf_{i+1} = \conf_i{\color{blue}\ominus}\alpha_i\frac{\partial h}{\partial \conf}^{+}(\conf_{i}) h(\conf_i)
+\conf_{i+1} = \conf_i{\ominus}\alpha_i\frac{\partial h}{\partial \conf}^{+}(\conf_{i}) h(\conf_i)
 \end{equation}
@@
 Class\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1HierarchicalIterative.html}{\texttt{HierarchicalIterative}} implements the above Newton method.
-{\color{blue}
+{
 Several line search methods are implemented:
@@
 To optimize constraint resolution, we perform variable substitution whenever possible in order to reduce the number of variables as well as the dimension of the resulting implicit constraint.
-Here we describe the method first published in~{\color{blue}Mirabel \textit{et al}}~\cite{MirLam2018}.
+Here we describe the method first published in~{Mirabel \textit{et al}}~\cite{MirLam2018}.
 Unlike in the former paper, the description we give in Algorithm~\ref{alg:ExplicitConstraintSet} is closer to the real implementation.
@@
 
-Instead, we propose an alternative architecture where paths store the constraints they are subject to and apply the constraints at path evaluation {\color{blue}(\textit{i.e.} when computing the configuration at a given parameter)}.
+Instead, we propose an alternative architecture where paths store the constraints they are subject to and apply the constraints at path evaluation {(\textit{i.e.} when computing the configuration at a given parameter)}.
 Let $P\in C^1([0,T],\CS)$ be a path without constraint defined on an interval $[0,T]$, and $\mathbf{proj}$ a projector onto a submanifold defined by numerical constraints (\textit{i.e.} an instance of\\ \href{https://gepettoweb.laas.fr/hpp/hpp-constraints/doxygen-html/classhpp_1_1constraints_1_1solver_1_1BySubstitution.html}{\texttt{SolverBySubstitution}}).
@@
 Thus, before inserting a projected path in a roadmap for example, it is necessary to detect possible discontinuities.
-{\color{blue}Hauser} \cite{Hauser-RSS-13} proposes a solution to this problem.
-{\color{blue}In a previous paper} \cite{mirabel:hal-01360409}, {\color{blue}we} described two algorithms to check whether a projected path is continuous.
+{Hauser} \cite{Hauser-RSS-13} proposes a solution to this problem.
+{In a previous paper} \cite{mirabel:hal-01360409}, {we} described two algorithms to check whether a projected path is continuous.
 These algorithms are implemented by
 classes \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1pathProjector_1_1Dichotomy.html}{\texttt{pathProjector::Dichotomy}} and \href{https://gepettoweb.laas.fr/hpp/hpp-core/doxygen-html/classhpp_1_1core_1_1pathProjector_1_1Progressive.html}{\texttt{pathProjector::Progressive}}.
-{\color{blue} Note that when a path is not continuous, the algorithms return a continuous portion of the path starting at the beginning of the path.
+{ Note that when a path is not continuous, the algorithms return a continuous portion of the path starting at the beginning of the path.
 This enables function \EXTEND in Algorithm~\ref{algo:M-RRT} to create a new node.}
@@
   \item each object is either grasped by a robot, or lies in a stable contact pose,
-  \item {\color{blue} the volumes occupied by the links of the robots and by
+  \item { the volumes occupied by the links of the robots and by
     the objects are pair-wise disjoint.}
@@
 $$
-\CS = \CS_{r_1}{\color{blue}\times\cdots}\times\CS_{r_{nr}}\times SE(3)^{no}
+\CS = \CS_{r_1}{\times\cdots}\times\CS_{r_{nr}}\times SE(3)^{no}
 $$
@@
 \end{definition}
-    {\color{blue}Note that $\reals^3\times SO(3)$ and $SE(3)$ have
+    {Note that $\reals^3\times SO(3)$ and $SE(3)$ have
       different group operators, exponential maps and
@@
   \end{align*}
-\item {\color{blue}if $flags=(\true,\true,\false,\true,\true,\false)$
+\item {if $flags=(\true,\true,\false,\true,\true,\false)$
   the grasp is satisfied iff the center of $\handle(\conf)$ is on the
@@
 Then it returns the linear interpolation between the configurations, projected onto the submanifold defined by the transition constraints.
-If the path is in collision {\color{blue}or discontinuous}, only the {\color{blue}continuous} collision-free part at the beginning of the path is returned.
+If the path is in collision {or discontinuous}, only the {continuous} collision-free part at the beginning of the path is returned.
 
@@
     \EndIf
-    \If {$p$ collision-free \color{blue}and continuous} {\Return $p$}
-    \Else { \Return collision-free {\color{blue}continuous} portion of $p$ starting at $\conf_{near}$}
+    \If {$p$ collision-free and continuous} {\Return $p$}
+    \Else { \Return collision-free {continuous} portion of $p$ starting at $\conf_{near}$}
     \EndIf
@@
     \textbf{constraints} \\
-    {\color{blue}$place_1(\conf)=0$} \\
-    {\color{blue}$place_1/comp(\conf) = place_1/comp(\conf_{near})$} \\
+    {$place_1(\conf)=0$} \\
+    {$place_1/comp(\conf) = place_1/comp(\conf_{near})$} \\
     $pregrasp_{11}(\conf) = 0$ \\
@@
     \textbf{constraints} \\
-    {\color{blue}$place_1(\conf)=0$} \\
-    {\color{blue}$place_1/comp(\conf) = place_1/comp(\conf_{near})$} \\
+    {$place_1(\conf)=0$} \\
+    {$place_1/comp(\conf) = place_1/comp(\conf_{near})$} \\
     $grasp_{11}(\conf) = 0$ \\
@@
     \textbf{constraints} \\
-    {\color{blue}$grasp_{11}(\conf)=0$} \\
-    {\color{blue}$grasp_{11}/comp(\conf) = grasp_{11}/comp(\conf_{2})$} \\
+    {$grasp_{11}(\conf)=0$} \\
+    {$grasp_{11}/comp(\conf) = grasp_{11}/comp(\conf_{2})$} \\
     $preplace_{1}(\conf) = 0$\\
@@
   \item roadmaps,
-  \item validation of configurations and paths, {\color{blue} notice that this
+  \item validation of configurations and paths, { notice that this
     includes an implementation of continuous collision checking first proposed
@@
 \href{https://github.com/humanoid-path-planner/hpp_benchmark/tree/v4.10.0/2020-07-23}{\texttt{hpp\_benchmark}} package. Here we only present a few test cases.
-The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with {\color{blue}32 Gigabytes of RAM and 3 Gigabytes of cache memory}.
+The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with {32 Gigabytes of RAM and 9 Megabytes of cache memory}.
 For each test case, we report the minimum, maximum, mean and standard deviation of the time of computation on the one hand, and of the number of nodes in the roadmap built to solve the problem, on the other.
@@
 The goal configuration is shown on the right.
-The algorithm is a constrained RRT close to the one described in {\color{blue}Dalibard \textit{et al}}~\cite{DalNakLamLau2009}}.
+The algorithm is a constrained RRT close to the one described in {Dalibard \textit{et al}}~\cite{DalNakLamLau2009}}.
   \label{fig:benchmark-hrp2}
@@
 One test case concerns constrained motion planning.
-The robot is an HRP-2 humanoid robot in quasi-static equilibrium that can slide on the ground ( {\color{blue}Figure~\ref{fig:benchmark-hrp2}} ).
-This type of motion can be post-processed into a walking motion using the method described in {\color{blue}Dalibard \textit{et al}}~\cite{dalibard:hal-00654175}.
+The robot is an HRP-2 humanoid robot in quasi-static equilibrium that can slide on the ground ( {Figure~\ref{fig:benchmark-hrp2}} ).
+This type of motion can be post-processed into a walking motion using the method described in {Dalibard \textit{et al}}~\cite{dalibard:hal-00654175}.
 The results are displayed in Table~\ref{tab:benchmark-hrp2}.
@@
   \end{center}
-  \caption{Experimental results for HRP-2 sliding on the ground {\color{blue}(36 degrees of freedom)}: time of computation and number of nodes.}
+  \caption{Experimental results for HRP-2 sliding on the ground {(36 degrees of freedom)}: time of computation and number of nodes.}
   \label{tab:benchmark-hrp2}
@@
   \end{center}
-  \caption{Experimental results for Baxter robot manipulating two boxes on a table {\color{blue}(31 degrees of freedom)}.}
+  \caption{Experimental results for Baxter robot manipulating two boxes on a table {(31 degrees of freedom)}.}
   \label{tab:baxter-manipulation-boxes}
@@
   \caption{Experimental results for PR-2 robot manipulating a box on a table
-     {\color{blue}(39 degrees of freedom)}.}
+     {(39 degrees of freedom)}.}
   \label{tab:pr2-manipulation-two-hands}
@@
   \caption{Experimental results for Romeo robot manipulating a placard
-    {\color{blue}(67 degrees of freedom)}.}
+    {(67 degrees of freedom)}.}
   \label{tab:romeo-placard}
@@
   \caption{Experimental results for construction set assembly
-    {\color{blue}(36 degrees of freedom)}.}
+    {(36 degrees of freedom)}.}
   \label{tab:construction-set}
@@
 easier to reach. The transition between the pregrasp waypoint and the
-grasp $\cap$ placement waypoint is {\color{blue}almost} always collision-free.
+grasp $\cap$ placement waypoint is {almost} always collision-free.
 
@@
 A good candidate is a generalization of RMR*~\cite{schmitt17icra}. Also we intend to focus on manipulation path optimization since paths computed by
-random algorithms are too long to be applied to real robots as such. {\color{blue}Finally, we would like to generalize the reduction property proposed by Siméon
+random algorithms are too long to be applied to real robots as such. {Finally, we would like to generalize the reduction property proposed by Siméon
   \textit{et al}~\cite{simeon2004ijrr} . The constraint graph representation is
